### 2020-11-17 paper_figures.Rmd
Initial commit

---
title: "Figures for Paper"
author: "Anna-Leigh Brown"
date: "06/10/2020"
output: 
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here")
```

```{r setup_function, echo=FALSE}
####helper funciton that gets the overall model p-value####
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

pretty_effect_plot = function(model, p_value_cutoff = 0.05, title = "Model 1"){
  # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
    p_value = lmp(model)
    r_squared = round(summary(model)$adj.r.squared,3)
    pretty_plot = broom::tidy(model) %>% 
        mutate(p_star = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  p.value > 0.1 ~ "NS")) %>% 
        filter(p.value < p_value_cutoff) %>%
        mutate(term = fct_reorder(term, -abs(estimate))) %>%
        ggplot(aes(x = term, y = estimate,fill = estimate > 0,show.legend = F)) +
        geom_col() +
        coord_flip() +
        ggtitle(title,subtitle = glue::glue("Model p-value: {p_value} \n Adj. R-squared: {r_squared}")) +
        geom_text(aes(x=term, y= -0.01, label=p_star)) 
}
```

# Summary statistics
First let's look at only including tissues with detectable stmn2 or UNC13A CE
```{r}
clean_data_table = fread(file.path(here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1)

print(glue::glue("Number of unique patients: {clean_data_table[,length(unique(participant_id))]}"))
print(glue::glue("Number of unique tissue samples: {clean_data_table[,length(unique(sample))]}"))
print("Patients Per Disease Category")
clean_data_table[,length(unique(participant_id)),by = disease]
print("Tissues Per Disease Category")
clean_data_table[,length(unique(sample)),by = disease]
print("Number of patients per rs12973192 genotype")
clean_data_table[,length(unique(participant_id)),by = call]
print("Number of tissues per disease")
clean_data_table[,.N,by = c("disease","tissue_clean")]
print("Number of partcipants by mutation and  disease")
clean_data_table[,length(unique(participant_id)),by = c("mutations","disease")]

print(glue::glue("Number of patients per pathology:"))
clean_data_table[,length(unique(participant_id)),by = .(pathology)]
```
# UNC13A cryptic is an event that is specific to TDP-43 proteinopathy

FTLD-non-TDP are those with TAU and FUS aggregates

Non-tdp ALS are those with SOD1 or FUS mutations. The n's are quite low on this 
unfortunately, only 8 ALS with SOD1 and 2 with FUS mutations.

```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(disease_group2 = case_when(disease == "Control" ~ "Control",
                                 pathology %in% c("FTD-TDP-A","FTD-TDP-B","FTD-TDP-C") ~ "FTLD-TDP",
                                 pathology %in% c("FTD-TAU","FTD-FUS") ~ "FTLD \n non-TDP",
                                 mutations %in% c("SOD1","FUS") ~ "ALS \n non-TDP",
                                 T ~ "ALS-TDP")) %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel 3'", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel 5'", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel 5'")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

```

Looking at disease tissue only, so just taking the cords in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.
```{r}

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals")

boxplot_table %>% 
    filter(disease_tissue == T) %>% 
    group_by(disease_group2) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
    ggplot(aes(x = detection_name, y = inclusion_reads * 10^6)) + 
    geom_boxplot() + 
    geom_jitter(height = 0) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    ylab("UNC13A cryptic inclusion \n reads per million") + 
    theme(text = element_text(size = 18)) + 
    xlab("N samples")

####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "FTLD \n non-TDP",
                              "FTLD-TDP")) %>% 
    filter(tissue_clean %in% c("Cerebellum","Frontal Cortex","Temporal Cortex")) %>% 
    ggplot(aes(x = disease_group2, y = value * 10^6)) + 
    geom_boxplot() + 
    geom_jitter(height = 0) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_grid(vars(junction_name),vars(tissue_clean)) + 
    ylab("UNC13A split \n reads per million") +
    theme(text = element_text(size = 18)) + 
    xlab("")

```

Splitting by the type of junction in ALS
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "ALS \n non-TDP",
                              "ALS-TDP")) %>% 
    ggplot(aes(x = disease_group2, y = value * 10^6, 
               fill = junction_name,
               group = junction_name)) + 
    geom_boxplot() + 
    geom_point(pch = 21,position = position_dodge(0.8)) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_wrap(vars(tissue_clean)) + 
    ylab("UNC13A cryptic inclusion \n reads per million") +
    # theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
    scale_fill_manual(values = colorblind_pal()(4)[2:4])

```

```{r}
clean_data_table %>% 
    mutate(disease_group2 = case_when(disease == "Control" ~ "Control",
                                 pathology %in% c("FTD-TDP-A","FTD-TDP-B","FTD-TDP-C") ~ "FTLD-TDP",
                                 pathology %in% c("FTLD-TAU","FTLD-FUS") ~ "FTLD-non-TDP",
                                 mutations %in% c("SOD1","FUS") ~ "ALS \n non-TDP",
                                 T ~ "ALS-TDP")) %>% 
    filter(disease_group2 %in% c("ALS \n non-TDP","ALS-TDP")) %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,disease_group2,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_1",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    dplyr::select(-variable,-value) %>% 
    unique() %>% 
    ggplot(aes(x = disease_group2, y = inclusion_reads * 10^6)) + 
    geom_boxplot() + 
    geom_jitter(height = 0) + 
    facet_wrap(~tissue_clean,nrow = 2) + 
    scale_y_continuous(trans = 'log10') + 
    ggpubr::theme_pubr() +
    ylab("UNC13A cryptic inclusion reads per million")  
```

# Relationship between UNC13A and STMN2 cryptic PSI

We noted in the cell lines that there was a linear relationship between the amount
of TARDBP KD efficiency and the amount of observed UNC13A cryptic. As a quantitative measure
of TDP-43 proteinopathy does not exist for these patients, no ranked staining, we used the amount of STMN2 
present in a tissue to rank the tissue by the amount of TDP-43 proteinopathy, as we've already shown
you can use that as a measure. Here we show the relationship in samples with 
at least 10 STMN2 annotated junctions and tissues where the PSI of UNC13A cryptic was also detectable, as measured by
STMN2 cryptic occurring in the tissue.

## Scatter plot showing the relationship in non-logspace between STMN2 and UNC13A cryptic PSI
```{r}
####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
clean_data_table %>% 
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "stmn_2_cryptic_psi_leaf",
                      y = "unc13a_cryptic_leaf_psi",
                      add = "reg.line",
                      cor.coef = T) +
    ylab("UNC13A Cryptic PSI ") + 
    xlab("STMN2 Cryptic PSI ")

####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
clean_data_table %>% 
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
    mutate(call = fct_relevel(call,"C/C", "C/G", "G/G")) %>%
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "stmn_2_cryptic_psi_leaf",
                      y = "unc13a_cryptic_leaf_psi",
                      add = "reg.line",
                      color = "call") +
    ylab("UNC13A Cryptic PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
    stat_cor(aes(color = call),show.legend = FALSE) +
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a"))
```

You can see a clear relationship, which appears to be modulated by the genotype
of rs12973192. This lead us to see if genotype was actually predictive of the 
amount of UNC13A cryptic psi. Because PSI's tend to be log linear, we chose
to model UNC13A in the log10 + 1 (log10(1) is zero) and we selected for tissues
which show TDP-43 pathology, as we assume this is an event which only begins to
show clear after TDP-43 pathology arises.

## Scatter plot showing the relationship in log + 1 between STMN2 and UNC13A cryptic PSI

```{r}

clean_data_table %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1)) %>% 
    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn_2_cryptic_psi_leaf_plusone",
                      y = "log10_unc13a_cryptic_leaf_psi_plusone",
                      color = 'call',fill = 'call',                      
                      add = "reg.line") +
    ylab("Log10 UNC13A Cryptic PSI + 1") + 
    xlab("Log10 STMN2 Cryptic PSI + 1 ") +
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) + 
    scale_fill_manual(values = c("#88CCEE","#44AA99","#105e2a"))

```

We can perform this analysis separated by tissue, but low numbers
make the results more complicated to interpret. Additionally, the tissue only
matters in so much as we know that certain tissues in certain diseases
are affected by TDP-43 pathology, whereas others aren't, tissue is not the 
most interesting or important variable, TDP-43 pathology is the question.

```{r}
####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
clean_data_table %>% 
    filter(tissue_clean%in% c("Cervical_Spinal_Cord",
                              "Motor_Cortex",
                              "Frontal_Cortex","Temporal_Cortex")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(cryptic_psi > 0 ) %>% 
    filter(stmn_2_cryptic_psi > 0 ) %>% 
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "stmn_2_cryptic_psi",
                      y = "unc13a_cryptic_leaf_psi",
                      color = 'call',                      
                      add = "reg.line") +
    ylab("UNC13A Cryptic PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
    facet_wrap(~tissue_clean)

```

## Modeling UNC13A cryptic using glms and mixed models

This brought us to the idea that we could even potentially model
UNC13A cryptic pathology using genotype and the amount of TDP-43 pathology. 
We've coded the number of G alleles here as a number and are looking into tissue
with TDP-43 pathology(measure with STMN2 cryptic PSI).

```{r}

new_regress = clean_data_table %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 )

another_model2 = lm(log10(unc13a_cryptic_leaf_psi + 1) ~ 
                   log10(stmn_2_cryptic_psi_leaf + 1) * number_g_alleles + pathology + tissue_clean,
              new_regress)

print(pretty_effect_plot(another_model2,p_value_cutoff = 1,"UNC13A Cryptic PSI"))
```

We find that at a cutoff of 0.01 for significant effects, only the interaction
between the number of G alleles and the amount of STMN2 pathology is predictive
of UNC13A cryptic inclusion. the absolute amount of STMN2 pathology does not
vary between genotypes. 
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

new_regress = new_regress %>% 
    mutate(log10_stmn_2_cryptic_psi_leaf = log10(stmn_2_cryptic_psi_leaf))

ggboxplot(new_regress, x = "call", y = "log10_stmn_2_cryptic_psi_leaf",
          color = "call", palette = "jco")+
  stat_compare_means() +
  stat_compare_means(comparisons = genotype_comparisons)

ggboxplot(new_regress, x = "call", y = "stmn_2_cryptic_psi_leaf",
          color = "call", palette = "jco")+
  stat_compare_means() +
  stat_compare_means(comparisons = genotype_comparisons)

new_regress %>% 
    ggplot() + 
    stat_ecdf(aes(color = call, x = log10_stmn_2_cryptic_psi_leaf_plusone))
```

Now I'm going to test out using mixed models.

```{r}
library(lme4)
unc_mixed = lmer(log10_unc13a_cryptic_leaf_psi_plusone ~ log10_stmn_2_cryptic_psi_leaf_plusone * number_g_alleles + (1 | pathology), data = new_regress)
summary(unc_mixed)
unc_mixed2 = lmer(log10_unc13a_cryptic_leaf_psi_plusone ~ log10_stmn_2_cryptic_psi_leaf_plusone * number_g_alleles+ (1 | pathology) + (1 | tissue_clean), data = new_regress)
summary(unc_mixed2)

```

This leads us to exploring the affects of UNC13A psi separately for tissues
with 'low' STMN2 and those with 'high' STMN2. This distinction is simply cutting
the distribution of STMN2 psi in half and assigning each tissue to the 'low' or 
'high' group as so
```{r}

```

### 2020-11-17 paper_figures.Rmd
first update on using Rprojects and github

@@
        "scales",
-       "here")
+       "here",
+       "ggthemes")
 ```
@@
     ggplot() + 
-    stat_ecdf(aes(color = call, x = log10_stmn_2_cryptic_psi_leaf_plusone))
+    stat_ecdf(aes(color = call, x = log10_stmn_2_cryptic_psi_leaf))
 ```

### 2020-11-17 paper_figures.Rmd
tweaks

@@
 
+First we look at detection rate in tissues affected by TDP-43 proteinopathy,
+For FTLD this is frontal and temporal Cortices, and
+for ALS this is lumbar, cervical, and thoracic spinal cord samples as well as 
+motor cortex. For controls we also take all 6 tissues, 
+frontal,temporal,motor cortices and the lumbar, cervical, and thoracic spinal cords.
+
+(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
+the ALS sample tissues are not balanced and not every participant has samples in every tissue)
 ```{r}
@@
     filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
-    dplyr::select(sample,participant_id,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
+    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
     melt() %>% 
@@
 
-boxplot_table %>% 
+disease_comparisons = list( c("Control","ALS-TDP"),
+                           c("Control","ALS \n non-TDP"),
+                           c("Control","FTLD-TDP"),
+                           c("Control","FTLD \n non-TDP" ))
+
+table_to_test = boxplot_table %>% 
     filter(disease_tissue == T) %>% 
@@
     mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
+    dplyr::select(detection_name,
+                  participant_id,
+                  inclusion_reads,
+                  disease_group2) %>% 
+    unique()
+
+test_pair = pairwise.wilcox.test(table_to_test$inclusion_reads, table_to_test$detection_name,
+                     p.adjust.method = "BH") %>% broom::tidy()
+
+test_pair = test_pair %>% 
+  mutate(p_value_draw = case_when(p.value < 0.0001~ "***",
+                                  p.value < 0.01 ~ "**",
+                                   p.value < 0.05 ~ "*",
+                          TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>% 
+  mutate(y.position = seq(0.25,by = 0.1,length.out = 9))
+
+table_to_test %>% 
     ggplot(aes(x = detection_name, y = inclusion_reads * 10^6)) + 
@@
     theme(text = element_text(size = 18)) + 
-    xlab("N samples")
+    xlab("N samples") + 
+    stat_pvalue_manual(test_pair %>% filter(p.value < 0.05),
+                       label = "p_value_draw") +
+   stat_compare_means()
 
@@
     filter(tissue_clean %in% c("Cerebellum","Frontal Cortex","Temporal Cortex")) %>% 
-    ggplot(aes(x = disease_group2, y = value * 10^6)) + 
-    geom_boxplot() + 
-    geom_jitter(height = 0) + 
+  dplyr::select(disease_group2,
+                tissue_clean,
+                junction_name,
+                value) %>% 
+  unique() %>% 
+    ggplot(aes(x = disease_group2, 
+               y = value * 10^6,
+               fill = junction_name)) + 
+    geom_boxplot(show.legend = F) + 
+    geom_jitter(height = 0,show.legend = F) + 
     scale_y_log10() + 
@@
     facet_grid(vars(junction_name),vars(tissue_clean)) + 
-    ylab("UNC13A split \n reads per million") +
+    ylab("UNC13A cryptic \n reads per million") +
     theme(text = element_text(size = 18)) + 
-    xlab("")
+    xlab("") +
+    scale_fill_manual(values = colorblind_pal()(4)[2:4])
 
@@
                               "ALS-TDP")) %>% 
-    ggplot(aes(x = disease_group2, y = value * 10^6, 
-               fill = junction_name,
-               group = junction_name)) + 
-    geom_boxplot() + 
-    geom_point(pch = 21,position = position_dodge(0.8)) + 
+      filter(tissue_clean %in% c("Cerebellum","Motor Cortex","Temporal Cortex") | grepl("Cord",tissue_clean)) %>% 
+  dplyr::select(disease_group2,
+                tissue_clean,
+                junction_name,
+                value) %>% 
+  unique() %>% 
+  mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
+    ggplot(aes(x = disease_group2, 
+               y = value * 10^6,
+               fill = junction_name)) + 
+    geom_boxplot(show.legend = F) + 
+    geom_jitter(height = 0,show.legend = F) + 
     scale_y_log10() + 
     ggpubr::theme_pubr() +
-    facet_wrap(vars(tissue_clean)) + 
-    ylab("UNC13A cryptic inclusion \n reads per million") +
-    # theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
-    scale_fill_manual(values = colorblind_pal()(4)[2:4])
-
+    facet_grid(vars(junction_name),vars(tissue_clean)) + 
+    ylab("UNC13A cryptic \n reads per million") +
+    theme(text = element_text(size = 18)) + 
+    xlab("") +
+    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
+  theme(axis.text.x =  element_text(size = 14))
 ```
@@
     filter(STMN2_annotated_leaf > 10) %>% 
-    filter(stmn_2_cryptic_psi_leaf > 0 )
+    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%  
+    mutate(log10_stmn_2_cryptic_psi_leaf = log10(stmn_2_cryptic_psi_leaf)) %>% 
+    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
+    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1))
 
@@
 genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))
-
-new_regress = new_regress %>% 
-    mutate(log10_stmn_2_cryptic_psi_leaf = log10(stmn_2_cryptic_psi_leaf))
 

### 2020-11-24 paper_figures.Rmd
Adding more of the data

@@
        "here",
-       "ggthemes")
+       "ggthemes",
+       "ggrepel")
 ```
@@
 ```
+# iPSC splicing
 
-# Summary statistics
+Read in the splicing results returned by MAJIQ and make a volcano plot, only highlight 
+genes of interest with a label.
+
+```{r}
+junc_colors = c("#648FFF","#fe6101")
+ipsc_splicing = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/ipsc_splicing_results.csv")
+
+ipsc_splicing %>% 
+  mutate(junction_name = glue::glue("{gene_name} - {paste_into_igv_junction}")) %>% 
+  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
+  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
+  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
+  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
+  mutate(label_junction = ifelse(log10_test_stat == 16 & deltaPSI > 0 , junction_name,"")) %>%
+  ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
+  geom_point(aes(fill = `Novel Junction`,alpha = graph_alpha), pch = 21, size = 4) + 
+  geom_text_repel(aes(label = label_junction), point.padding = 0.3,
+                  min.segment.length = unit(0, 'lines')) +
+  geom_hline(yintercept = -log10(1 - .9)) + 
+  guides(alpha = FALSE, size = FALSE) + 
+  scale_fill_manual(values = junc_colors) + 
+  theme(legend.position = 'top') + 
+  ggpubr::theme_pubr() + 
+  xlab("delta PSI") + 
+  ylab("-Log 10 Test Statistic") + 
+  theme(text = element_text(size = 18))
+```
+
+# UNC13A Cryptic PSI across TDP-43 KD experiments
+
+```{r}
+rel_rna_cryptic_amount = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv")
+
+rel_rna_cryptic_amount %>% 
+    ggbarplot(,
+              x = "source",
+              add = c("mean_se","jitter"),
+              y = "cryptic_psi",
+              fill = 'condition',
+              color = 'condition',
+              position = position_dodge(0.8)) + 
+    ggpubr::theme_pubr() +
+    scale_fill_manual(
+        values = c("#40B0A6","#E1BE6A")
+    ) + 
+    scale_color_manual(
+        values = c("#1C2617","#262114")
+    ) +
+    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
+                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
+                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
+                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
+                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
+    ylab("UNC13A Cryptic PSI") + 
+    xlab("") + 
+    guides(color = FALSE)
+
+```
+
+# Paient summary statistics
 First let's look at only including tissues with detectable stmn2 or UNC13A CE
@@
                               "C/C", "C/G", "G/G")) %>% 
-    mutate(number_g_alleles = as.numeric(call) - 1)
+    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
+    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
+    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi_leaf), n = 2))) %>%
+    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
+                                      stmn2_psi_groups == 2 ~ "High STMN2",
+                                      TRUE ~ "No STMN2"))
 
@@
   mutate(n_sample_detected = sum(detected)) %>% 
-  select(disease_group2,n_sample,n_sample_detected) %>% 
+  dplyr::select(disease_group2,n_sample,n_sample_detected) %>% 
   unique() %>% 
@@
 
-Now I'm going to test out using mixed models.
+We see that detection rate by genotype varies.
 
 ```{r}
-library(lme4)
-unc_mixed = lmer(log10_unc13a_cryptic_leaf_psi_plusone ~ log10_stmn_2_cryptic_psi_leaf_plusone * number_g_alleles + (1 | pathology), data = new_regress)
-summary(unc_mixed)
-unc_mixed2 = lmer(log10_unc13a_cryptic_leaf_psi_plusone ~ log10_stmn_2_cryptic_psi_leaf_plusone * number_g_alleles+ (1 | pathology) + (1 | tissue_clean), data = new_regress)
-summary(unc_mixed2)
+overall_fisher = clean_data_table %>%
+
+  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
+  dplyr::select(participant_id,unc13a_detected,call) %>% 
+  unique() %>% 
+  group_by(call) %>% 
+  mutate(n_sample = n_distinct(participant_id)) %>% 
+  mutate(n_sample_detected = sum(unc13a_detected)) %>% 
+  dplyr::select(call,n_sample,n_sample_detected) %>% 
+  unique()
+
+over_p = overall_fisher %>% 
+  column_to_rownames('call') %>% 
+  fisher.test() %>% 
+  broom::tidy() %>% 
+  .$p.value
+
+overall_fisher %>% 
+  mutate(detection_rate = n_sample_detected / n_sample) %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
+  geom_col(show.legend = F) + 
+  ggpubr::theme_pubr() + 
+  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
+  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
+  theme(text = element_text(size = 18)) + 
+  xlab("N individuals") + 
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))
 
@@
 
-This leads us to exploring the affects of UNC13A psi separately for tissues
-with 'low' STMN2 and those with 'high' STMN2. This distinction is simply cutting
-the distribution of STMN2 psi in half and assigning each tissue to the 'low' or 
-'high' group as so
+Although this difference is not significant, with the Fisher's exact giving
+a p-value of `r round(over_p,2)`.
+
 ```{r}
+detection_table = clean_data_table %>% 
+  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
+  group_by(call,unc13a_detected,stmn2_psi_groups) %>% 
+  add_count(name = "genotype_detected") %>% 
+  dplyr::select(call,unc13a_detected,genotype_detected,stmn2_psi_groups) %>% 
+  unique() %>% 
+  pivot_wider(names_from = "unc13a_detected",
+              values_from = "genotype_detected") %>% 
+  dplyr::rename(unc_not_detected = `FALSE`, unc_cryptic_detected = `TRUE`) %>%
+  mutate(total_tissue = (unc_not_detected + unc_cryptic_detected)) %>% 
+  mutate(detection_rate = unc_cryptic_detected / total_tissue) %>% 
+  ungroup() %>% 
+  mutate(stmn2_psi_groups = fct_relevel(stmn2_psi_groups, "No STMN2", after = 0)) %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {total_tissue} )"))
 
+count <- 0
+breaks_fun <- function(x) {
+  count <<- count + 1L
+  switch(
+    count,
+    detection_table$detection_name[1:3],
+    detection_table$detection_name[4:6],
+    detection_table$detection_name[7:9],
+  )
+}
+
+detection_table %>% 
+  ggplot(aes(x = 1, y = detection_rate,fill = call)) + 
+  geom_col(show.legend = F,position = 'dodge2') + 
+  facet_wrap(~stmn2_psi_groups) +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) + 
+  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
+  ylab("Percent of Samples \n UNC13A Cryptic Detected") + 
+  theme(text = element_text(size = 18))  
+  # xlab("N Tissues") +
+  # scale_x_continuous(labels = breaks_fun, breaks = c(0.75, 1, 1.25), limits = c(0, NA)) 
 ```

### 2020-11-25 paper_figures.Rmd
Continuing to fill out the Rmd

@@
   mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
-  mutate(label_junction = ifelse(log10_test_stat == 16 & deltaPSI > 0 , junction_name,"")) %>%
+  mutate(label_junction = case_when(gene_name %in% c("UNC13A",
+                                                     "UNC13B","PFKP","SETD5",
+                                                     "ATG4B","STMN2") & 
+                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
+                                      deltaPSI > 0 ~ junction_name,
+                                    T ~ "")) %>% 
   ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
@@
 
+We have 2 PSI's for the UNC13A cryptic, one which includes both the short and
+long form of the cryptic, and one which does not. While we're not mentioning in 
+figure 1, given that the longer form of the cryptic appears in control cerebellum, 
+one could argue that that junction should not be included in the PSI calculation
+here for UNC13A cryptic per se. In practice, this makes very little difference in the
+conclusions made on the cell lines, so I've included both for completeness.
+
+The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
+The NB cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
+There was variability on the Klim hMN set on allelic expression. 
 ```{r}
 rel_rna_cryptic_amount = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv")
+rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
+                                          UNC13A_5prime +  UNC13A_5prime_2 + 
+                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
+                                          UNC13A_5prime +  UNC13A_5prime_2 + 
+                                          UNC13A_5prime_3)]
+rel_rna_cryptic_amount %>% 
+    ggbarplot(,
+              x = "source",
+              add = c("mean_se","jitter"),
+              y = "cryptic_psi_full",
+              fill = 'condition',
+              color = 'condition',
+              position = position_dodge(0.8)) + 
+    ggpubr::theme_pubr() +
+    scale_fill_manual(
+        values = c("#40B0A6","#E1BE6A")
+    ) + 
+    scale_color_manual(
+        values = c("#1C2617","#262114")
+    ) +
+    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
+                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
+                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
+                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
+                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
+    ylab("UNC13A Cryptic PSI") + 
+    xlab("") + 
+    guides(color = FALSE)
 
@@
     guides(color = FALSE) 
-
 ```
@@
 clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
     mutate(call = fct_relevel(call,
@@
 
-print(pretty_effect_plot(another_model2,p_value_cutoff = 1,"UNC13A Cryptic PSI"))
+the_pretty_plot = pretty_effect_plot(another_model2,p_value_cutoff = 1,"UNC13A Cryptic PSI") + ggpubr::theme_pubr()
+print(the_pretty_plot)
 ```
@@
 overall_fisher = clean_data_table %>%
-
+    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
   mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
@@
   dplyr::select(call,n_sample,n_sample_detected) %>% 
-  unique()
+  unique() %>% 
+  mutate(n_non_detected = n_sample - n_sample_detected) %>% 
+  dplyr::select(-n_sample)
 
@@
 overall_fisher %>% 
+  mutate(n_sample = n_non_detected + n_sample_detected) %>% 
   mutate(detection_rate = n_sample_detected / n_sample) %>% 
@@
   ggpubr::theme_pubr() + 
-  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
-  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
+  scale_y_continuous(labels = scales::percent) + 
+  ylab("% of TDP-43 Proteionopathy Patients \n UNC13A Cryptic Detected") + 
   theme(text = element_text(size = 18)) + 
@@
 
+Only looking at disease relevant tissue and in patients
 ```{r}
 detection_table = clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+    mutate(call = fct_relevel(call,
+                              "C/C", "C/G", "G/G")) %>%
+    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi), n = 2))) %>%
+    mutate(stmn2_psi_groups = ifelse(is.na(stmn2_psi_groups),"  No STMN2",stmn2_psi_groups)) %>%
+    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
+                                        stmn2_psi_groups == 2 ~ "High STMN2",
+                                        TRUE ~ stmn2_psi_groups)) %>%
   mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
@@
 
-count <- 0
-breaks_fun <- function(x) {
-  count <<- count + 1L
-  switch(
-    count,
-    detection_table$detection_name[1:3],
-    detection_table$detection_name[4:6],
-    detection_table$detection_name[7:9],
-  )
-}
-
 detection_table %>% 
-  ggplot(aes(x = 1, y = detection_rate,fill = call)) + 
+  ggplot(aes(x = call, y = detection_rate,fill = call)) + 
   geom_col(show.legend = F,position = 'dodge2') + 
@@
   ylab("Percent of Samples \n UNC13A Cryptic Detected") + 
-  theme(text = element_text(size = 18))  
-  # xlab("N Tissues") +
-  # scale_x_continuous(labels = breaks_fun, breaks = c(0.75, 1, 1.25), limits = c(0, NA)) 
+  theme(text = element_text(size = 18))  +
+  ggpubr::theme_pubr() + 
+  geom_text(aes(label = total_tissue,y = 0),vjust = 1,size = 6) + 
+  theme(text = element_text(size = 18)) +
+  ggpubr::theme_pubr() +
+  xlab("N Tissues") 
 ```
+
+UNC13A cryptic PSI
+
+```{r}
+clean_data_table %>%   
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+    mutate(call = fct_relevel(call,
+                              "C/C", "C/G", "G/G")) %>%
+    filter(stmn2_annotated > 10) %>% 
+    filter(cryptic_psi > 0) %>% 
+    ggbarplot( x = "call", 
+               y = "cryptic_psi", 
+               color = 'call',
+              position = position_dodge(0.8),
+              add = c("mean_se","jitter")) + 
+    ylab("UNC13A Cryptic PSI") + 
+    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
+    theme(text = element_text(size = 20)) + 
+    stat_compare_means(aes(group = call), label = "p.format") + 
+    stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif")
+```
+
+```{r}
+clean_data_table %>%   
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+    mutate(call = fct_relevel(call,
+                              "C/C", "C/G", "G/G")) %>%
+    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi), n = 2))) %>%
+    mutate(stmn2_psi_groups = ifelse(is.na(stmn2_psi_groups),"  No STMN2",stmn2_psi_groups)) %>%
+    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
+                                        stmn2_psi_groups == 2 ~ "High STMN2",
+                                        TRUE ~ stmn2_psi_groups)) %>%
+    filter(cryptic_psi > 0) %>%
+    ggbarplot( x = "call", 
+               y = "cryptic_psi", 
+               color = 'call',
+               facet.by = 'stmn2_psi_groups',
+              position = position_dodge(0.8),
+              add = c("mean_se","jitter")) + 
+    ylab("UNC13A Cryptic PSI") + 
+    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
+    theme(text = element_text(size = 20)) + 
+    stat_compare_means(aes(group = call), label = "p.format") + 
+    stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif")  + 
+  xlab("")
+```

### 2020-11-26 paper_figures.Rmd
Update paper_figures.Rmd
Adding on the first VAF figures

@@
 junc_colors = c("#648FFF","#fe6101")
-ipsc_splicing = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/ipsc_splicing_results.csv")
+ipsc_splicing = fread(file.path(here(),"data","ipsc_splicing_results.csv")
 
@@
 ```{r}
-rel_rna_cryptic_amount = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv")
+
+rel_rna_cryptic_amount = fread(file.path(here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
 rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
@@
 Only looking at disease relevant tissue and in patients 
+we see that the detection rate depends on both the level of TPD-43 pathology present, 
+again I acknowledge STMN2 is a proxy, but it is what is is with the dataset, and the geneotype 
+of the patients.
+
+This difference is also not significant with a fisher's exact test, but
+the
+
 ```{r}
@@
 ```
+
+## Variant allele expression in TDP-43 KD experiments
+
+Across our KD's, we observed a clear allelic bias in the NB cells
+
+```{r}
+kd_vafs = fread(file.path(here(),"data","all_kds_unc13.snp.out.tsv"))
+
+kd_vafs %>% 
+    left_join(rel_rna_cryptic_amount) %>% 
+    filter(!is.na(condition)) %>% 
+    mutate(VAF = alt_reads / total) %>% 
+    filter(condition != "control" ) %>% 
+    filter(source %in% c("sh_dzap","shsy5y_normed","ipsc_normed_ward")) %>% 
+    ggplot(aes(x = TARDBP,y = VAF, fill = source,size = total)) + 
+    geom_point(pch = 21)  + 
+    ylim(0,1)
+
+kd_vafs %>% 
+    left_join(rel_rna_cryptic_amount) %>% 
+    filter(!is.na(condition)) %>% 
+    filter(condition != "control" ) %>%     
+    mutate(VAF = alt_reads / total) %>% 
+    filter(source %in% c("sh_dzap","shsy5y_normed","ipsc_normed_ward")) %>% 
+    dplyr::select(sample,ref_reads,alt_reads,TARDBP,source,VAF) %>% 
+    data.table::melt(id.vars = c("sample","TARDBP",'source','VAF')) %>% 
+    mutate(sample = fct_reorder(sample,-TARDBP)) %>% 
+    ggplot(aes(x = sample,y = value, fill = variable)) + 
+    geom_col(pch = 21, size = 4) +
+    coord_flip()
+```
+
+Also look at the coverage in the Liu nuclear facs sorted neurons
+
+```{r}
+liu_vafs = fread(file.path(here(),"data","liu_facs_neurons_unc13.snp.out.tsv"))
+liu_vafs %>% 
+    mutate(VAF = alt_reads / total) %>% 
+    data.table::melt(id.vars = c("sample",'VAF','total')) %>% 
+    mutate(allele = ifelse(variable == "ref_reads", "C","G")) %>% 
+    mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
+    mutate(sample_name = gsub("_unsorted", "", sample_name)) %>% 
+    mutate(condition = sub(".*_", "", sample)) %>% 
+    ggplot(aes(x = condition,y = value, fill = allele)) + 
+    geom_col(pch = 21, size = 4) +
+    facet_wrap(~sample_name,scales = "free_y")
+```

### 2020-12-11 paper_figures.Rmd
some model scripts which I will probably delete later

@@
 
-```{r setup_function, echo=FALSE}
-####helper funciton that gets the overall model p-value####
-lmp <- function (modelobject) {
-    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
-    f <- summary(modelobject)$fstatistic
-    p <- pf(f[1],f[2],f[3],lower.tail=F)
-    attributes(p) <- NULL
-    return(p)
-}
-
-pretty_effect_plot = function(model, p_value_cutoff = 0.05, title = "Model 1"){
-  # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
-    p_value = lmp(model)
-    r_squared = round(summary(model)$adj.r.squared,3)
-    pretty_plot = broom::tidy(model) %>% 
-        mutate(p_star = case_when(p.value < 0.001 ~ "***",
-                                  p.value < 0.01 ~ "**",
-                                  p.value < 0.05 ~ "*",
-                                  p.value < 0.1 ~ ".",
-                                  p.value > 0.1 ~ "NS")) %>% 
-        filter(p.value < p_value_cutoff) %>%
-        mutate(term = fct_reorder(term, -abs(estimate))) %>%
-        ggplot(aes(x = term, y = estimate,fill = estimate > 0,show.legend = F)) +
-        geom_col() +
-        coord_flip() +
-        ggtitle(title,subtitle = glue::glue("Model p-value: {p_value} \n Adj. R-squared: {r_squared}")) +
-        geom_text(aes(x=term, y= -0.01, label=p_star)) 
-}
-```
 # iPSC splicing
@@
 junc_colors = c("#648FFF","#fe6101")
-ipsc_splicing = fread(file.path(here(),"data","ipsc_splicing_results.csv")
+ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
+ipsc_de = fread(file.path(here::here(),"data","ipsc_differential_expression.csv"))
 
 ipsc_splicing %>% 
-  mutate(junction_name = glue::glue("{gene_name} - {paste_into_igv_junction}")) %>% 
+  mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
+                                                     "UNC13B","PFKP","SETD5",
+                                                     "ATG4B","STMN2") & 
+                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
+                                      deltaPSI > 0 ~ gene_name,
+                                    T ~ "")) %>% 
+  mutate(junction_name = make.unique(junction_name,sep = ".")) %>% 
   mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
@@
   mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
-  mutate(label_junction = case_when(gene_name %in% c("UNC13A",
+  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                      "UNC13B","PFKP","SETD5",
@@
 
-rel_rna_cryptic_amount = fread(file.path(here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
+rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
 rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
@@
     ) +
-    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
-                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
-                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
-                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
-                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
     ylab("UNC13A Cryptic PSI") + 
+    xlab("") + 
+    guides(color = FALSE)
+
+unca = rel_rna_cryptic_amount %>% 
+  ggbarplot(,
+            x = "source",
+            add = c("mean_se","jitter"),
+            y = "cryptic_psi",
+            fill = 'condition',
+            color = 'condition',
+            position = position_dodge(0.8)) + 
+  ggpubr::theme_pubr() +
+  scale_fill_manual(
+      values = c("#40B0A6","#E1BE6A")
+  ) + 
+  scale_color_manual(
+      values = c("#1C2617","#262114")
+  ) +
+  ylab("UNC13A Cryptic PSI") + 
+  xlab("") + 
+  guides(color = FALSE)
+
+stmn2 = rel_rna_cryptic_amount %>% 
+  ggbarplot(,
+            x = "source",
+            add = c("mean_se","jitter"),
+            y = "stmn_2_cryptic_psi",
+            fill = 'condition',
+            color = 'condition',
+            position = position_dodge(0.8)) + 
+  ggpubr::theme_pubr() +
+  scale_fill_manual(
+      values = c("#40B0A6","#E1BE6A")
+  ) + 
+  scale_color_manual(
+      values = c("#1C2617","#262114")
+  ) +
+  ylab("STMN2 Cryptic PSI") + 
   xlab("") + 
@@
               add = c("mean_se","jitter"),
-              y = "cryptic_psi",
+              y = "normalized_unc13b_ir",
+              fill = 'condition',
+              color = 'condition',
+              position = position_dodge(0.8)) + 
+    ggpubr::theme_pubr() +
+    scale_fill_manual(
+        values = c("#40B0A6","#E1BE6A")
+    ) + 
+    scale_color_manual(
+        values = c("#1C2617","#262114")
+    ) +
+    ylab("UNC13B Intron Retention Ratio") + 
+    xlab("") + 
+    guides(color = FALSE)
+
+rel_rna_cryptic_amount %>% 
+    ggbarplot(,
+              x = "source",
+              add = c("mean_se","jitter"),
+              y = "normalized_unc13a_ir",
               fill = 'condition',
@@
                               "sh_dzap" = "Appocher \n NB cells \n C/G")) +
-    ylab("UNC13A Cryptic PSI") + 
+    ylab("UNC13A Intron Retention Ratio") + 
     xlab("") + 
     guides(color = FALSE)
+
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
+    theme(legend.position = 'bottom')
+ggarrange(unca,stmn2)
 ```
@@
 ```{r}
-clean_data_table = fread(file.path(here(),"data","nygc_junction_information.csv"))
+clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
 clean_data_table = clean_data_table %>%     
@@
 
+melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
 ```
@@
   xlab("N individuals")
+```
+
+# Detection UNC13A by tissue
+
+```{r}
+boxplot_table %>% 
+  filter(disease_tissue == T) %>% 
+  mutate(detected = inclusion_reads > 0) %>% 
+  dplyr::select(participant_id,disease_group2,detected,tissue_clean) %>% 
+  unique() %>% 
+  group_by(disease_group2,tissue_clean) %>% 
+  mutate(n_sample = n_distinct(participant_id)) %>% 
+  mutate(n_sample_detected = sum(detected)) %>% 
+  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
+  unique() %>% 
+  mutate(detection_rate = n_sample_detected / n_sample) %>% 
+  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
+  mutate(detection_name = glue::glue("{disease_group2} \n {tissue_clean} \n ( {n_sample} )")) %>% 
+  ggplot() +
+  geom_col(aes(x = disease_group2, y = detection_rate)) + 
+  ggpubr::theme_pubr() + 
+  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
+  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
+  theme(text = element_text(size = 18)) + 
+  xlab("N individuals") + 
+  facet_wrap(~tissue_clean) +
+  theme(axis.text.x =  element_text(size = 14)) + 
+      theme_bw() + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+  theme(text = element_text(size = 18))  
+```
+
+# Detection UNC13A by tissue and Genotype
+
+```{r}
+boxplot_table %>% 
+  filter(disease_tissue == T) %>% 
+  mutate(detected = inclusion_reads > 0) %>% 
+  dplyr::select(participant_id,disease_group2,detected,tissue_clean) %>% 
+  left_join(clean_data_table %>% dplyr::select(participant_id,call)) %>% 
+  unique() %>% 
+  group_by(disease_group2,tissue_clean,call) %>% 
+  mutate(n_sample = n_distinct(participant_id)) %>% 
+  mutate(n_sample_detected = sum(detected)) %>% 
+  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
+  unique() %>% 
+  mutate(detection_rate = n_sample_detected / n_sample) %>% 
+  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
+  ggplot() +
+  geom_col(aes(x = disease_group2, y = detection_rate,fill = call),position = "dodge") + 
+  ggpubr::theme_pubr() + 
+  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
+  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
+  theme(text = element_text(size = 18)) + 
+  xlab("N individuals") + 
+  facet_wrap(~tissue_clean) +
+  theme(axis.text.x =  element_text(size = 14)) + 
+      theme_bw() + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+  theme(text = element_text(size = 18))  
+```
+
+```{r}
 
@@
     ylab("UNC13A cryptic \n reads per million") +
-    theme(text = element_text(size = 18)) + 
     xlab("") +
-    scale_fill_manual(values = colorblind_pal()(4)[2:4])
+    theme_bw() + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+      theme(text = element_text(size = 18))
 
@@
     ylab("UNC13A cryptic \n reads per million") +
-    theme(text = element_text(size = 18)) + 
     xlab("") +
     scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
-  theme(axis.text.x =  element_text(size = 14))
+  theme(axis.text.x =  element_text(size = 14)) + 
+      theme_bw() + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+  theme(text = element_text(size = 18))  
 ```
@@
 
+# Correlation between STMN2 Cryptic PSI and UNC13A TPM
+
+Separated by tissue only in ALS/FTLD-TDP
+```{r}
+
+clean_data_table %>% 
+    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
+  geom_point() + 
+  stat_cor(size = 8) + 
+  geom_smooth(method = lm, se = FALSE,color = "black") +
+  xlab("STMN2 Cryptic PSI") +
+  ylab("UNC13A TPM") + 
+  ggpubr::theme_pubr() + 
+  theme(text = element_text(size = 18)) +
+  facet_wrap(~tissue_clean,scales = "free_y")
+```
+
+No Tissue separation
+
+```{r}
+clean_data_table %>% 
+    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
+  geom_point() + 
+  stat_cor(size = 8) + 
+  geom_smooth(method = lm, se = FALSE,color = "black") +
+  xlab("STMN2 Cryptic PSI") +
+  ylab("UNC13A TPM") + 
+  ggpubr::theme_pubr() + 
+  theme(text = element_text(size = 18)) +
+  ylim(0,50)
+```
+
+```{r}
+#tissue separated
+clean_data_table %>% 
+      filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
+  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM)) + 
+  geom_point() + 
+  stat_cor(size = 8) + 
+  geom_smooth(method = lm, se = FALSE,color = "black") +
+  ylab("UNC13A TPM") +
+  xlab("UNC13A Cryptic PSI") + 
+  ggpubr::theme_pubr() + 
+  theme(text = element_text(size = 18)) +
+  facet_wrap(~tissue_clean,scales = 'free')
+
+clean_data_table %>% 
+      filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
+  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
+  geom_point() + 
+  stat_cor(label.y = 38,size = 8) + 
+  geom_smooth(method = lm, se = FALSE,color = "black") +
+  ylab("UNC13A TPM") +
+  xlab("STMN2 Cryptic PSI") + 
+  ylim(0,40) +
+  ggpubr::theme_pubr() + 
+  theme(text = element_text(size = 18))
+
+```
+
 # Relationship between UNC13A and STMN2 cryptic PSI
@@
 
-## Scatter plot showing the relationship in non-logspace between STMN2 and UNC13A cryptic PSI
+# Relationship between STMN2 and UNC13A cryptic PSI in patients
 ```{r}
@@
 clean_data_table %>% 
-    filter(STMN2_annotated_leaf > 10) %>% 
-    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
-    group_by(call) %>% 
-    ggpubr::ggscatter(., 
-                      x = "stmn_2_cryptic_psi_leaf",
-                      y = "unc13a_cryptic_leaf_psi",
-                      add = "reg.line",
-                      cor.coef = T) +
-    ylab("UNC13A Cryptic PSI ") + 
-    xlab("STMN2 Cryptic PSI ")
-
-####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
-clean_data_table %>% 
-    filter(STMN2_annotated_leaf > 10) %>% 
-    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
-    mutate(call = fct_relevel(call,"C/C", "C/G", "G/G")) %>%
-    group_by(call) %>% 
-    ggpubr::ggscatter(., 
-                      x = "stmn_2_cryptic_psi_leaf",
-                      y = "unc13a_cryptic_leaf_psi",
-                      add = "reg.line",
-                      color = "call") +
-    ylab("UNC13A Cryptic PSI ") + 
-    xlab("STMN2 Cryptic PSI ") +
-    stat_cor(aes(color = call),show.legend = FALSE) +
-    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a"))
-```
-
-You can see a clear relationship, which appears to be modulated by the genotype
-of rs12973192. This lead us to see if genotype was actually predictive of the 
-amount of UNC13A cryptic psi. Because PSI's tend to be log linear, we chose
-to model UNC13A in the log10 + 1 (log10(1) is zero) and we selected for tissues
-which show TDP-43 pathology, as we assume this is an event which only begins to
-show clear after TDP-43 pathology arises.
-
-## Scatter plot showing the relationship in log + 1 between STMN2 and UNC13A cryptic PSI
-
-```{r}
+  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) + 
+  geom_point() + 
+  stat_cor(size = 8) + 
+  geom_smooth(method = lm, se = FALSE,color = "black") +
+  xlab("STMN2 Cryptic PSI") +
+  ylab("UNC13A Cryptic PSI") + 
+  ggpubr::theme_pubr() + 
+  theme(text = element_text(size = 18))
 
@@
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-    mutate(call = fct_relevel(call,
-                              "C/C", "C/G", "G/G")) %>%
-    filter(STMN2_annotated_leaf > 10) %>% 
-    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
-    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1)) %>% 
-    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
-    group_by(call) %>% 
-    ggpubr::ggscatter(., 
-                      x = "log10_stmn_2_cryptic_psi_leaf_plusone",
-                      y = "log10_unc13a_cryptic_leaf_psi_plusone",
-                      color = 'call',fill = 'call',                      
-                      add = "reg.line") +
-    ylab("Log10 UNC13A Cryptic PSI + 1") + 
-    xlab("Log10 STMN2 Cryptic PSI + 1 ") +
-    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) + 
-    scale_fill_manual(values = c("#88CCEE","#44AA99","#105e2a"))
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  data.table::melt() %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  mutate(variable = ifelse(variable == "stmn_2_cryptic_psi_leaf", "STMN2", "UNC13A")) %>% 
+  ggbarplot( x = "tissue_clean", 
+               y = "value", 
+               color = 'variable',
+              position = position_dodge(0.8),
+              add = c("mean_se","jitter")) +
+  ylab("Cryptic PSI event") + 
+  facet_wrap(~variable,scales = "free") + 
+  scale_color_manual(values = colorblind_pal()(4)[1:3]) +
+  theme(legend.position = 'none') + 
+  xlab("") +
+  theme(text = element_text(size = 18))
 
@@
 ```{r}
-####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
 clean_data_table %>% 
-    filter(tissue_clean%in% c("Cervical_Spinal_Cord",
-                              "Motor_Cortex",
-                              "Frontal_Cortex","Temporal_Cortex")) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
     mutate(call = fct_relevel(call,
                               "C/C", "C/G", "G/G")) %>%
-    filter(STMN2_annotated_leaf > 10) %>% 
-    filter(cryptic_psi > 0 ) %>% 
-    filter(stmn_2_cryptic_psi > 0 ) %>% 
-    group_by(call) %>% 
+  filter(!grepl("Cord",tissue_clean)) %>% 
+    filter(call != "C/G") %>% 
     ggpubr::ggscatter(., 
@@
                       color = 'call',                      
-                      add = "reg.line") +
+                      add = "reg.line",
+                      ) +
     ylab("UNC13A Cryptic PSI ") + 
     xlab("STMN2 Cryptic PSI ") +
-    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
-    facet_wrap(~tissue_clean)
+  stat_cor(aes(                      x = stmn_2_cryptic_psi,
+                      y = unc13a_cryptic_leaf_psi,
+                      color = call ),show.legend = FALSE) + 
+  scale_color_manual(values = c("#88CCEE","#105e2a")) +
+  theme(text = element_text(size = 18))
 
@@
 
-## Modeling UNC13A cryptic using glms and mixed models
+# Detection rate by genotype
 
-This brought us to the idea that we could even potentially model
-UNC13A cryptic pathology using genotype and the amount of TDP-43 pathology. 
-We've coded the number of G alleles here as a number and are looking into tissue
-with TDP-43 pathology(measure with STMN2 cryptic PSI).
-
-```{r}
-
-new_regress = clean_data_table %>%
-    mutate(call = fct_relevel(call,
-                              "C/C", "C/G", "G/G")) %>%
-    filter(STMN2_annotated_leaf > 10) %>% 
-    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%  
-    mutate(log10_stmn_2_cryptic_psi_leaf = log10(stmn_2_cryptic_psi_leaf)) %>% 
-    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
-    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1))
-
-another_model2 = lm(log10(unc13a_cryptic_leaf_psi + 1) ~ 
-                   log10(stmn_2_cryptic_psi_leaf + 1) * number_g_alleles + pathology + tissue_clean,
-              new_regress)
-
-the_pretty_plot = pretty_effect_plot(another_model2,p_value_cutoff = 1,"UNC13A Cryptic PSI") + ggpubr::theme_pubr()
-print(the_pretty_plot)
-```
-
-We find that at a cutoff of 0.01 for significant effects, only the interaction
-between the number of G alleles and the amount of STMN2 pathology is predictive
-of UNC13A cryptic inclusion. the absolute amount of STMN2 pathology does not
-vary between genotypes. 
-```{r}
-genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))
-
-ggboxplot(new_regress, x = "call", y = "log10_stmn_2_cryptic_psi_leaf",
-          color = "call", palette = "jco")+
-  stat_compare_means() +
-  stat_compare_means(comparisons = genotype_comparisons)
-
-ggboxplot(new_regress, x = "call", y = "stmn_2_cryptic_psi_leaf",
-          color = "call", palette = "jco")+
-  stat_compare_means() +
-  stat_compare_means(comparisons = genotype_comparisons)
-
-new_regress %>% 
-    ggplot() + 
-    stat_ecdf(aes(color = call, x = log10_stmn_2_cryptic_psi_leaf))
-```
-
-We see that detection rate by genotype varies.
+## Barplot of detection by genotype
 
@@
   column_to_rownames('call') %>% 
-  fisher.test() %>% 
+  chisq.test() %>% 
   broom::tidy() %>% 
@@
 Only looking at disease relevant tissue and in patients 
-we see that the detection rate depends on both the level of TPD-43 pathology present, 
-again I acknowledge STMN2 is a proxy, but it is what is is with the dataset, and the geneotype 
-of the patients.
-
-This difference is also not significant with a fisher's exact test, but
-the
+we see that the detection rate depends on both the level of TPD-43 pathology present, as measured by STMN2 CE expression,
 
@@
 ```{r}
+genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))
+
 clean_data_table %>%   
@@
                               "C/C", "C/G", "G/G")) %>%
-    filter(stmn2_annotated > 10) %>% 
-    filter(cryptic_psi > 0) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0) %>% 
     ggbarplot( x = "call", 
-               y = "cryptic_psi", 
+               y = "unc13a_cryptic_leaf_psi", 
                color = 'call',
@@
                                         TRUE ~ stmn2_psi_groups)) %>%
-    filter(cryptic_psi > 0) %>%
+    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
     ggbarplot( x = "call", 
@@
 
-## Variant allele expression in TDP-43 KD experiments
+# Variant allele expression in TDP-43 KD experiments
 
@@
 ```{r}
-kd_vafs = fread(file.path(here(),"data","all_kds_unc13.snp.out.tsv"))
+kd_vafs = fread(file.path(here::here(),"data","all_kds_unc13.snp.out.tsv"))
 
@@
     filter(condition != "control" ) %>% 
-    filter(source %in% c("sh_dzap","shsy5y_normed","ipsc_normed_ward")) %>% 
-    ggplot(aes(x = TARDBP,y = VAF, fill = source,size = total)) + 
+    filter(alt_reads > 5) %>%  
+    ggplot(aes(x = stmn_2_cryptic_psi,y = cryptic_psi, fill = -(VAF - 0.5),size = total)) + 
     geom_point(pch = 21)  + 
-    ylim(0,1)
+    coord_fixed() +
+  ylim(0,0.8) + 
+  xlim(0,0.8) +
+  geom_abline() + 
+    scale_fill_gradient2()
 
@@
 ```{r}
-liu_vafs = fread(file.path(here(),"data","liu_facs_neurons_unc13.snp.out.tsv"))
+liu_vafs = fread(file.path(here::here(),"data","liu_facs_neurons_unc13.snp.out.tsv"))
+liu_stmn2 = fread(file.path(here::here(),"data","liu_stmn2_and_unc13aaggregated.clean.annotated.bed"))
+liu_stmn2[,sample := gsub(".SJ.out","",V4)]
+liu_stmn2 = liu_stmn2 %>% 
+  unique() %>% 
+    pivot_wider(id_cols = 'sample',
+                names_from = "V7",
+                values_from = 'V5',
+                values_fill = 0)
 liu_vafs %>% 
@@
     geom_col(pch = 21, size = 4) +
-    facet_wrap(~sample_name,scales = "free_y")
+    facet_wrap(~sample_name,scales = "free")
+
+liu_stmn2 %>% 
+   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>% 
+   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime)) %>% 
+   left_join(liu_vafs) %>% 
+    mutate(alt_vaf = (alt_reads)/total) %>% 
+    mutate(condition = ifelse(grepl("negative",sample),"negative", "positive")) %>% 
+             filter(!grepl("unsorted",sample)) %>%
+
+    ggplot(aes(y = unc13a_psi,x = stmn2_psi, fill = alt_vaf)) + 
+    geom_point(pch = 21,size = 5) + 
+    ggtitle("STMN2 PSI and the VAF of the G allele \n in the Liu Nuclei ") +
+  facet_grid(~condition) + 
+  scale_fill_viridis_c(option = "plasma") + 
+  coord_fixed() + 
+  geom_abline()
 ```
+
+## Variant allele expression in patients
+
+```{r}
+nycg_vafs = fread(file.path(here::here(),"data","NYGC_all_samples_UNC13A_snp_coverage.tsv"),fill = T)
+nycg_vafs %>% 
+    left_join(clean_data_table) %>% 
+    filter(call == "C/G") %>% 
+    mutate(fraction_risk = alt_reads / total) %>% 
+    filter(!is.na(fraction_risk)) %>% 
+    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
+    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) + 
+    xlab("STMN2 Cryptic PSI") +
+    ylab("UNC13A Cryptic PSI") +
+    geom_abline() + 
+    scale_fill_gradient2()
+```

### 2020-12-11 another_try_at_models.R

@@ -1,0 +1,80 @@
+library(tidymodels)
+library(usemodels)
+
+set.seed(42)
+clean_data_table = data.table::fread(file.path(here::here(),"data","nygc_junction_information.csv"))
+
+model_df <- clean_data_table %>% 
+    filter(UNC13A_annotated_leaf > 10) %>% 
+    filter(grepl("S-TDP|D-TDP",disease_group2)) %>% 
+    filter(disease_tissue == T) %>% 
+    mutate(stmn2_psi_by_number_g_alleles = (number_g_alleles + 1) * stmn_2_cryptic_psi_leaf) %>% 
+    dplyr::select(age,
+                  unc13a_cryptic_leaf_psi,
+                  stmn2_psi_by_number_g_alleles,
+                  stmn_2_cryptic_psi_leaf,
+                  number_g_alleles,
+                  disease_group2,
+                  tissue_clean,
+                  reported_mutations,
+                  sex) %>% 
+    mutate(disease_group2 = snakecase::to_snake_case(disease_group2)) %>% 
+    dplyr::rename(disease_group = disease_group2)
+
+
+
+model_split <- initial_split(model_df, strata = unc13a_cryptic_leaf_psi)
+model_train <- training(model_split)
+model_test <- testing(model_split)
+
+set.seed(234)
+model_folds <- bootstraps(model_train, strata = unc13a_cryptic_leaf_psi)
+
+# now a random forest -----------------------------------------------------
+
+ranger_recipe <- 
+    recipe(formula = unc13a_cryptic_leaf_psi ~ ., data = model_df) %>% 
+    step_string2factor(one_of("disease_group", "tissue_clean", "reported_mutations", 
+                              "sex")) 
+
+ranger_spec <- 
+    rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
+    set_mode("regression") %>% 
+    set_engine("ranger") 
+
+ranger_workflow <- 
+    workflow() %>% 
+    add_recipe(ranger_recipe) %>% 
+    add_model(ranger_spec) 
+
+set.seed(10356)
+doParallel::registerDoParallel()
+ranger_tune <-
+    tune_grid(ranger_workflow, resamples = model_folds,
+              grid = 10)
+show_best(ranger_tune, metric = "rmse")
[…]

### 2020-12-11 continuing_with_tidy_models.R

@@ -1,0 +1,254 @@
+library(tidymodels) # Includes the recipes package
+library(tidyposterior)
+library(rstanarm)
+set.seed(42)
+clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
+
+model_df <- clean_data_table %>% 
+    filter(unc13a_annotated > 10) %>% 
+    filter(grepl("S-TDP|D-TDP",disease_group2)) %>% 
+    filter(disease_tissue == T) %>% 
+    mutate(unc13a_cryptic_leaf_psi_log10 = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
+    mutate(number_g_alleles = number_g_alleles + 1) %>% 
+    dplyr::select(sample,age,
+    unc13a_cryptic_leaf_psi_log10,
+    stmn_2_cryptic_psi_leaf,
+    number_g_alleles,
+    disease_group2,
+    tissue_clean,
+    reported_mutations,
+    sex)
+
+model_split <- initial_split(model_df, prob = 0.80, strata = unc13a_cryptic_leaf_psi_log10)
+model_train <- training(model_split)
+model_test  <-  testing(model_split)
+
+####log transform and 
+simple_rec <- 
+    recipe(unc13a_cryptic_leaf_psi_log10 ~ .,
+           data = model_train) %>%
+    update_role(sample, new_role = "sample_id") %>% 
+    step_dummy(all_predictors()) %>% 
+    step_log(stmn_2_cryptic_psi_leaf, offset = 1)  %>%  
+    step_interact(terms =  ~ stmn_2_cryptic_psi_leaf:number_g_alleles)
+
+# models specs ----------------------------------------------------------
+
+standard_lm <- linear_reg() %>% 
+    set_engine("lm")
+
+# linear workflow  ----------------------------------------------------------
+
+lm_wf <- 
+    workflow() %>% 
+    add_model(standard_lm) %>% 
+    add_recipe(simple_rec)
+
+lm_fit <- fit(lm_wf, model_train)
+lm_fit %>% 
+    # This returns the parsnip object:
+    pull_workflow_fit() %>% 
+    # Now tidy the linear model object:
+    tidy() %>% 
+    filter(p.value < 0.1)
+
+
+# random forest workflow -----------------------------------------------------------
+
+rf_model <- 
+    rand_forest(trees = 1000) %>% 
[…]

### 2020-12-11 work_in_progress.R

@@ -1,0 +1,136 @@
+library(data.table)
+library(tidyverse)
+library(tidymodels)
+
+library(broom)
+library(iml)
+
+clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
+model_df <- clean_data_table %>% 
+    filter(UNC13A_annotated_leaf > 10) %>% 
+    filter(grepl("S-TDP|D-TDP",disease_group2)) %>% 
+    filter(disease_tissue == T) %>% 
+    mutate(stmn2_psi_by_number_g_alleles = (number_g_alleles + 1) * stmn_2_cryptic_psi_leaf) %>% 
+    dplyr::select(age,
+                  unc13a_cryptic_leaf_psi,
+                  stmn2_psi_by_number_g_alleles,
+                  stmn_2_cryptic_psi_leaf,
+                  number_g_alleles,
+                  disease_group2,
+                  tissue_clean,
+                  reported_mutations,
+                  sex) %>% 
+    mutate(disease_group2 = snakecase::to_snake_case(disease_group2)) %>% 
+    dplyr::rename(disease_group = disease_group2)
+
+ranger_recipe <- 
+    recipe(formula = unc13a_cryptic_leaf_psi ~ ., data = model_df) %>% 
+    step_string2factor(one_of("disease_group", "tissue_clean", "reported_mutations", 
+                              "sex")) 
+
+test_ex = bake(prep(ranger_recipe),new_data = model_df)
+
+test_ex = bake(prep(ranger_recipe),new_data = model_df)
+linear_model <- lm(unc13a_cryptic_leaf_psi~., data = test_ex)
+test_ex %>% 
+    cbind(.,.pred = predict(linear_model,test_ex)) %>% 
+    ggplot(aes(x = unc13a_cryptic_leaf_psi, y = .pred)) + 
+    geom_point() + 
+    stat_cor()
+#Coef plot
+linear_model %>% 
+    tidy(conf.int = TRUE) %>% 
+    dplyr::slice(-1) %>% 
+    mutate(significant_coeff = p.value < 0.05) %>% 
+    mutate(term = fct_reorder(term,estimate)) %>% 
+    ggplot(aes(x = term, y = estimate, fill = significant_coeff)) + 
[…]

### 2021-02-01 paper_figures.Rmd
update git ignore - add some metadata

@@
 ```{r}
-junc_colors = c("#648FFF","#fe6101")
+
 ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
@@
 
-ipsc_splicing %>% 
+splicing_dots_tables <- ipsc_splicing %>% 
   mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
@@
                                     T ~ "")) %>% 
-  mutate(junction_name = make.unique(junction_name,sep = ".")) %>% 
   mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
@@
                                       deltaPSI > 0 ~ junction_name,
+                                    T ~ ""))
+
+fig1a = ggplot() +
+  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions != 0),
+             aes(x = deltaPSI, y =log10_test_stat,
+                 alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 4) + 
+  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
+             aes(x = deltaPSI, y =log10_test_stat,
+                 alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 4) + 
+  geom_text_repel(data = splicing_dots_tables[junction_name != ""],
+                  aes(x = deltaPSI, y =log10_test_stat,
+                      label = label_junction,
+                      color = as.character(de_novo_junctions)), point.padding = 0.3,
+                    nudge_y = 0.2,
+                  min.segment.length = 0,
+                  box.padding  = 2,
+                  size=6,show.legend = F) +
+  geom_hline(yintercept = -log10(1 - .9)) + 
+     scale_fill_manual(name = "",
+                      breaks = c("Annotated Junction", "Novel Junction"),
+                      values = c("Annotated Junction" = "#648FFF", "Novel Junction" = "#fe6101") ) +
+       scale_color_manual(name = "",
+                      breaks = c("0", "1"),
+                      values = c("1" = "#648FFF", "0" = "#fe6101") ) +
+  guides(alpha = FALSE, size = FALSE) + 
+  theme(legend.position = 'top') + 
+  ggpubr::theme_pubr() + 
+  xlab("delta PSI") + 
+  ylab(expression(paste("-Lo", g[10], " Test Statistic"))) +
+  theme(text = element_text(size = 24)) +
+  theme(legend.text=element_text(size=22)) +
+  xlim(-1,1)
+
+de_table = ipsc_de %>% 
+  mutate(contains_cryptic = gene_name %in% splicing_dots_tables[cryptic_junction == T,unique(gene_name)]) %>% 
+  mutate(contains_cryptic = as.character(as.numeric(contains_cryptic))) %>% 
+  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
+                                                     "UNC13B","PFKP","SETD5",
+                                                     "ATG4B","STMN2","TARDBP") ~ gene_name,
                                     T ~ "")) %>% 
-  ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
-  geom_point(aes(fill = `Novel Junction`,alpha = graph_alpha), pch = 21, size = 4) + 
-  geom_text_repel(aes(label = label_junction), point.padding = 0.3,
-                  min.segment.length = unit(0, 'lines')) +
-  geom_hline(yintercept = -log10(1 - .9)) + 
+  mutate(graph_alpha = ifelse(padj < 0.1, 1, 0.2)) %>% 
+  mutate(y_data = -log10(padj))
+
+fig2a = ggplot(data = de_table) + 
+  geom_point(data = de_table %>% filter(contains_cryptic == "0"),
+         aes(x = log2FoldChange, y = -log10(padj),
+             alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 4) + 
+    geom_point(data = de_table %>% filter(contains_cryptic == "1"),
+         aes(x = log2FoldChange, y = -log10(padj),
+             alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 4) + 
+    geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
+                  aes(x = log2FoldChange, 
+                      y = -log10(padj),
+                      label = label_junction,
+                      color = as.character(contains_cryptic)),
+                   nudge_y = 5,
+                  min.segment.length = 0,
+                  box.padding  = 4,
+                  size=6,show.legend = F) +
+       scale_fill_manual(name = "",
+                      breaks = c("No Cryptic", "Contains Cryptic"),
+                      values = c("No Cryptic" = "#648FFF", "Contains Cryptic" = "#fe6101") ) +
+         scale_color_manual(name = "",
+                      breaks = c("1", "0"),
+                      values = c("1" = "#fe6101", "0" = "#648FFF") ) +
+  xlim(-7.5,7.5) + 
+  ylab(expression(paste("-Lo", g[10], " P-value"))) +
     guides(alpha = FALSE, size = FALSE) + 
-  scale_fill_manual(values = junc_colors) + 
   theme(legend.position = 'top') + 
   ggpubr::theme_pubr() + 
-  xlab("delta PSI") + 
-  ylab("-Log 10 Test Statistic") + 
-  theme(text = element_text(size = 18))
-```
-
-# UNC13A Cryptic PSI across TDP-43 KD experiments
+  xlab(expression(paste("Lo", g[2], " Fold Change"))) +
+  theme(text = element_text(size = 24)) +
+  theme(legend.text=element_text(size=22)) +
+    geom_hline(yintercept = -log10(0.1))  +
+  geom_vline(xintercept=c(0), linetype="dotted")
+
+```
+
+# UNC13A CE PSI across TDP-43 KD experiments
 
@@
                                           UNC13A_5prime_3)]
-rel_rna_cryptic_amount %>% 
+####barplot UNC13A CE PSI - full five####
+unc13a_cryptic_full = rel_rna_cryptic_amount %>% 
     ggbarplot(,
@@
     ggpubr::theme_pubr() +
-    scale_fill_manual(
+    scale_fill_manual(name = "",
         values = c("#40B0A6","#E1BE6A")
     ) + 
-    scale_color_manual(
+    scale_color_manual(name = "",
         values = c("#1C2617","#262114")
     ) +
-    ylab("UNC13A Cryptic PSI") + 
+    ylab("UNC13A CE PSI") + 
     xlab("") + 
-    guides(color = FALSE)
+    guides(color = FALSE) +
+    theme(text = element_text(size = 20,family = 'sans'), 
+          legend.text = element_text(size = 36,family = 'sans'),
+          axis.title.y = element_text(size = 28),
+          axis.text.y = element_text(size = 28))
+
+####barplot UNC13B NMD PSI####
+unc13b_psi_nmd = rel_rna_cryptic_amount %>% 
+    ggbarplot(,
+              x = "source",
+              add = c("mean_se","jitter"),
+              y = "unc13b_nmd_exon_psi",
+              fill = 'condition',
+              color = 'condition',
+              position = position_dodge(0.8)) + 
+    ggpubr::theme_pubr() +
+    scale_fill_manual(name = "",
+        values = c("#40B0A6","#E1BE6A")
+    ) + 
+    scale_color_manual(name = "",
+        values = c("#1C2617","#262114")
+    ) +
+    ylab("UNC13B \nNMD Exon PSI") + 
+    xlab("") + 
+    guides(color = FALSE) +
+    theme(text = element_text(size = 20,family = 'sans'), 
+          legend.text = element_text(size = 36,family = 'sans'),
+          axis.title.y = element_text(size = 28),
+          axis.text.y = element_text(size = 28))
+
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
+    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
+    ggpubr::theme_pubr() +
+    theme(text = element_text(size = 20)) +
+    scale_x_continuous(labels = scales::percent) +
+    ylab("UNC13A CE PSI") +
+    theme(legend.title=element_blank()) +
+    theme(legend.position = 'bottom') 
+####UNC13A RNA and UNC13A Cryptic####
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
+    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12) +
+    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
+    ggpubr::theme_pubr() +
+    theme(text = element_text(size = 20)) +
+    scale_x_continuous(labels = scales::percent) +
+    ylab("UNC13A CE PSI") +
+    expand_limits(y = 1) + 
+    theme(legend.title=element_blank()) +
+    theme(legend.position = 'bottom') +
+      theme(text = element_text(size = 18,family = 'sans'), 
+          legend.text = element_text(size = 20,family = 'sans'),
+          axis.title = element_text(size = 32),
+          axis.text = element_text(size = 32))
+####UNC13A RNA and UNC13A Cryptic####
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
+    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
+    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
+    ggpubr::theme_pubr() +
+    theme(text = element_text(size = 20)) +
+    scale_x_continuous(labels = scales::percent) +
+    ylab("UNC13A CE PSI") +
+    expand_limits(y = 1) + 
+    theme(legend.title=element_blank()) +
+    theme(legend.position = 'bottom') +
+      theme(text = element_text(size = 18,family = 'sans'), 
+          legend.text = element_text(size = 20,family = 'sans'),
+          axis.title = element_text(size = 32),
+          axis.text = element_text(size = 32))
+####UNC13A RNA and UNC13A IR####
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12) +
+    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
+    ggpubr::theme_pubr() +
+    theme(text = element_text(size = 20)) +
+    scale_x_continuous(labels = scales::percent) +
+    ylab("UNC13A Normalized IR") +
+    theme(legend.title=element_blank()) +
+    theme(legend.position = 'bottom') 
+####TARDBP RNA and UNC13A IR####
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = TARDBP, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
+    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
+    ggpubr::theme_pubr() +
+    theme(text = element_text(size = 20)) +
+    scale_x_continuous(labels = scales::percent) +
+    ylab("UNC13A Normalized IR") +
+    theme(legend.title=element_blank()) +
+    theme(legend.position = 'bottom')
 
@@
             add = c("mean_se","jitter"),
-            y = "cryptic_psi",
+            y = "cryptic_psi_full",
             fill = 'condition',
@@
   ) +
-  ylab("UNC13A Cryptic PSI") + 
+  ylab("UNC13A CE PSI") + 
   xlab("") + 
@@
 
-rel_rna_cryptic_amount %>% 
+####barplot UNC13B normalized IR####
+unc13b_ir = rel_rna_cryptic_amount %>% 
     ggbarplot(,
@@
     ) +
-    ylab("UNC13B Intron Retention Ratio") + 
+    ylab("Normalized UNC13B \nIntron Retention Ratio") + 
     xlab("") + 
-    guides(color = FALSE)
-
-rel_rna_cryptic_amount %>% 
+    guides(color = FALSE) + 
+      theme(legend.position = "none") +
+  theme(text = element_text(size = 18)) +
+      guides(color = FALSE) + 
+      theme(text = element_text(size = 20,family = 'sans'), 
+          legend.text = element_text(size = 36,family = 'sans'),
+          axis.title.y = element_text(size = 28),
+          axis.text.y = element_text(size = 28))
+####barplot UNC13A normalized IR####
+unc13a_ir = rel_rna_cryptic_amount %>% 
     ggbarplot(,
@@
     ) +
-    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
-                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
-                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
-                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
-                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
-    ylab("UNC13A Intron Retention Ratio") + 
+    ylab("Normalized UNC13A \nIntron Retention Ratio") + 
     xlab("") + 
-    guides(color = FALSE)
-
+    theme(legend.position = "none") +
+    guides(color = FALSE) + 
+      theme(text = element_text(size = 20,family = 'sans'), 
+          legend.text = element_text(size = 36,family = 'sans'),
+          axis.title.y = element_text(size = 28),
+          axis.text.y = element_text(size = 28))
+
+ggarrange(unc13a_cryptic_full,unc13b_psi_nmd,unc13a_ir,unc13b_ir,common.legend = T)
+####scatterplot stmn2 normalized TARDBP####
+
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = TARDBP, y = stmn_2_cryptic_psi, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
+    theme(legend.position = 'bottom') + 
+  facet_wrap(~source)
+####scatterplot UNC13BIR normalized TARDBP####
 rel_rna_cryptic_amount %>% 
@@
     theme(legend.position = 'bottom')
-ggarrange(unca,stmn2)
+####scatterplot UNC13AIR and UNC13A Crptic####
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13a_ir)) +
+    theme(legend.position = 'bottom')
+####scatterplot UNC13BIR and UNC13A Crptic####
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
+    theme(legend.position = 'bottom')+ 
+  facet_wrap(~source)
+####scatterplot UNC13BIR and UNC13B NMD####
+rel_rna_cryptic_amount %>% 
+    filter(condition != "control") %>% 
+    ggplot() + 
+    geom_point(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
+    theme(legend.position = 'bottom')+ 
+  facet_wrap(~source)
+
+####barplot UNC13A RNA level####
+
 ```
@@
     mutate(number_g_alleles = as.numeric(call) - 1) %>% 
-    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
-    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi_leaf), n = 2))) %>%
-    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
-                                      stmn2_psi_groups == 2 ~ "High STMN2",
-                                      TRUE ~ "No STMN2"))
+    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi))
 
@@
 boxplot_table = clean_data_table %>% 
-    mutate(disease_group2 = case_when(disease == "Control" ~ "Control",
-                                 pathology %in% c("FTD-TDP-A","FTD-TDP-B","FTD-TDP-C") ~ "FTLD-TDP",
-                                 pathology %in% c("FTD-TAU","FTD-FUS") ~ "FTLD \n non-TDP",
-                                 mutations %in% c("SOD1","FUS") ~ "ALS \n non-TDP",
-                                 T ~ "ALS-TDP")) %>% 
     mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
@@
     melt() %>% 
-    filter(grepl("_3prime|_5prime_",variable)) %>% 
+    filter(grepl("_3prime|_5prime_1",variable)) %>% 
     group_by(sample) %>% 
@@
     unique() %>% 
-    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel 3'", 
-                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel 5'", 
-                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel 5'")) %>% 
+    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
+                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
+                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
     mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
@@
 ```
-
+# Detection of UNC13A
 Looking at disease tissue only, so just taking the cords in ALS and the 
@@
   ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
-  theme(text = element_text(size = 18)) + 
+  theme(text = element_text(size = 24)) + 
   xlab("N individuals")
@@
 boxplot_table %>% 
-  filter(disease_tissue == T) %>% 
+  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>% 
   mutate(detected = inclusion_reads > 0) %>% 
-  dplyr::select(participant_id,disease_group2,detected,tissue_clean) %>% 
+  dplyr::select(sample,disease_group2,detected,tissue_clean) %>% 
   unique() %>% 
   group_by(disease_group2,tissue_clean) %>% 
-  mutate(n_sample = n_distinct(participant_id)) %>% 
+  mutate(n_sample = n_distinct(sample)) %>% 
   mutate(n_sample_detected = sum(detected)) %>% 
@@
   mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
-  mutate(detection_name = glue::glue("{disease_group2} \n {tissue_clean} \n ( {n_sample} )")) %>% 
+  ungroup() %>% 
+  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>% 
+  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
+                                    "Frontal Cortex",
+                                    "Lumbar Spinal Cord",
+                                    "Motor Cortex",
+                                    "Thoracic Spinal Cord")) %>% 
+    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
+  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>% 
   ggplot() +
-  geom_col(aes(x = disease_group2, y = detection_rate)) + 
-  ggpubr::theme_pubr() + 
-  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
-  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
+  geom_col(aes(x = detection_name, y = detection_rate,fill = tissue_clean),position = position_dodge2(width = 0.8, preserve = "single")) + 
+  ggpubr::theme_pubr() + 
+  ylab("Percent of Tissues \n UNC13A Cryptic Detected") + 
   theme(text = element_text(size = 18)) + 
-  xlab("N individuals") + 
-  facet_wrap(~tissue_clean) +
+  xlab("N tissues") + 
+  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
+    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) + 
   theme(axis.text.x =  element_text(size = 14)) + 
@@
     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
-  theme(text = element_text(size = 18))  
+  theme(text = element_text(size = 18))  +
+  scale_fill_manual(values = c("#4C97B5","#C87156","#4C97B5","#C87156","#4C97B5","#C87156")) +
+  theme(legend.position = "none") 
 ```
@@
 ```{r}
+
 boxplot_table %>% 
-  filter(disease_tissue == T) %>% 
+  left_join(clean_data_table %>% dplyr::select(sample, call)) %>% 
+  unique() %>% 
+  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>% 
   mutate(detected = inclusion_reads > 0) %>% 
-  dplyr::select(participant_id,disease_group2,detected,tissue_clean) %>% 
-  left_join(clean_data_table %>% dplyr::select(participant_id,call)) %>% 
+  dplyr::select(sample,disease_group2,detected,tissue_clean,call) %>% 
   unique() %>% 
   group_by(disease_group2,tissue_clean,call) %>% 
-  mutate(n_sample = n_distinct(participant_id)) %>% 
+  mutate(n_sample = n_distinct(sample)) %>% 
   mutate(n_sample_detected = sum(detected)) %>% 
@@
   mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
+  ungroup() %>% 
+  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>% 
+  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
+                                    "Frontal Cortex",
+                                    "Lumbar Spinal Cord",
+                                    "Motor Cortex",
+                                    "Thoracic Spinal Cord")) %>% 
+    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
+  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>% 
   ggplot() +
-  geom_col(aes(x = disease_group2, y = detection_rate,fill = call),position = "dodge") + 
-  ggpubr::theme_pubr() + 
-  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
-  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
+  geom_col(aes(x = detection_name, y = detection_rate,fill = call),position = position_dodge2(width = 0.8, preserve = "single")) + 
+  ggpubr::theme_pubr() + 
+  ylab("Percent of Tissues \n UNC13A Cryptic Detected") + 
   theme(text = element_text(size = 18)) + 
-  xlab("N individuals") + 
-  facet_wrap(~tissue_clean) +
+  xlab("N tissues") + 
+  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
+    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) + 
   theme(axis.text.x =  element_text(size = 14)) + 
@@
     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
-  theme(text = element_text(size = 18))  
+  theme(text = element_text(size = 18))  +
+  theme(legend.position = "none") 
+```
+
+# Junction type expression
+
+```{r}
+boxplot_table %>% 
+  filter(tissue_clean %in% c("Frontal Cortex", "Lumbar Spinal Cord", "Cervical Spinal Cord","Motor Cortex", "Temporal Cortex", "Cerebellum") )%>% 
+    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
+                                    "Frontal Cortex",
+                                    "Lumbar Spinal Cord",
+                                    "Temporal Cortex",
+                                    "Motor Cortex")) %>% 
+        mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
+  dplyr::select(disease_group2,
+                tissue_clean,
+                junction_name,
+                value) %>% 
+    unique() %>% 
+    ggplot(aes(x = disease_group2, 
+               y = value * 10^6,
+               fill = junction_name)) + 
+    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1)) + 
+    geom_point(pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
+    scale_y_log10() + 
+    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) + 
+    ylab("UNC13A cryptic \n reads per million") +
+    theme(text = element_text(size = 12)) + 
+    xlab("") +
+    scale_fill_manual(values = colorblind_pal()(4)[2:4]) 
 ```
@@
                   inclusion_reads,
-                  disease_group2) %>% 
+                  disease_group2,
+                  tissue_clean,
+                  disease_tissue) %>% 
     unique()
@@
                           TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>% 
-  mutate(y.position = seq(0.25,by = 0.1,length.out = 9))
+  mutate(y.position = seq(0.25,by = 0.1,length.out = 7))
 
@@
     ylab("UNC13A cryptic inclusion \n reads per million") + 
-    theme(text = element_text(size = 18)) + 
+    theme(text = element_text(size = 24)) + 
     xlab("N samples") + 
     stat_pvalue_manual(test_pair %>% filter(p.value < 0.05),
-                       label = "p_value_draw") +
-   stat_compare_means()
-
-####Inclusion reads by if TDP-potential####
+                       label = "p_value_draw",size = 8) +
+   stat_compare_means(size = 8)
+
+```
+
+# UNC13A inclusion separated by tissue and not type
+
+```{r}
+
+boxplot_table %>% 
+  filter(junction_name %in% c("  Novel Donor"," Short Novel Acceptor" )) %>% 
+  group_by(sample) %>% 
+  summarise(new_inc = sum(value),
+            disease_group2 = disease_group2,
+            tissue_clean = tissue_clean,
+            sample = sample) %>% 
+  ungroup() %>% 
+  unique() %>% 
+    filter(!(tissue_clean %in% c("Hippocampus","Occipital Cortex","Cerebellum"))) %>% 
+    ggplot(aes(x = disease_group2, y = new_inc * 10^6)) + 
+    geom_boxplot() + 
+    geom_jitter(height = 0) + 
+    scale_y_log10() + 
+    facet_wrap(~tissue_clean) +
+    ggpubr::theme_pubr() +
+    ylab("UNC13A cryptic inclusion \n reads per million") + 
+    theme(text = element_text(size = 24)) + 
+    xlab("N samples")
+
+```
+
+# UNC13A Inclusion Reads by type and tissue
+
+```{r}
+####ALS reads per tissue####
 boxplot_table %>% 
   filter(disease_group2 %in% c("Control",
-                              "FTLD \n non-TDP",
-                              "FTLD-TDP")) %>% 
+                              "ALS \n non-TDP",
+                              "ALS-TDP")) %>% 
+      filter(tissue_clean %in% c("Motor Cortex","Cervical Spinal Cord","Lumbar Spinal Cord")) %>% 
+  dplyr::select(disease_group2,
+                tissue_clean,
+                junction_name,
+                value) %>% 
+  unique() %>% 
+  mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
+    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex")) %>% 
+      mutate(y_val = value * 10^6) %>% 
+    ggbarplot( x = "disease_group2", 
+               y = "y_val", 
+               color = 'junction_name',
+               fill = 'junction_name',
+               position = position_dodge(0.8),
+               add = c("mean_se","jitter"),
+               facet.by = c("junction_name","tissue_clean"))+ 
+    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
+       scale_color_manual(
+        values = c("#1C2617","#262114","#262115")
+    ) +
+    theme(axis.text.x =  element_text(size = 24)) + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+    theme(text = element_text(size = 24))  +
+    ylab("UNC13A cryptic \n reads per million")  +
+    xlab("") + 
+    theme(legend.position = 'none') +
+    theme(legend.title = element_blank()) +
+  ylim(0,0.2) + 
+  theme(
+  strip.background.y = element_blank(),
+  strip.text.y  = element_blank()
+) +
+  guides(color = FALSE)
+
+####FTLD reads per tissue####
+boxplot_table %>% 
     filter(tissue_clean %in% c("Cerebellum","Frontal Cortex","Temporal Cortex")) %>% 
@@
       unique() %>% 
-    ggplot(aes(x = disease_group2, 
-               y = value * 10^6,
-               fill = junction_name)) + 
-    geom_boxplot(show.legend = F) + 
-    geom_jitter(height = 0,show.legend = F) + 
-    scale_y_log10() + 
-    ggpubr::theme_pubr() +
-    facet_grid(vars(junction_name),vars(tissue_clean)) + 
+      mutate(y_val = value * 10^6) %>% 
+      mutate(tissue_clean = fct_relevel(tissue_clean,"Cerebellum",after = Inf)) %>% 
+      mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
+    ggbarplot( x = "disease_group2", 
+               y = "y_val", 
+               color = 'junction_name',
+               fill = 'junction_name',
+               position = position_dodge(0.8),
+               add = c("mean_se","jitter"),
+               facet.by = c("junction_name","tissue_clean")) + 
+    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
+       scale_color_manual(
+        values = c("#1C2617","#262114","#262115")
+    ) +
+    theme(axis.text.x =  element_text(size = 24)) + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+    theme(text = element_text(size = 24))  +
     ylab("UNC13A cryptic \n reads per million")  +
     xlab("") + 
-    theme_bw() + 
-    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
-      theme(text = element_text(size = 18))
-
-```
-
-Splitting by the type of junction in ALS
-```{r}
-####Inclusion reads by if TDP-potential####
-boxplot_table %>% 
-  filter(disease_group2 %in% c("Control",
-                              "ALS \n non-TDP",
-                              "ALS-TDP")) %>% 
-      filter(tissue_clean %in% c("Cerebellum","Motor Cortex","Temporal Cortex") | grepl("Cord",tissue_clean)) %>% 
-  dplyr::select(disease_group2,
-                tissue_clean,
-                junction_name,
-                value) %>% 
-  unique() %>% 
-  mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
-    ggplot(aes(x = disease_group2, 
-               y = value * 10^6,
-               fill = junction_name)) + 
-    geom_boxplot(show.legend = F) + 
-    geom_jitter(height = 0,show.legend = F) + 
-    scale_y_log10() + 
-    ggpubr::theme_pubr() +
-    facet_grid(vars(junction_name),vars(tissue_clean)) + 
-    ylab("UNC13A cryptic \n reads per million") +
+    theme(legend.position = 'top') +
+    theme(legend.title = element_blank()) +
+    ylim(0,0.2) +
+    theme(
+  strip.background.y = element_blank(),
+  strip.text.y  = element_blank()
+) +
+    scale_x_discrete(guide = guide_axis(n.dodge=2)) +
+  guides(color = FALSE)
+
+```
+
+# UNC13A TPM across disease states
+
+Separated by tissue only in ALS/FTLD-TDP
+```{r}
+
+clean_data_table %>% 
+  filter(UNC13A_annotated_leaf > 10) %>% 
+  mutate(unc13_d = unc13a_cryptic_leaf_psi_full > 0) %>% 
+    filter(disease_tissue == T) %>% 
+    ggbarplot(,
+              x = "disease_group2",
+              add = c("mean_se","jitter"),
+              y = "UNC13A_TPM",
+              fill = 'unc13_d',
+              color = 'unc13_d',
+              facet.by = 'tissue_clean',
+              position = position_dodge(0.8)) + 
+    ggpubr::theme_pubr() +
+    scale_fill_manual(name = "",
+        values = c("#40B0A6","#E1BE6A")
+    ) + 
+    scale_color_manual(name = "",
+        values = c("#1C2617","#262114")
+    ) +
+    ylab("UNC13A TPM") + 
     xlab("") + 
-    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
-  theme(axis.text.x =  element_text(size = 14)) + 
-      theme_bw() + 
-    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
-  theme(text = element_text(size = 18))  
-```
-
-```{r}
-clean_data_table %>% 
-    mutate(disease_group2 = case_when(disease == "Control" ~ "Control",
-                                 pathology %in% c("FTD-TDP-A","FTD-TDP-B","FTD-TDP-C") ~ "FTLD-TDP",
-                                 pathology %in% c("FTLD-TAU","FTLD-FUS") ~ "FTLD-non-TDP",
-                                 mutations %in% c("SOD1","FUS") ~ "ALS \n non-TDP",
-                                 T ~ "ALS-TDP")) %>% 
-    filter(disease_group2 %in% c("ALS \n non-TDP","ALS-TDP")) %>% 
-    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
-    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
-    dplyr::select(sample,disease_group2,tissue_clean,contains("_library_norm")) %>% 
-    melt() %>% 
-    filter(grepl("_3prime|_5prime_1",variable)) %>% 
-    group_by(sample) %>% 
-    mutate(inclusion_reads = sum(value)) %>% 
-    ungroup() %>% 
-    dplyr::select(-variable,-value) %>% 
-    unique() %>% 
-    ggplot(aes(x = disease_group2, y = inclusion_reads * 10^6)) + 
-    geom_boxplot() + 
-    geom_jitter(height = 0) + 
-    facet_wrap(~tissue_clean,nrow = 2) + 
-    scale_y_continuous(trans = 'log10') + 
-    ggpubr::theme_pubr() +
-    ylab("UNC13A cryptic inclusion reads per million")  
+    guides(color = FALSE) +
+    theme(text = element_text(size = 20)) +
+  facet_wrap(~tissue_clean, scales = 'free') +
+  ggtitle("UNC13A Cryptic Detected") +
+  stat_compare_means()
 ```
@@
 
+# Correlation between UNC13A CE PSI and UNC13A TPM
+
 ```{r}
@@
 clean_data_table %>% 
-      filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
-  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM)) + 
+  filter(unc13a_cryptic_leaf_psi_full > 0 ) %>% 
+    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  ggplot(aes(x = unc13a_cryptic_leaf_psi_full, y = UNC13A_TPM, color = disease_group2)) + 
   geom_point() + 
@@
   ylab("UNC13A TPM") +
-  xlab("UNC13A Cryptic PSI") + 
+  xlab("UNC13A CE PSI") + 
   ggpubr::theme_pubr() + 
@@
 clean_data_table %>% 
-      filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
-  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(unc13a_cryptic_leaf_psi_full > 0) %>% 
+  ggplot(aes(x = unc13a_cryptic_leaf_psi_full, y = UNC13A_TPM)) + 
   geom_point() + 
@@
   ylab("UNC13A TPM") +
-  xlab("STMN2 Cryptic PSI") + 
-  ylim(0,40) +
-  ggpubr::theme_pubr() + 
-  theme(text = element_text(size = 18))
-
-```
-
-# Relationship between UNC13A and STMN2 cryptic PSI
-
-We noted in the cell lines that there was a linear relationship between the amount
-of TARDBP KD efficiency and the amount of observed UNC13A cryptic. As a quantitative measure
-of TDP-43 proteinopathy does not exist for these patients, no ranked staining, we used the amount of STMN2 
-present in a tissue to rank the tissue by the amount of TDP-43 proteinopathy, as we've already shown
-you can use that as a measure. Here we show the relationship in samples with 
-at least 10 STMN2 annotated junctions and tissues where the PSI of UNC13A cryptic was also detectable, as measured by
-STMN2 cryptic occurring in the tissue.
-
-# Relationship between STMN2 and UNC13A cryptic PSI in patients
+  xlab("UNC13A CE PSI") + 
+  ggpubr::theme_pubr() + 
+  theme(text = element_text(size = 22)) +
+  ylim(0,40)
+
+```
+
+# Relationship between STMN2 and UNC13A CE PSI in patients
 ```{r}
@@
 clean_data_table %>% 
+      filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0) %>% 
   ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) + 
   geom_point() + 
-  stat_cor(size = 8) + 
+  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") + 
   geom_smooth(method = lm, se = FALSE,color = "black") +
   xlab("STMN2 Cryptic PSI") +
-  ylab("UNC13A Cryptic PSI") + 
+  ylab("UNC13A CE PSI") + 
   ggpubr::theme_pubr() + 
   theme(text = element_text(size = 18)) 
-
+```
+
+# STMN2 and UNC13A CE PSI in tissue
+
+```{r}
 clean_data_table %>% 
@@
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi_full) %>% 
+  mutate(unc13a_normalized_psi_to_stmn2 = log2((unc13a_cryptic_leaf_psi_full +1)/ (stmn_2_cryptic_psi_leaf + 1))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(tissue_clean) %>% 
+  mutate(psi_mean = median(unc13a_normalized_psi_to_stmn2)) %>% 
+  ungroup() %>% 
+  mutate(tissue_clean = fct_reorder(tissue_clean,-psi_mean)) %>% 
+  ggbarplot( x = "tissue_clean", 
+               y = "unc13a_normalized_psi_to_stmn2", 
+              position = position_dodge(0.8),
+              add = c("mean_se","jitter"),
+             facet.by = 'disease_group2') +
+  ylab("Log2Fold Ratio \n UNC13A Cryptic to STMN2 Cryptic") + 
+  xlab("") +
+  theme(text = element_text(size = 24)) +
+  theme(legend.position = 'bottom') +
+  facet_wrap(~disease_group2, scales = 'free_x')
+####boxplot of full fix cryptic psi in als-tdp and ftld-tdp by genotype and tissue####
+clean_data_table %>% 
+    filter(disease_tissue == T) %>% 
+  filter(unc13a_cryptic_leaf_psi_full > 0 ) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
   dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  ggboxplot( x = "tissue_clean", 
+               y = "unc13a_cryptic_leaf_psi", 
+             fill = "call",
+             facet.by = 'disease_group2') +
+  xlab("") +
+  geom_jitter(height = 0) +
+  theme(text = element_text(size = 24)) +
+  theme(legend.position = 'bottom') +
+  facet_wrap(~disease_group2, scales = 'free')
+
+clean_data_table %>% 
+    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,STMN2_TPM,UNC13A_TPM) %>% 
+  mutate(unc13a_normalized_psi_to_stmn2 = log2((UNC13A_TPM +1)/ (STMN2_TPM + 1))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(tissue_clean) %>% 
+  mutate(psi_mean = median(unc13a_normalized_psi_to_stmn2)) %>% 
+  ungroup() %>% 
+  mutate(tissue_clean = fct_reorder(tissue_clean,-psi_mean)) %>% 
+  ggbarplot( x = "tissue_clean", 
+               y = "unc13a_normalized_psi_to_stmn2", 
+              position = position_dodge(0.8),
+              add = c("mean_se","jitter"),
+             facet.by = 'disease_group2') +
+  ylab("Log2Fold Ratio \n UNC13A Cryptic to STMN2 Cryptic") + 
+  xlab("") +
+  theme(text = element_text(size = 24)) +
+  theme(legend.position = 'bottom') +
+  facet_wrap(~disease_group2, scales = 'free_x')
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(participant_id,  sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi_full) %>%  add_count(participant_id) %>% 
+  filter(n == 5) %>% 
+  dplyr::select(-n) %>% 
   data.table::melt() %>% 
   mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  mutate(variable = ifelse(variable == "stmn_2_cryptic_psi_leaf", "STMN2", "UNC13A")) %>% 
-  ggbarplot( x = "tissue_clean", 
-               y = "value", 
-               color = 'variable',
-              position = position_dodge(0.8),
-              add = c("mean_se","jitter")) +
-  ylab("Cryptic PSI event") + 
-  facet_wrap(~variable,scales = "free") + 
-  scale_color_manual(values = colorblind_pal()(4)[1:3]) +
+  ggplot(aes(x = tissue_clean, y = value, color = participant_id)) + 
+  geom_point() +
+  geom_line(aes(group = participant_id)) +
+  theme(text = element_text(size = 24)) +
   theme(legend.position = 'none')  +
-  xlab("") +
-  theme(text = element_text(size = 18))
-
-```
-
-We can perform this analysis separated by tissue, but low numbers
-make the results more complicated to interpret. Additionally, the tissue only
-matters in so much as we know that certain tissues in certain diseases
-are affected by TDP-43 pathology, whereas others aren't, tissue is not the 
-most interesting or important variable, TDP-43 pathology is the question.
-
-```{r}
-clean_data_table %>% 
+  facet_grid(~variable)
+  scale_color_manual(values = c("#D55E00","#0072B2"))
+
+```
+
+# Correlation UNC13A PSI and STMN2 PSI with genotype effect
+
+```{r}
+tissue_to_show_correlation = clean_data_table %>% 
   filter(disease_tissue == T) %>% 
@@
   filter(!grepl("Cord",tissue_clean)) %>% 
-    filter(call != "C/G") %>% 
+    mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
+    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
+  filter(is.finite(log10_unc13a) & is.finite(log10_stmn2))
+
+tissue_to_show_correlation %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+    mutate(call = fct_relevel(call,
+                              "C/C", "C/G", "G/G")) %>%
+  filter(!grepl("Cord",tissue_clean)) %>% 
+  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
+    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
     ggpubr::ggscatter(., 
-                      x = "stmn_2_cryptic_psi",
-                      y = "unc13a_cryptic_leaf_psi",
+                      x = "log10_stmn2",
+                      y = "log10_unc13a",
                       color = 'call',                      
-                      add = "reg.line",
-                      ) +
-    ylab("UNC13A Cryptic PSI ") + 
+                      add = "reg.line", size = 3
+                      ) +
+    ylab("UNC13A CE PSI ") + 
     xlab("STMN2 Cryptic PSI ") +
-  stat_cor(aes(                      x = stmn_2_cryptic_psi,
+  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                       y = unc13a_cryptic_leaf_psi,
-                      color = call ),show.legend = FALSE) + 
-  scale_color_manual(values = c("#88CCEE","#105e2a")) +
-  theme(text = element_text(size = 18))
+                      color = call),
+           method = 'spearman',
+           cor.coef.name = 'rho',
+           show.legend = FALSE,
+           size = 14) + 
+    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  theme(legend.title = element_blank()) +
+  theme(text = element_text(size = 24),legend.text = element_text(size = 32))
 
@@
   dplyr::select(-n_sample)
+
+over_p = overall_fisher %>% 
+  column_to_rownames('call') %>% 
+  chisq.test() %>% 
+  broom::tidy() %>% 
+  .$p.value
+
+overall_fisher %>% 
+  mutate(n_sample = n_non_detected + n_sample_detected) %>% 
+  mutate(detection_rate = n_sample_detected / n_sample) %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
+  geom_col(show.legend = F) + 
+  ggpubr::theme_pubr() + 
+  scale_y_continuous(labels = scales::percent) + 
+  ylab("% of TDP-43 Proteionopathy Patients \n UNC13A Cryptic Detected") + 
+  theme(text = element_text(size = 18)) + 
+  xlab("N individuals") + 
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))
+
+```
+
+## Barplot of detection by genotype and tissue
+
+```{r}
+clean_data_table %>%
+    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
+  dplyr::select(sample,unc13a_detected,call,tissue_clean) %>% 
+  unique() %>% 
+  group_by(call,tissue_clean) %>% 
+  mutate(n_sample = n_distinct(sample)) %>% 
+  mutate(n_sample_detected = sum(unc13a_detected)) %>% 
+  dplyr::select(call,n_sample,n_sample_detected) %>% 
+  unique() %>% 
+  mutate(n_non_detected = n_sample - n_sample_detected) %>% 
+  mutate(detection_rate = n_sample_detected / n_sample) %>% 
+  ggplot(aes(x = tissue_clean, y = detection_rate, fill = call)) +
+  geom_col(show.legend = F,position = position_dodge2()) + 
+  geom_text(aes(label = n_sample,y = 0),vjust = 1,size = 11,position = position_dodge(width = 1)) + 
+  ggpubr::theme_pubr() + 
+  scale_y_continuous(labels = scales::percent) + 
+  ylab("% of Tissue samples \n UNC13A Cryptic Detected") + 
+  theme(text = element_text(size = 18)) + 
+  xlab("N individuals") + 
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))
 
@@
 detection_table %>% 
+  mutate(error = sqrt((detection_rate * (1-detection_rate))/total_tissue)) %>% 
   ggplot(aes(x = call, y = detection_rate,fill = call)) + 
   geom_col(show.legend = F,position = 'dodge2') + 
+  geom_errorbar(aes(ymin = detection_rate - error, ymax = detection_rate + error),
+                  position = position_dodge(0.9),
+                width=0.2) +
   facet_wrap(~stmn2_psi_groups) +
@@
   scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
-  ylab("Percent of Samples \n UNC13A Cryptic Detected") + 
-  theme(text = element_text(size = 18))  +
-  ggpubr::theme_pubr() + 
-  geom_text(aes(label = total_tissue,y = 0),vjust = 1,size = 6) + 
-  theme(text = element_text(size = 18)) +
-  ggpubr::theme_pubr() +
+  ylab("Percent of Samples Detected") + 
+  ggpubr::theme_pubr() + 
+  geom_text(aes(label = total_tissue,y = 0),vjust = 1,size = 11) + 
+  ggpubr::theme_pubr() +
+    theme(text = element_text(size = 32)) +
   xlab("N Tissues") 
@@
 
-UNC13A cryptic PSI
+UNC13A CE PSI
 
@@
               add = c("mean_se","jitter")) + 
-    ylab("UNC13A Cryptic PSI") + 
+    ylab("UNC13A CE PSI") + 
     scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
@@
 
+# Liu Facs Neurons
 Also look at the coverage in the Liu nuclear facs sorted neurons
@@
                 values_fill = 0)
+
+liu_stmn2 <- liu_stmn2 %>% 
+   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>% 
+   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2)) 
 liu_vafs %>% 
@@
     facet_wrap(~sample_name,scales = "free")
+
+liu_stmn2 %>% 
+  dplyr::select(stmn2_psi,sample,unc13a_psi) %>% 
+  data.table::melt() %>% 
+   mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
+    filter(!grepl("unsorted",sample_name)) %>% 
+    mutate(condition = sub(".*_", "", sample)) %>% 
+    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>% 
+    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>% 
+    ggplot(aes(x = condition,y = value, fill = variable)) + 
+    geom_col(size = 4,position = 'dodge') +
+    facet_wrap(~sample_name,scales = "free_x") + 
+    scale_fill_manual(values = c("#004D40","#882255")) +
+    theme_bw() + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+    theme(text = element_text(size = 24))  +
+      xlab("") +
+    ylab("PSI")  +
+  theme(legend.position =  'top') +
+  theme(legend.title=element_blank())
 
@@
     xlab("STMN2 Cryptic PSI") +
-    ylab("UNC13A Cryptic PSI") +
+    ylab("UNC13A CE PSI") +
     geom_abline() + 
     scale_fill_gradient2()
-```
+
+nycg_vafs %>% 
+    left_join(clean_data_table) %>% 
+    filter(call == "C/G") %>% 
+    mutate(fraction_risk = alt_reads / total) %>% 
+    filter(!is.na(fraction_risk)) %>% 
+    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+    mutate(stmn2_unc_fract = ((unc13a_cryptic_leaf_psi * 100) / (stmn_2_cryptic_psi_leaf * 100))) %>% 
+    ggplot(aes(x = stmn2_unc_fract, y = fraction_risk)) +
+    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) + 
+    xlab("STMN2 Cryptic PSI") +
+    ylab("UNC13A CE PSI") +
+    scale_fill_gradient2()
+```
+
+# UNC13A TPM by tissue and disease
+
+```{r}
+ftld_comp <- list( c("Control", "FTLD \n non-TDP"), 
+                        c("Control", "FTLD-TDP"), c("FTLD-TDP", "FTLD \n non-TDP") )
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(tissue_clean == "Frontal_Cortex") %>% 
+  ggplot(aes(x = disease_group2, y = UNC13A_TPM)) + 
+  geom_boxplot() + 
+  facet_wrap(~tissue_clean,scales = "free") + 
+  stat_compare_means(comparisons = ftld_comp)
+```

### 2021-02-01 another_try_at_models.R

@@ -1,80 +1,0 @@
-library(tidymodels)
-library(usemodels)
-
-set.seed(42)
-clean_data_table = data.table::fread(file.path(here::here(),"data","nygc_junction_information.csv"))
-
-model_df <- clean_data_table %>% 
-    filter(UNC13A_annotated_leaf > 10) %>% 
-    filter(grepl("S-TDP|D-TDP",disease_group2)) %>% 
-    filter(disease_tissue == T) %>% 
-    mutate(stmn2_psi_by_number_g_alleles = (number_g_alleles + 1) * stmn_2_cryptic_psi_leaf) %>% 
-    dplyr::select(age,
-                  unc13a_cryptic_leaf_psi,
-                  stmn2_psi_by_number_g_alleles,
-                  stmn_2_cryptic_psi_leaf,
-                  number_g_alleles,
-                  disease_group2,
-                  tissue_clean,
-                  reported_mutations,
-                  sex) %>% 
-    mutate(disease_group2 = snakecase::to_snake_case(disease_group2)) %>% 
-    dplyr::rename(disease_group = disease_group2)
-
-
-
-model_split <- initial_split(model_df, strata = unc13a_cryptic_leaf_psi)
-model_train <- training(model_split)
-model_test <- testing(model_split)
-
-set.seed(234)
-model_folds <- bootstraps(model_train, strata = unc13a_cryptic_leaf_psi)
-
-# now a random forest -----------------------------------------------------
-
-ranger_recipe <- 
-    recipe(formula = unc13a_cryptic_leaf_psi ~ ., data = model_df) %>% 
-    step_string2factor(one_of("disease_group", "tissue_clean", "reported_mutations", 
-                              "sex")) 
-
-ranger_spec <- 
-    rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
-    set_mode("regression") %>% 
-    set_engine("ranger") 
-
-ranger_workflow <- 
-    workflow() %>% 
-    add_recipe(ranger_recipe) %>% 
-    add_model(ranger_spec) 
-
-set.seed(10356)
-doParallel::registerDoParallel()
-ranger_tune <-
-    tune_grid(ranger_workflow, resamples = model_folds,
-              grid = 10)
-show_best(ranger_tune, metric = "rmse")
[…]

### 2021-02-01 continuing_with_tidy_models.R

@@ -1,254 +1,0 @@
-library(tidymodels) # Includes the recipes package
-library(tidyposterior)
-library(rstanarm)
-set.seed(42)
-clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
-
-model_df <- clean_data_table %>% 
-    filter(unc13a_annotated > 10) %>% 
-    filter(grepl("S-TDP|D-TDP",disease_group2)) %>% 
-    filter(disease_tissue == T) %>% 
-    mutate(unc13a_cryptic_leaf_psi_log10 = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
-    mutate(number_g_alleles = number_g_alleles + 1) %>% 
-    dplyr::select(sample,age,
-    unc13a_cryptic_leaf_psi_log10,
-    stmn_2_cryptic_psi_leaf,
-    number_g_alleles,
-    disease_group2,
-    tissue_clean,
-    reported_mutations,
-    sex)
-
-model_split <- initial_split(model_df, prob = 0.80, strata = unc13a_cryptic_leaf_psi_log10)
-model_train <- training(model_split)
-model_test  <-  testing(model_split)
-
-####log transform and 
-simple_rec <- 
-    recipe(unc13a_cryptic_leaf_psi_log10 ~ .,
-           data = model_train) %>%
-    update_role(sample, new_role = "sample_id") %>% 
-    step_dummy(all_predictors()) %>% 
-    step_log(stmn_2_cryptic_psi_leaf, offset = 1)  %>%  
-    step_interact(terms =  ~ stmn_2_cryptic_psi_leaf:number_g_alleles)
-
-# models specs ----------------------------------------------------------
-
-standard_lm <- linear_reg() %>% 
-    set_engine("lm")
-
-# linear workflow  ----------------------------------------------------------
-
-lm_wf <- 
-    workflow() %>% 
-    add_model(standard_lm) %>% 
-    add_recipe(simple_rec)
-
-lm_fit <- fit(lm_wf, model_train)
-lm_fit %>% 
-    # This returns the parsnip object:
-    pull_workflow_fit() %>% 
-    # Now tidy the linear model object:
-    tidy() %>% 
-    filter(p.value < 0.1)
-
-
-# random forest workflow -----------------------------------------------------------
-
-rf_model <- 
-    rand_forest(trees = 1000) %>% 
[…]

### 2021-02-01 work_in_progress.R
remove the modeling stuff

@@ -1,136 +1,0 @@
-library(data.table)
-library(tidyverse)
-library(tidymodels)
-
-library(broom)
-library(iml)
-
-clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
-model_df <- clean_data_table %>% 
-    filter(UNC13A_annotated_leaf > 10) %>% 
-    filter(grepl("S-TDP|D-TDP",disease_group2)) %>% 
-    filter(disease_tissue == T) %>% 
-    mutate(stmn2_psi_by_number_g_alleles = (number_g_alleles + 1) * stmn_2_cryptic_psi_leaf) %>% 
-    dplyr::select(age,
-                  unc13a_cryptic_leaf_psi,
-                  stmn2_psi_by_number_g_alleles,
-                  stmn_2_cryptic_psi_leaf,
-                  number_g_alleles,
-                  disease_group2,
-                  tissue_clean,
-                  reported_mutations,
-                  sex) %>% 
-    mutate(disease_group2 = snakecase::to_snake_case(disease_group2)) %>% 
-    dplyr::rename(disease_group = disease_group2)
-
-ranger_recipe <- 
-    recipe(formula = unc13a_cryptic_leaf_psi ~ ., data = model_df) %>% 
-    step_string2factor(one_of("disease_group", "tissue_clean", "reported_mutations", 
-                              "sex")) 
-
-test_ex = bake(prep(ranger_recipe),new_data = model_df)
-
-test_ex = bake(prep(ranger_recipe),new_data = model_df)
-linear_model <- lm(unc13a_cryptic_leaf_psi~., data = test_ex)
-test_ex %>% 
-    cbind(.,.pred = predict(linear_model,test_ex)) %>% 
-    ggplot(aes(x = unc13a_cryptic_leaf_psi, y = .pred)) + 
-    geom_point() + 
-    stat_cor()
-#Coef plot
-linear_model %>% 
-    tidy(conf.int = TRUE) %>% 
-    dplyr::slice(-1) %>% 
-    mutate(significant_coeff = p.value < 0.05) %>% 
-    mutate(term = fct_reorder(term,estimate)) %>% 
-    ggplot(aes(x = term, y = estimate, fill = significant_coeff)) + 
[…]

### 2021-02-11 paper_figures.Rmd
peptide plot and data

@@
 
-```{r}
+```{r volcano_plots}
 
@@
              aes(x = deltaPSI, y =log10_test_stat,
-                 alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 4) + 
+                 alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 10) + 
   geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
              aes(x = deltaPSI, y =log10_test_stat,
-                 alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 4) + 
+                 alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 10) + 
   geom_text_repel(data = splicing_dots_tables[junction_name != ""],
@@
   geom_hline(yintercept = -log10(1 - .9)) + 
+  geom_vline(xintercept = -0,linetype="dotted") + 
      scale_fill_manual(name = "",
@@
 
-fig2a = ggplot(data = de_table) + 
+fig1b = ggplot(data = de_table) + 
   geom_point(data = de_table %>% filter(contains_cryptic == "0"),
          aes(x = log2FoldChange, y = -log10(padj),
-             alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 4) + 
+             alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 10) + 
     geom_point(data = de_table %>% filter(contains_cryptic == "1"),
          aes(x = log2FoldChange, y = -log10(padj),
-             alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 4) + 
+             alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 10)+ 
     geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
@@
 
+print(fig1a)
+print(fig1b)
 ```
@@
 
-We have 2 PSI's for the UNC13A cryptic, one which includes both the short and
-long form of the cryptic, and one which does not. While we're not mentioning in 
-figure 1, given that the longer form of the cryptic appears in control cerebellum, 
-one could argue that that junction should not be included in the PSI calculation
-here for UNC13A cryptic per se. In practice, this makes very little difference in the
-conclusions made on the cell lines, so I've included both for completeness.
-
 The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
-The NB cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
+The SK-N-DZ cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
 There was variability on the Klim hMN set on allelic expression.
-```{r}
+
+```{r read in }
 
@@
                                           UNC13A_5prime_3)]
-####barplot UNC13A CE PSI - full five####
-unc13a_cryptic_full = rel_rna_cryptic_amount %>% 
+```
+
+## barplot UNC13A CE PSI - full five
+
+```{r}
+rel_rna_cryptic_amount %>% 
     ggbarplot(,
@@
     guides(color = FALSE) +
+    scale_y_continuous(labels = scales::percent) +
     theme(text = element_text(size = 20,family = 'sans'), 
@@
           axis.text.y = element_text(size = 28))
-
-####barplot UNC13B NMD PSI####
-unc13b_psi_nmd = rel_rna_cryptic_amount %>% 
+```
+
+## barplot UNC13B NMD PSI
+
+```{r}
+rel_rna_cryptic_amount %>% 
     ggbarplot(,
@@
           axis.text.y = element_text(size = 28))
+```
+
+#### TARDBP RNA and UNC13A Cryptic
+```{r}
 
@@
     theme(legend.position = 'bottom') 
-####UNC13A RNA and UNC13A Cryptic####
+```
+
+#### UNC13A RNA and UNC13A Cryptic####
+
+```{r}
 rel_rna_cryptic_amount %>% 
@@
     geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
-    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12) +
+    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12,
+             method = 'spearman',
+           cor.coef.name = 'rho') +
     geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
@@
     scale_x_continuous(labels = scales::percent) +
+      scale_y_continuous(labels = scales::percent) +
     ylab("UNC13A CE PSI") +
@@
           axis.text = element_text(size = 32))
-####UNC13A RNA and UNC13A Cryptic####
-rel_rna_cryptic_amount %>% 
-    filter(condition != "control") %>% 
-    ggplot() + 
-    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
-    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
-    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
-    ggpubr::theme_pubr() +
-    theme(text = element_text(size = 20)) +
-    scale_x_continuous(labels = scales::percent) +
-    ylab("UNC13A CE PSI") +
-    expand_limits(y = 1) + 
-    theme(legend.title=element_blank()) +
-    theme(legend.position = 'bottom') +
-      theme(text = element_text(size = 18,family = 'sans'), 
-          legend.text = element_text(size = 20,family = 'sans'),
-          axis.title = element_text(size = 32),
-          axis.text = element_text(size = 32))
-####UNC13A RNA and UNC13A IR####
+```
+#### UNC13A RNA and UNC13A IR
+```{r}
 rel_rna_cryptic_amount %>% 
@@
     geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
-    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12) +
+    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
     geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
@@
     theme(legend.position = 'bottom') 
-####TARDBP RNA and UNC13A IR####
+```
+#### TARDBP RNA and UNC13A IR
+
+```{r}
 rel_rna_cryptic_amount %>% 
@@
     theme(legend.position = 'bottom') 
-
-unca = rel_rna_cryptic_amount %>% 
+```
+
+#### barplot UNC13A relative RNA across experiments
+
+```{r}
+rel_rna_cryptic_amount %>% 
   ggbarplot(,
@@
   guides(color = FALSE) 
-
+```
+
+```{r}
 stmn2 = rel_rna_cryptic_amount %>% 
@@
 
-ggarrange(unc13a_cryptic_full,unc13b_psi_nmd,unc13a_ir,unc13b_ir,common.legend = T)
 ####scatterplot stmn2 normalized TARDBP####
-
 rel_rna_cryptic_amount %>% 
@@
     stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
-    theme(legend.position = 'bottom') + 
-  facet_wrap(~source)
+    theme(legend.position = 'bottom') 
 ####scatterplot UNC13BIR normalized TARDBP####
@@
   facet_wrap(~source)
-####scatterplot UNC13BIR and UNC13B NMD####
+```
+
+#### scatterplot UNC13BIR and UNC13B NMD
+
+```{r}
 rel_rna_cryptic_amount %>% 
@@
 
-####barplot UNC13A RNA level####
-
-```
-
-# Paient summary statistics
-First let's look at only including tissues with detectable stmn2 or UNC13A CE
+```
+
+#### barplot protein abundances iPSC KD
+
+```{r}
+peptides = fread(file.path(here::here(),"data","peptide_abdundance.csv"))
+
+peptides %>% 
+  mutate(gene = fct_relevel(gene, "UNC13A","UNC13B")) %>% 
+  ggbarplot(,
+            x = "condition",
+            add = c("mean_se"),
+            y = "protein",
+            fill = 'condition',
+            color = 'condition',
+            position = position_dodge(0.8),
+            facet.by = 'gene') + 
+  ggpubr::theme_pubr() +
+  scale_fill_manual(
+      values = c("#40B0A6","#E1BE6A")
+  ) + 
+  scale_color_manual(
+      values = c("#1C2617","#262114")
+  ) +
+  ylab("Protein abundance") + 
+  xlab("") + 
+  geom_jitter(pch = 21,height = 0,aes(fill = condition),size = 2) + 
+  guides(color = FALSE) +
+  facet_wrap(~gene, scales = 'free') +
+  theme(legend.position = 'none') + 
+  scale_y_continuous(labels = scientific) + 
+  stat_compare_means(comparisons = list(c("Control","TDP-43 KD")),
+                     label = 'p.signif',tip.length = 0,
+                     size = 10)
+
+```
+
+# Patient summary statistics
+
 ```{r}
@@
 the ALS sample tissues are not balanced and not every participant has samples in every tissue)
+
 ```{r}
@@
 # Detection of UNC13A
-Looking at disease tissue only, so just taking the cords in ALS and the 
+Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the 
 frontal and temporal cortex of FTLD and then the cord and cortices in Controls.
+
 ```{r}
@@
 
-# Detection UNC13A by tissue
+# UNC13A detection rate by tissue
 
@@
   ggplot() +
-  geom_col(aes(x = detection_name, y = detection_rate,fill = tissue_clean),position = position_dodge2(width = 0.8, preserve = "single")) + 
+  geom_col(aes(x = detection_name, y = detection_rate,fill = detection_name),position = position_dodge2(width = 0.8, preserve = "single")) + 
   ggpubr::theme_pubr() + 
@@
   theme(axis.text.x =  element_text(size = 14)) + 
-      theme_bw() + 
     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
   theme(text = element_text(size = 18))  +
-  scale_fill_manual(values = c("#4C97B5","#C87156","#4C97B5","#C87156","#4C97B5","#C87156")) +
+  scale_fill_manual(values = c("#4C97B5","#4C97B5","#4C97B5","#4C97B5","#4C97B5","red","#C87156","#C87156")) +
   theme(legend.position = "none") 
@@
   theme(text = element_text(size = 18))  +
-  theme(legend.position = "none") 
-```
-
-# Junction type expression
+  theme(legend.position = "bottom") 
+```
+
+# Junction type expression by tissue - shorter exon only collapsed
 
@@
 boxplot_table %>% 
-  filter(tissue_clean %in% c("Frontal Cortex", "Lumbar Spinal Cord", "Cervical Spinal Cord","Motor Cortex", "Temporal Cortex", "Cerebellum") )%>% 
+    filter(tissue_clean %in% 
+               c("Frontal Cortex", 
+                 "Lumbar Spinal Cord", 
+                 "Cervical Spinal Cord",
+                 "Motor Cortex", 
+                 "Temporal Cortex",
+                 "Cerebellum") )%>% 
+    group_by(disease_group2,tissue_clean) %>% 
+    mutate(n_sample = n_distinct(sample)) %>% 
+    ungroup() %>% 
+    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
+    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
     mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
@@
                                       "Motor Cortex")) %>% 
-        mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
-  dplyr::select(disease_group2,
+    ggplot(aes(x = detection_name, 
+               y = inclusion_reads * 10^6,
+               fill ="red")) + 
+    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1)) + 
+    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
+    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) + 
+    ylab("UNC13A cryptic \n reads per million") +
+    theme(text = element_text(size = 12)) + 
+    xlab("") +
+    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
+    theme_bw() + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+    theme(text = element_text(size = 18))  +
+    theme(legend.position = "bottom") 
+```
+
+# Junction type expression by tissue - the longer acceptor
+
+```{r}
+clean_data_table %>% 
+    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
+    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
+    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
+    melt() %>% 
+    filter(grepl("_3prime|_5prime_1|_5prime_2",variable)) %>% 
+      # filter(grepl("_5prime_2",variable)) %>% 
+    unique() %>% 
+    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
+                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
+                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
+    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  dplyr::select(sample,disease_group2,
                 tissue_clean,
@@
     unique() %>% 
-    ggplot(aes(x = disease_group2, 
+    filter(tissue_clean %in% 
+               c("Frontal Cortex", 
+                 "Lumbar Spinal Cord", 
+                 "Cervical Spinal Cord",
+                 "Motor Cortex", 
+                 "Temporal Cortex",
+                 "Cerebellum"))%>% 
+    group_by(disease_group2,tissue_clean) %>% 
+    mutate(n_sample = n_distinct(sample)) %>% 
+    ungroup() %>% 
+    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
+    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
+    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
+                                      "Frontal Cortex",
+                                      "Lumbar Spinal Cord",
+                                      "Temporal Cortex",
+                                      "Motor Cortex")) %>% 
+    ggplot(aes(x = detection_name, 
                y = value * 10^6,
                fill = junction_name)) + 
-    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1)) + 
-    geom_point(pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
+    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single")) + 
+    geom_point(size = 2,pch = 21, alpha = 0.7, position = position_jitterdodge(jitter.width = 0.2))+ 
     scale_y_log10() + 
@@
     xlab("") +
-    scale_fill_manual(values = colorblind_pal()(4)[2:4]) 
+    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
+        theme_bw() + 
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+  theme(text = element_text(size = 18))  +
+  theme(legend.position = "bottom") 
 ```
@@
 
-# UNC13A inclusion separated by tissue and not type
-
-```{r}
-
-boxplot_table %>% 
-  filter(junction_name %in% c("  Novel Donor"," Short Novel Acceptor" )) %>% 
-  group_by(sample) %>% 
-  summarise(new_inc = sum(value),
-            disease_group2 = disease_group2,
-            tissue_clean = tissue_clean,
-            sample = sample) %>% 
+# UNC13A TPM by detection status and tissue
+
+```{r}
+
+clean_data_table %>% 
+  filter(UNC13A_annotated_leaf > 10) %>% 
+  mutate(unc13_d = ifelse(unc13a_cryptic_leaf_psi > 0,"detected","not-detected")) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  group_by(disease_group2,tissue_clean) %>% 
+  mutate(n_sample = n_distinct(sample)) %>% 
   ungroup() %>% 
-  unique() %>% 
-    filter(!(tissue_clean %in% c("Hippocampus","Occipital Cortex","Cerebellum"))) %>% 
-    ggplot(aes(x = disease_group2, y = new_inc * 10^6)) + 
-    geom_boxplot() + 
-    geom_jitter(height = 0) + 
-    scale_y_log10() + 
-    facet_wrap(~tissue_clean) +
-    ggpubr::theme_pubr() +
-    ylab("UNC13A cryptic inclusion \n reads per million") + 
-    theme(text = element_text(size = 24)) + 
-    xlab("N samples")
-
-```
-
-# UNC13A Inclusion Reads by type and tissue
-
-```{r}
-####ALS reads per tissue####
-boxplot_table %>% 
-  filter(disease_group2 %in% c("Control",
-                              "ALS \n non-TDP",
-                              "ALS-TDP")) %>% 
-      filter(tissue_clean %in% c("Motor Cortex","Cervical Spinal Cord","Lumbar Spinal Cord")) %>% 
-  dplyr::select(disease_group2,
-                tissue_clean,
-                junction_name,
-                value) %>% 
-  unique() %>% 
-  mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
-    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex")) %>% 
-      mutate(y_val = value * 10^6) %>% 
-    ggbarplot( x = "disease_group2", 
-               y = "y_val", 
-               color = 'junction_name',
-               fill = 'junction_name',
-               position = position_dodge(0.8),
-               add = c("mean_se","jitter"),
-               facet.by = c("junction_name","tissue_clean"))+ 
-    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
-       scale_color_manual(
-        values = c("#1C2617","#262114","#262115")
-    ) +
-    theme(axis.text.x =  element_text(size = 24)) + 
-    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
-    theme(text = element_text(size = 24))  +
-    ylab("UNC13A cryptic \n reads per million")  +
-    xlab("") + 
-    theme(legend.position = 'none') +
-    theme(legend.title = element_blank()) +
-  ylim(0,0.2) + 
-  theme(
-  strip.background.y = element_blank(),
-  strip.text.y  = element_blank()
-) +
-  guides(color = FALSE)
-
-####FTLD reads per tissue####
-boxplot_table %>% 
-    filter(tissue_clean %in% c("Cerebellum","Frontal Cortex","Temporal Cortex")) %>% 
-    dplyr::select(disease_group2,
-                tissue_clean,
-                junction_name,
-                value) %>% 
-      unique() %>% 
-      mutate(y_val = value * 10^6) %>% 
-      mutate(tissue_clean = fct_relevel(tissue_clean,"Cerebellum",after = Inf)) %>% 
-      mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
-    ggbarplot( x = "disease_group2", 
-               y = "y_val", 
-               color = 'junction_name',
-               fill = 'junction_name',
-               position = position_dodge(0.8),
-               add = c("mean_se","jitter"),
-               facet.by = c("junction_name","tissue_clean")) + 
-    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
-       scale_color_manual(
-        values = c("#1C2617","#262114","#262115")
-    ) +
-    theme(axis.text.x =  element_text(size = 24)) + 
-    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
-    theme(text = element_text(size = 24))  +
-    ylab("UNC13A cryptic \n reads per million")  +
-    xlab("") + 
-    theme(legend.position = 'top') +
-    theme(legend.title = element_blank()) +
-    ylim(0,0.2) +
-    theme(
-  strip.background.y = element_blank(),
-  strip.text.y  = element_blank()
-) +
-    scale_x_discrete(guide = guide_axis(n.dodge=2)) +
-  guides(color = FALSE)
-
-```
-
-# UNC13A TPM across disease states
-
-Separated by tissue only in ALS/FTLD-TDP
-```{r}
-
-clean_data_table %>% 
-  filter(UNC13A_annotated_leaf > 10) %>% 
-  mutate(unc13_d = unc13a_cryptic_leaf_psi_full > 0) %>% 
-    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>%  
+  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
+     filter(tissue_clean %in% 
+               c("Frontal Cortex", 
+                 "Lumbar Spinal Cord", 
+                 "Cervical Spinal Cord",
+                 "Motor Cortex", 
+                 "Temporal Cortex",
+                 "Cerebellum") )%>% 
+  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
+                                "Frontal Cortex",
+                                "Lumbar Spinal Cord",
+                                "Temporal Cortex",
+                                "Motor Cortex")) %>% 
     ggbarplot(,
-              x = "disease_group2",
+              x = "detection_name",
               add = c("mean_se","jitter"),
@@
     theme(text = element_text(size = 20)) +
-  facet_wrap(~tissue_clean, scales = 'free') +
-  ggtitle("UNC13A Cryptic Detected") +
-  stat_compare_means()
+  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
+  ggtitle("UNC13A Cryptic Detected - all tissue") +
+  stat_compare_means(aes(x  = detection_name, fill = unc13_d,y = UNC13A_TPM))
 ```
@@
 # Correlation between UNC13A CE PSI and UNC13A TPM
-
-```{r}
-#tissue separated
-clean_data_table %>% 
-  filter(unc13a_cryptic_leaf_psi_full > 0 ) %>% 
-    filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  ggplot(aes(x = unc13a_cryptic_leaf_psi_full, y = UNC13A_TPM, color = disease_group2)) + 
+## tissue separated
+
+```{r}
+clean_data_table %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) + 
   geom_point() + 
@@
   theme(text = element_text(size = 18)) +
-  facet_wrap(~tissue_clean,scales = 'free')
-
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  filter(unc13a_cryptic_leaf_psi_full > 0) %>% 
-  ggplot(aes(x = unc13a_cryptic_leaf_psi_full, y = UNC13A_TPM)) + 
+  facet_wrap(~tissue_clean,scales = 'free') +
+  scale_x_continuous(labels = scales::percent)
+
+```
+
+## not separated -
+
+```{r}
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0) %>% 
+  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13B_TPM)) + 
   geom_point() + 
-  stat_cor(label.y = 38,size = 8) + 
+  stat_cor(label.y = 38,size = 8,method = 's',cor.coef.name = 'rho') + 
   geom_smooth(method = lm, se = FALSE,color = "black") +
@@
   theme(text = element_text(size = 22)) +
-  ylim(0,40)
+  scale_x_continuous(labels = scales::percent)
 
@@
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
   filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
@@
   ggpubr::theme_pubr() + 
-  theme(text = element_text(size = 18)) 
+  theme(text = element_text(size = 18)) +
+  scale_y_continuous(labels = scales::percent)  +
+  scale_x_continuous(labels = scales::percent) 
 ```
@@
   facet_wrap(~disease_group2, scales = 'free_x')
-####boxplot of full fix cryptic psi in als-tdp and ftld-tdp by genotype and tissue####
+```
+
+# boxplot of cryptic psi in als-tdp and ftld-tdp by genotype and tissue
+
+```{r}
 clean_data_table %>% 
@@
   facet_wrap(~disease_group2, scales = 'free')
+
+```
+
+```{r}
+cortex_names = clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = (unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1)) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(call) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  .$detection_name %>% unique()
+
+```
+
+# boxplot of UNC over STMN2 psi in als-tdp and ftld-tdp by genotype and tissue
+
+```{r}
+cortex_comparisons = list(c("C/C \n( 121 )", "G/G \n( 54 )"),c("G/G \n( 54 )", "C/G \n( 110 )"),c("C/C \n( 121 )", "C/G \n( 110 )"))
+detected_comparisons = list(c("C/C \n( 26 )", "G/G \n( 21 )"),c("G/G \n( 21 )", "C/G \n( 26 )"),c("C/C \n( 26 )", "C/G \n( 26 )"))
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
+  group_by(call) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = detection_name)) + 
+  geom_boxplot() + 
+  geom_jitter(alpha = 0.3, height = 0) + 
+  stat_compare_means() +
+  ylab("STMN2 PSI")
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+    group_by(call) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+    mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call)) + 
+  geom_boxplot() + 
+  geom_jitter(alpha = 0.3, height = 0) + 
+  stat_compare_means() +
+  ylab("STMN2 PSI")
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+    group_by(call) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call)) + 
+  geom_boxplot() + 
+  geom_jitter(alpha = 0.3, height = 0) + 
+  stat_compare_means() +
+  facet_wrap(~disease_group2) +
+  stat_compare_means( label = "p.format") + 
+  # stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif") +
+  ylab("STMN2 PSI")
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+    group_by(call) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call, fill = call)) + 
+  geom_boxplot() + 
+  geom_jitter(alpha = 0.3, height = 0) + 
+  stat_compare_means() +
+  facet_wrap(~disease_group2) +
+  ylab("UNC13A PSI") +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24))
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = (unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1)) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(call) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
+  geom_boxplot(show.legend = F) + 
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
+  stat_compare_means( label = "p.format") + 
+  stat_compare_means(aes(group = detection_name),comparisons = cortex_comparisons,label = "p.signif") +
+  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 1,size = 1.5) +
+  ggtitle("All Cortex Samples")
+
+```
+
+```{r}
 
@@
 ```{r}
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(call) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
+  geom_boxplot(show.legend = F) + 
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
+  stat_compare_means( label = "p.format") + 
+  stat_compare_means(aes(group = detection_name),comparisons = detected_comparisons,label = "p.signif") +
+  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
+  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 0,size = 1.5)
+
 tissue_to_show_correlation = clean_data_table %>% 
@@
     mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
-    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
-  filter(is.finite(log10_unc13a) & is.finite(log10_stmn2))
+    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi))
 
 tissue_to_show_correlation %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
     mutate(call = fct_relevel(call,
@@
   theme(legend.position =  'top') +
-  theme(legend.title=element_blank())
+  theme(legend.title=element_blank()) +
+  stat_compare_means()
 

### 2021-02-11 paper_figures.Rmd
making rmd pretty

@@
 
+```{css, echo = FALSE}
+h1, #TOC>ul>li {
+  color: #000000;
+    font-weight: bold;
+}
+
+h2, #TOC>ul>ul>li {
+  color: #41494D;
+}
+
+```
+
 ```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
+knitr::opts_chunk$set(
+  comment = '', fig.width = 9, fig.height = 9
+)
 if (!require("pacman")) install.packages("pacman")
@@
 
-#### TARDBP RNA and UNC13A Cryptic
+## TARDBP RNA and UNC13A Cryptic
+
 ```{r}
@@
 
-#### UNC13A RNA and UNC13A Cryptic####
+## UNC13A RNA and UNC13A Cryptic
 
@@
 ```
-#### UNC13A RNA and UNC13A IR
+## UNC13A RNA and UNC13A IR
 ```{r}
@@
 ```
-#### TARDBP RNA and UNC13A IR
+## TARDBP RNA and UNC13A IR
 
@@
 
-#### barplot UNC13A relative RNA across experiments
+## barplot UNC13A relative RNA across experiments
 
@@
 
-#### scatterplot UNC13BIR and UNC13B NMD
+## scatterplot UNC13BIR and UNC13B NMD
 
@@
 
-#### barplot protein abundances iPSC KD
+# iPSC protein abundance mass spec
 
@@
 ```
-# Detection of UNC13A
+## Pecent of patients UNC13A detected
 Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the 
@@
 
-# UNC13A detection rate by tissue
+## Pecent of samples UNC13A detected by tissue
 
@@
 
-# Detection UNC13A by tissue and Genotype
+## Pecent of samples UNC13A detected by tissue and genotype
 
@@
 
-# Junction type expression by tissue - shorter exon only collapsed
-
-```{r}
+## UNC13A CE level across tissue
+
+This only plots the short novel acceptor and novel donor
+
+```{r onlyshort}
 boxplot_table %>% 
@@
 
-# Junction type expression by tissue - the longer acceptor
+## UNC13A CE level across tissue - split by junction type
 
@@
 
+## UNC13A CE level across disease types
+
 ```{r}
@@
 
-# UNC13A TPM by detection status and tissue
-
-```{r}
-
-clean_data_table %>% 
-  filter(UNC13A_annotated_leaf > 10) %>% 
-  mutate(unc13_d = ifelse(unc13a_cryptic_leaf_psi > 0,"detected","not-detected")) %>% 
-  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+# UNC13A RNA expression (TPM)
+
+## UNC13A TPM by disease and tissue
+
+```{r}
+comps = unique(clean_data_table$disease_group2) %>% gsub("Control", " Control",.) %>% 
+  combn(.,2,simplify = F)
+
+clean_data_table %>% 
   group_by(disease_group2,tissue_clean) %>% 
@@
   ungroup() %>% 
-  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>%  
+  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
   mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  filter(tissue_clean %in% 
@@
                               "Motor Cortex")) %>% 
-    ggbarplot(,
-              x = "detection_name",
-              add = c("mean_se","jitter"),
-              y = "UNC13A_TPM",
-              fill = 'unc13_d',
-              color = 'unc13_d',
-              facet.by = 'tissue_clean',
-              position = position_dodge(0.8)) + 
-    ggpubr::theme_pubr() +
-    scale_fill_manual(name = "",
-        values = c("#40B0A6","#E1BE6A")
-    ) + 
-    scale_color_manual(name = "",
-        values = c("#1C2617","#262114")
-    ) +
+    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
+    ggpubr::theme_pubr() +
+    geom_boxplot(aes(fill = disease_group2)) + 
+    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
     ylab("UNC13A TPM") + 
@@
   facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
-  ggtitle("UNC13A Cryptic Detected - all tissue") +
-  stat_compare_means(aes(x  = detection_name, fill = unc13_d,y = UNC13A_TPM))
-```
-
-# Correlation between STMN2 Cryptic PSI and UNC13A TPM
-
-Separated by tissue only in ALS/FTLD-TDP
-```{r}
-
+  theme(legend.position = 'none')
+```
+
+## Correlation STMN2 Cryptic PSI and UNC13A TPM
+
+only in ALS/FTLD-TDP
+
+```{r}
 clean_data_table %>% 
@@
   theme(text = element_text(size = 18)) +
-  facet_wrap(~tissue_clean,scales = "free_y")
-```
-
-No Tissue separation
-
-```{r}
+  ylim(0,50)
+```
+
+## Correlation STMN2 Cryptic PSI and UNC13A TPM - tissue separations
+
+```{r}
+
 clean_data_table %>% 
@@
   theme(text = element_text(size = 18)) +
-  ylim(0,50)
-```
-
-# Correlation between UNC13A CE PSI and UNC13A TPM
-## tissue separated
-
-```{r}
-clean_data_table %>% 
-  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
-    filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) + 
-  geom_point() + 
-  stat_cor(size = 8) + 
-  geom_smooth(method = lm, se = FALSE,color = "black") +
-  ylab("UNC13A TPM") +
-  xlab("UNC13A CE PSI") + 
-  ggpubr::theme_pubr() + 
-  theme(text = element_text(size = 18)) +
-  facet_wrap(~tissue_clean,scales = 'free') +
-  scale_x_continuous(labels = scales::percent)
-
-```
-
-## not separated -
+  facet_wrap(~tissue_clean,scales = "free_y")
+```
+
+## Correlation between UNC13A CE PSI and UNC13A TPM
 
@@
   theme(text = element_text(size = 22)) +
+  scale_x_continuous(labels = scales::percent)
+
+```
+
+## Correlation between UNC13A CE PSI and UNC13A TPM - tissue separated
+
+```{r}
+clean_data_table %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+    filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) + 
+  geom_point() + 
+  stat_cor(size = 8) + 
+  geom_smooth(method = lm, se = FALSE,color = "black") +
+  ylab("UNC13A TPM") +
+  xlab("UNC13A CE PSI") + 
+  ggpubr::theme_pubr() + 
+  theme(text = element_text(size = 18)) +
+  facet_wrap(~tissue_clean,scales = 'free') +
   scale_x_continuous(labels = scales::percent)
@@
 
-# STMN2 and UNC13A CE PSI in tissue
-
-```{r}
-clean_data_table %>% 
-    filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi_full) %>% 
-  mutate(unc13a_normalized_psi_to_stmn2 = log2((unc13a_cryptic_leaf_psi_full +1)/ (stmn_2_cryptic_psi_leaf + 1))) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  group_by(tissue_clean) %>% 
-  mutate(psi_mean = median(unc13a_normalized_psi_to_stmn2)) %>% 
-  ungroup() %>% 
-  mutate(tissue_clean = fct_reorder(tissue_clean,-psi_mean)) %>% 
-  ggbarplot( x = "tissue_clean", 
-               y = "unc13a_normalized_psi_to_stmn2", 
-              position = position_dodge(0.8),
-              add = c("mean_se","jitter"),
-             facet.by = 'disease_group2') +
-  ylab("Log2Fold Ratio \n UNC13A Cryptic to STMN2 Cryptic") + 
-  xlab("") +
-  theme(text = element_text(size = 24)) +
-  theme(legend.position = 'bottom') +
-  facet_wrap(~disease_group2, scales = 'free_x')
-```
-
-# boxplot of cryptic psi in als-tdp and ftld-tdp by genotype and tissue
-
-```{r}
-clean_data_table %>% 
-    filter(disease_tissue == T) %>% 
-  filter(unc13a_cryptic_leaf_psi_full > 0 ) %>% 
+# Effect of UNC13A genotype
+
+## UNC13A psi by genotype
+```{r}
+genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
   dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  ggboxplot( x = "tissue_clean", 
-               y = "unc13a_cryptic_leaf_psi", 
-             fill = "call",
-             facet.by = 'disease_group2') +
-  xlab("") +
-  geom_jitter(height = 0) +
-  theme(text = element_text(size = 24)) +
-  theme(legend.position = 'bottom') +
-  facet_wrap(~disease_group2, scales = 'free')
-
-```
-
-```{r}
-cortex_names = clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(grepl("Cortex",tissue_clean)) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(log2_unc = (unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1)) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  group_by(call) %>% 
-  mutate(n_sample = length(unique(sample))) %>% 
-  ungroup() %>% 
-  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
-  .$detection_name %>% unique()
-
-```
-
-# boxplot of UNC over STMN2 psi in als-tdp and ftld-tdp by genotype and tissue
-
-```{r}
-cortex_comparisons = list(c("C/C \n( 121 )", "G/G \n( 54 )"),c("G/G \n( 54 )", "C/G \n( 110 )"),c("C/C \n( 121 )", "C/G \n( 110 )"))
-detected_comparisons = list(c("C/C \n( 26 )", "G/G \n( 21 )"),c("G/G \n( 21 )", "C/G \n( 26 )"),c("C/C \n( 26 )", "C/G \n( 26 )"))
-
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
-  group_by(call) %>% 
-  mutate(n_sample = length(unique(sample))) %>% 
-  ungroup() %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
-  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = detection_name)) + 
-  geom_boxplot() + 
-  geom_jitter(alpha = 0.3, height = 0) + 
-  stat_compare_means() +
-  ylab("STMN2 PSI")
-
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-    group_by(call) %>% 
-  mutate(n_sample = length(unique(sample))) %>% 
-  ungroup() %>% 
-    mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
-  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call)) + 
-  geom_boxplot() + 
-  geom_jitter(alpha = 0.3, height = 0) + 
-  stat_compare_means() +
-  ylab("STMN2 PSI")
-
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-    group_by(call) %>% 
-  mutate(n_sample = length(unique(sample))) %>% 
-  ungroup() %>% 
-  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
-  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call)) + 
-  geom_boxplot() + 
-  geom_jitter(alpha = 0.3, height = 0) + 
-  stat_compare_means() +
-  facet_wrap(~disease_group2) +
-  stat_compare_means( label = "p.format") + 
-  # stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif") +
-  ylab("STMN2 PSI")
-
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(grepl("Cortex",tissue_clean)) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1))) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
   group_by(call) %>% 
@@
   geom_boxplot() + 
-  geom_jitter(alpha = 0.3, height = 0) + 
-  stat_compare_means() +
-  facet_wrap(~disease_group2) +
+  geom_jitter( height = 0) + 
+  facet_wrap(~disease_group2,scales = 'free') +
   ylab("UNC13A PSI") +
@@
   ggpubr::theme_pubr() +
-  theme(text = element_text(size = 24))
-
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(grepl("Cortex",tissue_clean)) %>% 
+  theme(text = element_text(size = 24)) +
+  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")
+
+```
+
+## UNC13A psi by genotype - tissue split
+```{r}
+genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
   dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(log2_unc = (unc13a_cryptic_leaf_psi +1 )/(stmn_2_cryptic_psi_leaf +1)) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
   group_by(call) %>% 
@@
   mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
-  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
-  geom_boxplot(show.legend = F) + 
-  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
-  stat_compare_means( label = "p.format") + 
-  stat_compare_means(aes(group = detection_name),comparisons = cortex_comparisons,label = "p.signif") +
-  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
+  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call, fill = call)) + 
+  geom_boxplot() + 
+  geom_jitter( height = 0) + 
+  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
+  ylab("UNC13A PSI") +
   xlab("UNC13A rs12973192 Genotype") +
@@
   theme(text = element_text(size = 24)) +
-  geom_hline(yintercept = 1,size = 1.5) +
-  ggtitle("All Cortex Samples")
-
-```
-
-```{r}
-
-clean_data_table %>% 
-    filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,STMN2_TPM,UNC13A_TPM) %>% 
-  mutate(unc13a_normalized_psi_to_stmn2 = log2((UNC13A_TPM +1)/ (STMN2_TPM + 1))) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  group_by(tissue_clean) %>% 
-  mutate(psi_mean = median(unc13a_normalized_psi_to_stmn2)) %>% 
+  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")
+
+```
+
+## UNC13A psi in detected tissues by genotype - tissue split
+```{r}
+
+clean_data_table %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  group_by(call) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
   ungroup() %>% 
-  mutate(tissue_clean = fct_reorder(tissue_clean,-psi_mean)) %>% 
-  ggbarplot( x = "tissue_clean", 
-               y = "unc13a_normalized_psi_to_stmn2", 
-              position = position_dodge(0.8),
-              add = c("mean_se","jitter"),
-             facet.by = 'disease_group2') +
-  ylab("Log2Fold Ratio \n UNC13A Cryptic to STMN2 Cryptic") + 
-  xlab("") +
+  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call, fill = call)) + 
+  geom_boxplot() + 
+  geom_jitter( height = 0) + 
+  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
+  ylab("UNC13A PSI") +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
   theme(text = element_text(size = 24)) +
-  theme(legend.position = 'bottom') +
-  facet_wrap(~disease_group2, scales = 'free_x')
-
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(participant_id,  sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi_full) %>%  add_count(participant_id) %>% 
-  filter(n == 5) %>% 
-  dplyr::select(-n) %>% 
-  data.table::melt() %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  ggplot(aes(x = tissue_clean, y = value, color = participant_id)) + 
-  geom_point() +
-  geom_line(aes(group = participant_id)) +
+  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")
+
+```
+
+## STMN2 psi by genotype 
+```{r}
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call, fill = call)) + 
+  geom_boxplot() + 
+  geom_jitter( height = 0) + 
+  facet_wrap(~disease_group2,scales = 'free') +
+  ylab("STMN2 PSI") +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
   theme(text = element_text(size = 24)) +
-  theme(legend.position = 'none')  +
-  facet_grid(~variable)
-  scale_color_manual(values = c("#D55E00","#0072B2"))
-
-```
-
-# Correlation UNC13A PSI and STMN2 PSI with genotype effect
-
-```{r}
+  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")
+
+```
+
+## STMN2 psi by genotype - tissue split
+```{r}
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call, fill = call)) + 
+  geom_boxplot() + 
+  geom_jitter( height = 0) + 
+  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
+  ylab("STMN2 PSI") +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")
+
+```
+
+## STMN2 psi in detected tissues by genotype - tissue split
+```{r}
+
+clean_data_table %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call, fill = call)) + 
+  geom_boxplot() + 
+  geom_jitter( height = 0) + 
+  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
+  ylab("STMN2 PSI") +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")
+
+```
+
+## Correlation UNC13A PSI and STMN2 PSI with genotype effect
+
+```{r}
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+    mutate(call = fct_relevel(call,
+                              "C/C", "C/G", "G/G")) %>%
+  filter(!grepl("Cord",tissue_clean)) %>% 
+  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
+    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
+    ggpubr::ggscatter(., 
+                      x = "log10_stmn2",
+                      y = "log10_unc13a",
+                      color = 'call',                      
+                      add = "reg.line", size = 3
+                      ) +
+    ylab("UNC13A CE PSI ") + 
+    xlab("STMN2 Cryptic PSI ") +
+  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
+                      y = unc13a_cryptic_leaf_psi,
+                      color = call),
+           method = 'spearman',
+           cor.coef.name = 'rho',
+           show.legend = FALSE,
+           size = 14) + 
+    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  theme(legend.title = element_blank()) +
+  theme(text = element_text(size = 24),legend.text = element_text(size = 32))
+
+```
+
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected
+
+```{r}
+detected_comparisons = list(c("C/C \n( 26 )", "G/G \n( 21 )"),c("G/G \n( 21 )", "C/G \n( 26 )"),c("C/C \n( 26 )", "C/G \n( 26 )"))
 clean_data_table %>% 
@@
   geom_hline(yintercept = 0,size = 1.5) 
-
-tissue_to_show_correlation = clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-    mutate(call = fct_relevel(call,
-                              "C/C", "C/G", "G/G")) %>%
-  filter(!grepl("Cord",tissue_clean)) %>% 
-    mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
-    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi))
-
-tissue_to_show_correlation %>% 
-    mutate(call = fct_relevel(call,
-                              "C/C", "C/G", "G/G")) %>%
-  filter(!grepl("Cord",tissue_clean)) %>% 
-  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
-    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
-    ggpubr::ggscatter(., 
-                      x = "log10_stmn2",
-                      y = "log10_unc13a",
-                      color = 'call',                      
-                      add = "reg.line", size = 3
-                      ) +
-    ylab("UNC13A CE PSI ") + 
-    xlab("STMN2 Cryptic PSI ") +
-  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
-                      y = unc13a_cryptic_leaf_psi,
-                      color = call),
-           method = 'spearman',
-           cor.coef.name = 'rho',
-           show.legend = FALSE,
-           size = 14) + 
-    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
-  theme(legend.title = element_blank()) +
-  theme(text = element_text(size = 24),legend.text = element_text(size = 32))
-
-```
-
-# Detection rate by genotype
-
-## Barplot of detection by genotype
+```
+
+## Detection rate by genotype
 
@@
   dplyr::select(-n_sample)
-
-over_p = overall_fisher %>% 
-  column_to_rownames('call') %>% 
-  chisq.test() %>% 
-  broom::tidy() %>% 
-  .$p.value
-
-overall_fisher %>% 
-  mutate(n_sample = n_non_detected + n_sample_detected) %>% 
-  mutate(detection_rate = n_sample_detected / n_sample) %>% 
-  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
-  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
-  geom_col(show.legend = F) + 
-  ggpubr::theme_pubr() + 
-  scale_y_continuous(labels = scales::percent) + 
-  ylab("% of TDP-43 Proteionopathy Patients \n UNC13A Cryptic Detected") + 
-  theme(text = element_text(size = 18)) + 
-  xlab("N individuals") + 
-  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))
-
-```
-
-## Barplot of detection by genotype and tissue
-
-```{r}
-clean_data_table %>%
-    filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
-  dplyr::select(sample,unc13a_detected,call,tissue_clean) %>% 
-  unique() %>% 
-  group_by(call,tissue_clean) %>% 
-  mutate(n_sample = n_distinct(sample)) %>% 
-  mutate(n_sample_detected = sum(unc13a_detected)) %>% 
-  dplyr::select(call,n_sample,n_sample_detected) %>% 
-  unique() %>% 
-  mutate(n_non_detected = n_sample - n_sample_detected) %>% 
-  mutate(detection_rate = n_sample_detected / n_sample) %>% 
-  ggplot(aes(x = tissue_clean, y = detection_rate, fill = call)) +
-  geom_col(show.legend = F,position = position_dodge2()) + 
-  geom_text(aes(label = n_sample,y = 0),vjust = 1,size = 11,position = position_dodge(width = 1)) + 
-  ggpubr::theme_pubr() + 
-  scale_y_continuous(labels = scales::percent) + 
-  ylab("% of Tissue samples \n UNC13A Cryptic Detected") + 
-  theme(text = element_text(size = 18)) + 
-  xlab("N individuals") + 
-  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))
 
@@
 
-UNC13A CE PSI
-
-```{r}
-genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))
-
-clean_data_table %>%   
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-    mutate(call = fct_relevel(call,
-                              "C/C", "C/G", "G/G")) %>%
-  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
-  filter(stmn_2_cryptic_psi_leaf > 0) %>% 
-    ggbarplot( x = "call", 
-               y = "unc13a_cryptic_leaf_psi", 
-               color = 'call',
-              position = position_dodge(0.8),
-              add = c("mean_se","jitter")) + 
-    ylab("UNC13A CE PSI") + 
-    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
-    theme(text = element_text(size = 20)) + 
-    stat_compare_means(aes(group = call), label = "p.format") + 
-    stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif")
-```
-
-```{r}
-clean_data_table %>%   
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-    mutate(call = fct_relevel(call,
-                              "C/C", "C/G", "G/G")) %>%
-    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi), n = 2))) %>%
-    mutate(stmn2_psi_groups = ifelse(is.na(stmn2_psi_groups),"  No STMN2",stmn2_psi_groups)) %>%
-    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
-                                        stmn2_psi_groups == 2 ~ "High STMN2",
-                                        TRUE ~ stmn2_psi_groups)) %>%
-    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
-    ggbarplot( x = "call", 
-               y = "cryptic_psi", 
-               color = 'call',
-               facet.by = 'stmn2_psi_groups',
-              position = position_dodge(0.8),
-              add = c("mean_se","jitter")) + 
-    ylab("UNC13A CE PSI") + 
-    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
-    theme(text = element_text(size = 20)) + 
-    stat_compare_means(aes(group = call), label = "p.format") + 
-    stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif")  + 
-  xlab("")
-```
-
 # Variant allele expression in TDP-43 KD experiments
 
-Across our KD's, we observed a clear allelic bias in the NB cells
+Across our KD's, we observed a clear allelic bias in the SK-DZ-N cells
 
@@
 
-kd_vafs %>% 
-    left_join(rel_rna_cryptic_amount) %>% 
-    filter(!is.na(condition)) %>% 
-    filter(condition != "control" ) %>%     
-    mutate(VAF = alt_reads / total) %>% 
-    filter(source %in% c("sh_dzap","shsy5y_normed","ipsc_normed_ward")) %>% 
-    dplyr::select(sample,ref_reads,alt_reads,TARDBP,source,VAF) %>% 
-    data.table::melt(id.vars = c("sample","TARDBP",'source','VAF')) %>% 
-    mutate(sample = fct_reorder(sample,-TARDBP)) %>% 
-    ggplot(aes(x = sample,y = value, fill = variable)) + 
-    geom_col(pch = 21, size = 4) +
-    coord_flip() 
 ```
@@
 # Liu Facs Neurons
+
 Also look at the coverage in the Liu nuclear facs sorted neurons
@@
 ```
-
-# UNC13A TPM by tissue and disease
-
-```{r}
-ftld_comp <- list( c("Control", "FTLD \n non-TDP"), 
-                        c("Control", "FTLD-TDP"), c("FTLD-TDP", "FTLD \n non-TDP") )
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(tissue_clean == "Frontal_Cortex") %>% 
-  ggplot(aes(x = disease_group2, y = UNC13A_TPM)) + 
-  geom_boxplot() + 
-  facet_wrap(~tissue_clean,scales = "free") + 
-  stat_compare_means(comparisons = ftld_comp)
-```

### 2021-02-24 paper_figures.Rmd
Add code for generating side by side stmn2 unc13a graph

@@
 clean_data_table = clean_data_table %>%     
-    mutate(call = fct_relevel(call,
+    mutate(rs12973192 = fct_relevel(rs12973192,
                               "C/C", "C/G", "G/G")) %>% 
-    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
+    mutate(number_g_alleles = as.numeric(rs12973192) - 1) %>% 
     mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi))
@@
 clean_data_table[,length(unique(sample)),by = disease]
-print("Number of patients per rs12973192 genotype")
-clean_data_table[,length(unique(participant_id)),by = call]
+print("Number of patients per UNC13A SNP genotype")
+unique(clean_data_table[,.(participant_id,rs12608932,rs12973192)]) %>% select(-participant_id) %>% table()
 print("Number of tissues per disease")
@@
 boxplot_table %>% 
-  left_join(clean_data_table %>% dplyr::select(sample, call)) %>% 
+  left_join(clean_data_table %>% dplyr::select(sample, rs12973192)) %>% 
   unique() %>% 
@@
   mutate(detected = inclusion_reads > 0) %>% 
-  dplyr::select(sample,disease_group2,detected,tissue_clean,call) %>% 
+  dplyr::select(sample,disease_group2,detected,tissue_clean,rs12973192) %>% 
   unique() %>% 
-  group_by(disease_group2,tissue_clean,call) %>% 
+  group_by(disease_group2,tissue_clean,rs12973192) %>% 
   mutate(n_sample = n_distinct(sample)) %>% 
@@
   ggplot() +
-  geom_col(aes(x = detection_name, y = detection_rate,fill = call),position = position_dodge2(width = 0.8, preserve = "single")) + 
+  geom_col(aes(x = detection_name, y = detection_rate,fill = rs12973192),position = position_dodge2(width = 0.8, preserve = "single")) + 
   ggpubr::theme_pubr() + 
@@
 
-# Relationship between STMN2 and UNC13A CE PSI in patients
-```{r}
-####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
+# Relationship between STMN2 and UNC13A CE PSI 
+## Scatter plot showing the correlation in non-log space in cortex UNC13A & STMN2 CE
+
+```{r}
 clean_data_table %>% 
@@
 
+## Side by side comparison UNC13A & STMN2 CE PSI
+```{r}
+clean_data_table %>% 
+  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  # filter(unc13a_cryptic_leaf_psi > 0) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  melt() %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
+    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
+                                    "Frontal Cortex",
+                                    "Lumbar Spinal Cord",
+                                    "Temporal Cortex")) %>% 
+  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
+  ggplot(aes(x = variable, y = value)) +
+  geom_boxplot(aes(color = variable)) +
+  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
+  scale_fill_manual(values = c("#004D40","#882255")) +
+  scale_color_manual(values = c("#004D40","#882255")) +
+  facet_wrap(~tissue_clean,nrow = 3,scales = 'free') +
+    scale_y_continuous(labels = scales::percent)  +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24))  +
+  xlab("") +
+  ylab("PSI")  +
+  theme(legend.position = 'none')
+
+```
+
+## Side by side comparison UNC13A & STMN2 CE PSI v2 
+```{r}
+clean_data_table %>% 
+  filter(UNC13A_annotated_leaf > 10) %>% 
+  filter(STMN2_annotated_leaf > 10) %>% 
+  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+  # filter(unc13a_cryptic_leaf_psi > 0) %>%
+  # filter(disease_tissue == T) %>% 
+    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
+  filter(tissue_clean %in% c("Motor Cortex", "Cervical Spinal Cord", "Frontal Cortex", "Lumbar Spinal Cord", "Temporal Cortex")) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_rat = log2((unc13a_cryptic_leaf_psi + 0.1) / (stmn_2_cryptic_psi_leaf + 0.1))) %>% 
+  group_by(tissue_clean, disease_group2) %>% 
+  mutate(average_log2Fold = mean(log2_rat)) %>% 
+  ungroup() %>% 
+  dplyr::select(-log2_rat) %>% 
+  pivot_longer(cols = c("stmn_2_cryptic_psi_leaf","unc13a_cryptic_leaf_psi")) %>% 
+  mutate(tissue_clean = fct_reorder(tissue_clean,average_log2Fold)) %>% 
+  mutate(variable = ifelse(grepl("stmn_2",name), "STMN2", "UNC13A")) %>%
+  ggplot(aes(x = tissue_clean, y = value,fill = name)) +
+  geom_boxplot() +
+  geom_point(pch = 21,height = 0,width = 0.2,position = position_jitterdodge()) +
+  scale_fill_manual(values = c("#004D40","#882255")) +
+  facet_wrap(~tissue_clean+disease_group2,nrow = 1,scales = 'free')  +
+  theme(text = element_text(size = 24))  +
+  xlab("") +
+  ylab("PSI")  +
+  theme(legend.position = 'none')
+```
+
+## Side by side comparison UNC13A & STMN2 CE PSI - v3
+```{r}
+clean_data_table %>% 
+  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  # filter(unc13a_cryptic_leaf_psi > 0) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  melt() %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
+    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
+                                    "Frontal Cortex",
+                                    "Lumbar Spinal Cord",
+                                    "Temporal Cortex")) %>% 
+  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
+  ggplot(aes(x = variable, y = value)) +
+  geom_boxplot(aes(color = variable)) +
+  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
+  scale_fill_manual(values = c("#004D40","#882255")) +
+  scale_color_manual(values = c("#004D40","#882255")) +
+  facet_wrap(~tissue_clean+disease_group2,nrow = 3,scales = 'free') +
+    scale_y_continuous(labels = scales::percent)  +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24))  +
+  xlab("") +
+  ylab("PSI")  +
+  theme(legend.position = 'none')
+
+```
+
+## Log2FC UNC13A & STMN2 CE PSI
+```{r}
+clean_data_table %>% 
+  filter(UNC13A_annotated_leaf > 10) %>% 
+  filter(STMN2_annotated_leaf > 10) %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
+
+  # mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>% 
+  ggplot(aes(x = tissue_clean,y = log2_unc, fill = tissue_clean)) + 
+  geom_boxplot(position = 'dodge') +
+  # geom_jitter(height = 0,width = 0.2,pch = 21) +
+  facet_wrap(~disease_group2,scales = "free_x") + 
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+  theme(text = element_text(size = 24))  +
+  xlab("") +
+  ylab("PSI")  +
+  theme(legend.position =  'top') +
+  theme(legend.title=element_blank()) +
+  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
+```
+
 # Effect of UNC13A genotype
@@
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
   mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
-  group_by(call) %>% 
+  group_by(rs12973192) %>% 
   mutate(n_sample = length(unique(sample))) %>% 
   ungroup() %>% 
-  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
-  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = call, fill = call)) + 
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) + 
   geom_boxplot() + 
@@
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
   mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
-  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = call, fill = call)) + 
+  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) + 
   geom_boxplot() + 
@@
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-    mutate(call = fct_relevel(call,
+    mutate(rs12973192 = fct_relevel(rs12973192,
                               "C/C", "C/G", "G/G")) %>%
@@
                       y = "log10_unc13a",
-                      color = 'call',                      
+                      color = 'rs12973192',                      
                       add = "reg.line", size = 3
@@
                       y = unc13a_cryptic_leaf_psi,
-                      color = call),
+                      color = rs12973192),
            method = 'spearman',
@@
 
-## Log2 Fold UNC / STMN2 CE in TDP-path with both detected
-
-```{r}
-detected_comparisons = list(c("C/C \n( 26 )", "G/G \n( 21 )"),c("G/G \n( 21 )", "C/G \n( 26 )"),c("C/C \n( 26 )", "C/G \n( 26 )"))
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(grepl("Cortex",tissue_clean)) %>% 
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex only
+
+```{r}
+detected_table_for_graph = clean_data_table %>% 
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
@@
   filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(rs12973192) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
+  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))
+
+comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)
+
+detected_table_for_graph %>% 
+  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
+  geom_boxplot(show.legend = F) + 
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
+  stat_compare_means( label = "p.format") + 
+  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
+  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
+  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 0,size = 1.5) 
+```
+
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord =
+
+```{r}
+detected_table_for_graph = clean_data_table %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(rs12973192) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
+  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))
+
+comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)
+
+detected_table_for_graph %>% 
+  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
+  geom_boxplot(show.legend = F) + 
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
+  stat_compare_means( label = "p.format") + 
+  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
+  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
+  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 0,size = 1.5) 
+```
+
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932
+
+```{r}
+detected_table_for_graph_rs12608932 = clean_data_table %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(rs12608932) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))
+
+comparisons = combn(unique(detected_table_for_graph_rs12608932$detection_name),2,simplify = F)
+
+detected_table_for_graph_rs12608932 %>% 
+  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
+  geom_boxplot(show.legend = F) + 
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
+  stat_compare_means( label = "p.format") + 
+  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
+  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
+  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
+  xlab("UNC13A rs12608932 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 0,size = 1.5) 
+```
+
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932
+
+```{r}
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
   mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>% 
   mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  group_by(call) %>% 
+  group_by(rs12608932) %>% 
   mutate(n_sample = length(unique(sample))) %>% 
   ungroup() %>% 
-  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
   ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
@@
   ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
-  xlab("UNC13A rs12973192 Genotype") +
+  xlab("UNC13A rs12608932 Genotype") +
   scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
@@
   mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
-  dplyr::select(participant_id,unc13a_detected,call) %>% 
+  dplyr::select(participant_id,unc13a_detected,rs12973192) %>% 
   unique() %>% 
-  group_by(call) %>% 
+  group_by(rs12973192) %>% 
   mutate(n_sample = n_distinct(participant_id)) %>% 
   mutate(n_sample_detected = sum(unc13a_detected)) %>% 
-  dplyr::select(call,n_sample,n_sample_detected) %>% 
+  dplyr::select(rs12973192,n_sample,n_sample_detected) %>% 
   unique() %>% 
@@
 over_p = overall_fisher %>% 
-  column_to_rownames('call') %>% 
+  column_to_rownames('rs12973192') %>% 
   chisq.test() %>% 
@@
   mutate(detection_rate = n_sample_detected / n_sample) %>% 
-  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
   ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
@@
 a p-value of `r round(over_p,2)`.
-
-Only looking at disease relevant tissue and in patients 
-we see that the detection rate depends on both the level of TPD-43 pathology present, as measured by STMN2 CE expression,
-
-```{r}
-detection_table = clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-    mutate(call = fct_relevel(call,
-                              "C/C", "C/G", "G/G")) %>%
-    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi), n = 2))) %>%
-    mutate(stmn2_psi_groups = ifelse(is.na(stmn2_psi_groups),"  No STMN2",stmn2_psi_groups)) %>%
-    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
-                                        stmn2_psi_groups == 2 ~ "High STMN2",
-                                        TRUE ~ stmn2_psi_groups)) %>%
-  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
-  group_by(call,unc13a_detected,stmn2_psi_groups) %>% 
-  add_count(name = "genotype_detected") %>% 
-  dplyr::select(call,unc13a_detected,genotype_detected,stmn2_psi_groups) %>% 
-  unique() %>% 
-  pivot_wider(names_from = "unc13a_detected",
-              values_from = "genotype_detected") %>% 
-  dplyr::rename(unc_not_detected = `FALSE`, unc_cryptic_detected = `TRUE`) %>%
-  mutate(total_tissue = (unc_not_detected + unc_cryptic_detected)) %>% 
-  mutate(detection_rate = unc_cryptic_detected / total_tissue) %>% 
-  ungroup() %>% 
-  mutate(stmn2_psi_groups = fct_relevel(stmn2_psi_groups, "No STMN2", after = 0)) %>% 
-  mutate(detection_name = glue::glue("{call} \n ( {total_tissue} )"))
-
-detection_table %>% 
-  mutate(error = sqrt((detection_rate * (1-detection_rate))/total_tissue)) %>% 
-  ggplot(aes(x = call, y = detection_rate,fill = call)) + 
-  geom_col(show.legend = F,position = 'dodge2') + 
-  geom_errorbar(aes(ymin = detection_rate - error, ymax = detection_rate + error),
-                  position = position_dodge(0.9),
-                width=0.2) +
-  facet_wrap(~stmn2_psi_groups) +
-  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) + 
-  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
-  ylab("Percent of Samples Detected") + 
-  ggpubr::theme_pubr() + 
-  geom_text(aes(label = total_tissue,y = 0),vjust = 1,size = 11) + 
-  ggpubr::theme_pubr() +
-    theme(text = element_text(size = 32)) +
-  xlab("N Tissues") 
-```
 
@@
   theme(legend.title=element_blank()) +
-  stat_compare_means()
+  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
 
@@
     left_join(clean_data_table) %>% 
-    filter(call == "C/G") %>% 
+    filter(rs12973192 == "C/G") %>% 
     mutate(fraction_risk = alt_reads / total) %>% 

### 2021-03-02 paper_figures.Rmd
added nmd
sy5y nmd data sheet and relative script to produce fig 2d

@@
 
+```
+
+# NMD experiment on sy5y cells
+
+```{r}
+
+for_bar_plot_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
+
+for_bar_plot_sy <- data.frame(for_bar_plot_sy)
+for_bar_plot_sy$condition <- as.factor(for_bar_plot_sy$condition)
+for_bar_plot_sy$condition <- factor(for_bar_plot_sy$condition,levels(for_bar_plot_sy$condition)[c(2,1)])
+
+nmd_plot = ggbarplot(for_bar_plot_sy,
+                        x = 'gene',
+                        add = c("mean_se","jitter"),
+                        y = 'value',
+                        fill = 'condition',
+                        color = 'condition',
+                        position = position_dodge(0.8),
+                        dot.size = 10) + 
+  scale_y_continuous(labels = scales::percent) + 
+  ggpubr::theme_pubr() +
+  scale_fill_manual(
+    values = c("#40B0A6","#E1BE6A")
+  ) + 
+  scale_color_manual(
+    values = c("#1C2617","#262114")
+  )
+
+plot(nmd_plot)
 ```

### 2021-03-03 paper_figures.Rmd
Update paper_figures.Rmd

@@
 title: "Figures for Paper"
-author: "Anna-Leigh Brown"
+author: "Anna-Leigh Brown + Matteo Zanovello"
 date: "06/10/2020"

### 2021-03-09 paper_figures.Rmd
including data for the plots of the fold changes

@@
   theme(legend.text=element_text(size=22)) +
-  xlim(-1,1)
+  xlim(-1,1) +
+  scale_x_continuous(labels = scales::percent)
 
@@
     ggbarplot(,
-              x = "source",
+              x = "condition",
               add = c("mean_se","jitter"),
@@
               color = 'condition',
-              position = position_dodge(0.8)) + 
+              position = position_dodge(0.8),facet.by = 'source') + 
     ggpubr::theme_pubr() +
@@
           axis.title.y = element_text(size = 28),
-          axis.text.y = element_text(size = 28))
+          axis.text.y = element_text(size = 28)) +
+    facet_wrap(~source) + 
+  stat_compare_means() +
+  scale_y_continuous(labels = scales::percent)
 ```
@@
     geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) + 
-    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
+    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12,method = 's',cor.coef.name = 'rho') +
     geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
@@
     theme(legend.title=element_blank()) +
-    theme(legend.position = 'bottom') 
+    theme(legend.position = 'bottom') +
+    scale_y_continuous(labels = scales::percent)
 ```
@@
     stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
-    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
+    geom_smooth(aes(x = UNC13A, y = normalized_unc13a_ir),color = "black",method = 'lm') +
     ggpubr::theme_pubr() +
@@
     stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
-    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
+    geom_smooth(aes(x = TARDBP, y = normalized_unc13a_ir),color = "black",method = 'lm') +
     ggpubr::theme_pubr() +
@@
 
-## barplot UNC13A relative RNA across experiments
+## barplot UNC13A cryptic PSI
 
@@
 
+## barplot STMN2 cryptic PSI
+
 ```{r}
@@
   guides(color = FALSE)
-
-####barplot UNC13B normalized IR####
+print(stmn2)
+```
+
+## barplot UNC13B normalized IR
+```{r}
 unc13b_ir = rel_rna_cryptic_amount %>% 
@@
           axis.text.y = element_text(size = 28))
-####barplot UNC13A normalized IR####
+
+print(unc13b_ir)
+```
+
+## barplot UNC13A normalized IR
+```{r}
 unc13a_ir = rel_rna_cryptic_amount %>% 
@@
           axis.text.y = element_text(size = 28))
-
-####scatterplot stmn2 normalized TARDBP####
+```
+
+## barplot UNC13A normalized RNA Level
+
+```{r}
+text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))
+
+rel_rna_cryptic_amount %>% 
+    ggbarplot(,
+              x = "source",
+              add = c("mean_se","jitter"),
+              y = "UNC13A",
+              fill = 'condition',
+              color = 'condition',
+              position = position_dodge(0.8)) + 
+    ggpubr::theme_pubr() +
+    scale_fill_manual(
+        values = c("#40B0A6","#E1BE6A")
+    ) + 
+    scale_color_manual(
+        values = c("#1C2617","#262114")
+    ) +
+    ylab("UNC13A") + 
+    xlab("") + 
+    theme(legend.position = "none") +
+    guides(color = FALSE) + 
+      theme(text = element_text(size = 20,family = 'sans'), 
+          legend.text = element_text(size = 36,family = 'sans'),
+          axis.title.y = element_text(size = 28),
+          axis.text.y = element_text(size = 28)) +
+    scale_y_continuous(labels = scales::percent) +
+    geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
+    geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)
+
+```
+
+## barplot UNC13B normalized RNA Level
+
+```{r}
+text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))
+
+rel_rna_cryptic_amount %>% 
+    ggbarplot(,
+              x = "source",
+              add = c("mean_se","jitter"),
+              y = "UNC13A",
+              fill = 'condition',
+              color = 'condition',
+              position = position_dodge(0.8)) + 
+    ggpubr::theme_pubr() +
+    scale_fill_manual(
+        values = c("#40B0A6","#E1BE6A")
+    ) + 
+    scale_color_manual(
+        values = c("#1C2617","#262114")
+    ) +
+    ylab("UNC13B") + 
+    xlab("") + 
+    theme(legend.position = "none") +
+    guides(color = FALSE) + 
+      theme(text = element_text(size = 20,family = 'sans'), 
+          legend.text = element_text(size = 36,family = 'sans'),
+          axis.title.y = element_text(size = 28),
+          axis.text.y = element_text(size = 28)) +
+    scale_y_continuous(labels = scales::percent) +
+    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
+    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)
+
+```
+
+## barplot TARDBP normalized RNA Level
+
+```{r}
+text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))
+
+rel_rna_cryptic_amount %>% 
+    ggbarplot(,
+              x = "source",
+              add = c("mean_se","jitter"),
+              y = "TARDBP",
+              fill = 'condition',
+              color = 'condition',
+              position = position_dodge(0.8)) + 
+    ggpubr::theme_pubr() +
+    scale_fill_manual(
+        values = c("#40B0A6","#E1BE6A")
+    ) + 
+    scale_color_manual(
+        values = c("#1C2617","#262114")
+    ) +
+    ylab("TARDBP") + 
+    xlab("") + 
+    theme(legend.position = "none") +
+    guides(color = FALSE) + 
+      theme(text = element_text(size = 20,family = 'sans'), 
+          legend.text = element_text(size = 36,family = 'sans'),
+          axis.title.y = element_text(size = 28),
+          axis.text.y = element_text(size = 28)) +
+    scale_y_continuous(labels = scales::percent) +
+    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
+    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)
+
+```
+## Scatterplot stmn2 normalized TARDBP
+```{r}
 rel_rna_cryptic_amount %>% 
@@
     theme(legend.position = 'bottom')
-####scatterplot UNC13BIR normalized TARDBP####
+```
+
+## Scatterplot UNC13BIR normalized TARDBP
+
+```{r}
 rel_rna_cryptic_amount %>% 
@@
     theme(legend.position = 'bottom')
-####scatterplot UNC13AIR and UNC13A Crptic####
+```
+
+## Scatterplot UNC13A-IR and UNC13A Crptic
+```{r}
 rel_rna_cryptic_amount %>% 
@@
     ggplot() + 
-    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
-    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13a_ir)) +
+    geom_point(aes(x = cryptic_psi_full, 
+                   y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
+    stat_cor(aes(x = cryptic_psi_full, 
+                 y = normalized_unc13a_ir)) +
     theme(legend.position = 'bottom')
-####scatterplot UNC13BIR and UNC13A Crptic####
+```
+
+## Scatterplot UNC13B-IR and UNC13A Crptic
+
+```{r}
 rel_rna_cryptic_amount %>% 
@@
 
-# NMD experiment on sy5y cells
+# NMD experiment on SH-SY5Y cells
 
@@
     values = c("#1C2617","#262114")
-  )
+  ) +
+  ylab("Percent of transcript after CHX treatment")
 
@@
                fill ="red")) + 
-    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1)) + 
+    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA)) + 
     geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
@@
   geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
-  scale_fill_manual(values = c("#004D40","#882255")) +
-  scale_color_manual(values = c("#004D40","#882255")) +
+  scale_fill_manual(values = c("#008574","#cd5582")) +
+  scale_color_manual(values = c("#008574","#cd5582")) +
   facet_wrap(~tissue_clean,nrow = 3,scales = 'free') +
@@
 # Effect of UNC13A genotype
-
-## UNC13A psi by genotype
+## UNC13A psi by genotype - rs12973192
+```{r}
+rs12973192_comparisons = clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  group_by(rs12973192) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
+  pull(detection_name) %>% 
+  unique() %>% 
+  combn(.,2,simplify = F)
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  group_by(rs12973192) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12973192)) + 
+  geom_boxplot() + 
+  geom_jitter( height = 0) + 
+  ylab("UNC13A PSI") +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  stat_compare_means() +
+  stat_compare_means(comparisons = rs12973192_comparisons,label = "p.signif") + 
+  scale_y_continuous(trans='log10')
+```
+
+## UNC13A psi by genotype - rs12608932
+```{r}
+
+rs12608932_comparisons = clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  group_by(rs12608932) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
+  pull(detection_name) %>% 
+  unique() %>% 
+  combn(.,2,simplify = F)
+
+clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
+  group_by(rs12608932) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
+  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12608932)) + 
+  geom_boxplot() + 
+  geom_jitter( height = 0) + 
+  ylab("UNC13A PSI") +
+  xlab("UNC13A rs12608932 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  stat_compare_means() +
+  stat_compare_means(comparisons = rs12608932_comparisons,label = "p.signif") + 
+  scale_y_continuous(trans='log10')
+
+```
+
+## UNC13A psi by genotype - rs12973192
 ```{r}
@@
 
-## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord =
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord
 
@@
   stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
-  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
-  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
-  xlab("UNC13A rs12608932 Genotype") +
-  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
-  ggpubr::theme_pubr() +
-  theme(text = element_text(size = 24)) +
-  geom_hline(yintercept = 0,size = 1.5) 
-```
-
-## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932
-
-```{r}
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(grepl("Cortex",tissue_clean)) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
-  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  group_by(rs12608932) %>% 
-  mutate(n_sample = length(unique(sample))) %>% 
-  ungroup() %>% 
-  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
-  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
-  geom_boxplot(show.legend = F) + 
-  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
-  stat_compare_means( label = "p.format") + 
-  stat_compare_means(aes(group = detection_name),comparisons = detected_comparisons,label = "p.signif") +
   ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 

### 2021-03-09 paper_figures.Rmd
updated script
added script to create nmd plot in dz cells (sup fig 2b)

--- 2021-03-02 paper_figures.Rmd
@@
 plot(nmd_plot)
+```
+
+```{r}
+
+for_bar_plot_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
+
+for_bar_plot_dz <- data.frame(for_bar_plot_dz)
+for_bar_plot_dz$condition <- as.factor(for_bar_plot_dz$condition)
+for_bar_plot_dz$condition <- factor(for_bar_plot_dz$condition,levels(for_bar_plot_dz$condition)[c(2,1)])
+
+nmd_plot2 = ggbarplot(for_bar_plot_dz,
+                        x = 'gene',
+                        add = c("mean_se","jitter"),
+                        y = 'value',
+                        fill = 'condition',
+                        color = 'condition',
+                        position = position_dodge(0.8),
+                        dot.size = 10) + 
+  scale_y_continuous(labels = scales::percent) + 
+  ggpubr::theme_pubr() +
+  scale_fill_manual(
+    values = c("#40B0A6","#E1BE6A")
+  ) + 
+  scale_color_manual(
+    values = c("#1C2617","#262114")
+  )
+
+plot(nmd_plot2)
 ```

### 2021-03-14 oscar_figures.Rmd
Riboseq volcano plots

---
title: "oscar_figures"
author: "Oscar Wilkins"
date: "14/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DESeq2)
library(here)
library(ggrepel)
```

## Ribosome profiling quality control

Perform quality control to check that reads are periodic and enriched across the llllllllllllllll

## Ribosome profiling

Use DESeq to detect differential ribosome footprints and create volcano plots

```{r message = FALSE}
gene_names <- read_csv(paste0(here(), "/data/gencode_v29_names.csv"))
counts_df <- read_csv(paste0(here(), "/data/riboseq/cds_feature_counts.csv"))
counts <- as.matrix(counts_df %>% column_to_rownames("gene_id"))
ipsc_splicing <- read_csv(paste0(here(), "/data/ipsc_splicing_results.csv"))
ipsc_de = read_csv(paste0(here(), "/data/ipsc_differential_expression.csv"))

col_data <- data.frame(sample = colnames(counts)) %>%
  mutate(condition = ifelse(str_detect(sample, "tdp"), "tdp_ko", "control"))

dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = col_data,
                              design= ~condition)
dds <- DESeq(dds)
results_name <- resultsNames(dds)[2]
res <- results(dds, name=results_name)

res_df <- data.frame(res) %>%
  rownames_to_column("gene_id") %>%
  left_join(gene_names, by = "gene_id")

cryptic_df <- ipsc_splicing %>%
  dplyr::select(gene_name, cryptic_junction) %>%
  unique() %>%
  filter(cryptic_junction)

res_df2 <- res_df %>%
  mutate(cryptic_junction = gene_name %in% cryptic_df$gene_name) %>%
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2","TARDBP") ~ gene_name)) %>%
  mutate(colour = ifelse(cryptic_junction, "#fe6101",  "#648FFF")) %>%
  arrange(colour)
```

```{r} 
p1 <- ggplot(res_df2, aes(x = log2FoldChange, y = -log10(pvalue), label = label_junction)) +
  geom_point(data = res_df2 %>% filter(!cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#648FFF") +
  geom_point(data = res_df2 %>% filter(cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#fe6101") +
  ggpubr::theme_pubr() +
  ylab("-Log10(p-value)") +
  xlab("Log2 fold change") +
  geom_text_repel(box.padding = 1.2, max.overlaps = Inf)

p2 <- p1 + ylim(0,10)
p1
p2
```

### 2021-03-14 oscar_figures.Rmd
Add periodicity plot

@@
 
-Perform quality control to check that reads are periodic and enriched across the llllllllllllllll
+Perform quality control to check that reads are periodic.
+
+```{r message=FALSE, warning=FALSE}
+read_length_filter <- 29  # length of reads of interest
+gencode_v29_info <- read_tsv(paste0(here(), "/data/longest_proteincoding_transcript_hs_details.txt"))
+ribo_df <- read_tsv(paste0(here(), "/data/riboseq/sh_tdpb.Aligned.toTranscriptome.out.bam.bed.gz"), 
+                    col_names = c("transcript_id", "start0", "end", "strand")) %>%
+    inner_join(gencode_v29_info, by = "transcript_id") %>%
+    dplyr::filter(end-start0 == read_length_filter) %>%
+    mutate(distance_from_start = start0-cds_start) %>%
+    group_by(distance_from_start) %>%
+    mutate(n_this_dist = n()) %>%
+    dplyr::select(distance_from_start, n_this_dist) %>%
+    unique()
+
+```
+
+```{r, warning=FALSE}
+ggplot(ribo_df, aes(x=distance_from_start, y=n_this_dist)) +
+    geom_bar(stat="identity") +
+    xlim(-50, 50) +
+    ylab("Counts") + 
+    xlab("Distance of 5' end from annotated start codon") +
+    ggtitle(paste0("Periodicity of ", read_length_filter, "nt reads from TDP-43 KD replicate B"))
+```
 
@@
 
-```{r message = FALSE}
-gene_names <- read_csv(paste0(here(), "/data/gencode_v29_names.csv"))
+```{r message = FALSE, warning=FALSE}
 counts_df <- read_csv(paste0(here(), "/data/riboseq/cds_feature_counts.csv"))
@@
   rownames_to_column("gene_id") %>%
-  left_join(gene_names, by = "gene_id")
+  left_join(gencode_v29_info, by = "gene_id")
 
@@
 
-```{r} 
+```{r, warning=FALSE} 
 p1 <- ggplot(res_df2, aes(x = log2FoldChange, y = -log10(pvalue), label = label_junction)) +

### 2021-03-14 oscar_figures.Rmd
Add barplot of targeted RNA-seq

@@
     inner_join(gencode_v29_info, by = "transcript_id") %>%
-    dplyr::filter(end-start0 == read_length_filter) %>%
+    dplyr::filter(end-start0 == read_length_filter, strand == "+") %>%
     mutate(distance_from_start = start0-cds_start) %>%
@@
 
-## Ribosome profiling
+## Ribosome profiling volcano plot
 
@@
 ```
+
+## Targeted RNA-seq
+
+Read in data from processed bam files
+
+```{r message=FALSE, warning=FALSE}
+specific_rt <- read_csv(paste0(here(), "/data/targeted_RNA_seq/specific_RT_primer/for_plot.csv")) %>%
+    select(risk, sample_name, n)
+
+random_hexamer <- read_csv(paste0(here(), "/data/targeted_RNA_seq/random_hexamer/values_for_plot.csv")) %>%
+    select(-sample_name) %>%
+    select(sample_name = actual_sample_name, risk, n)
+
+tg_rnaseq_df <- bind_rows(specific_rt, random_hexamer) %>%
+  group_by(sample_name) %>%
+  mutate(n_healthy = ifelse(risk == "Non-Risk", n, 0)) %>%
+  mutate(combined_healthy = sum(n_healthy)) %>%
+  select(-n_healthy) %>%
+  mutate(combined = sum(n)) %>%
+  group_by(sample_name, risk) %>%
+  mutate(combined_n = sum(n)) %>%
+  ungroup() %>%
+  select(-n) %>%
+  unique() %>%
+  group_by(sample_name) %>%
+  mutate(p_value = pbinom(combined_healthy, sum(combined_n), 0.5)) %>%  # single-tailed bionomial test
+  mutate(colour = ifelse(risk == "Non-Risk", "lightblue", "red")) %>%
+  ungroup() %>%
+  mutate(sample_name_ordered = fct_reorder(sample_name, dplyr::desc(combined_n)))
+
+ggplot(tg_rnaseq_df %>%
+         filter(combined_n > 0), aes(x = sample_name_ordered, y = combined_n, fill = risk)) +
+  geom_bar(stat="identity", position="dodge") +
+  geom_text(aes(x=sample_name, y=-2.5, label=signif(p_value,2)), size = 3.5) +
+  xlab("") +
+  ylab("Unique cDNAs") +
+  ggeasy::easy_add_legend_title("Allele") +
+  theme_minimal() +
+  ggpubr::theme_pubr() +
+  ggeasy::easy_rotate_x_labels()
+```
+
+## iCLIP
+
+##

### 2021-03-14 oscar_figures.Rmd
Add iCLIP peak plots

@@
 
+TODO: add scripts which deduplicate UMIs for each.
+
 ```{r message=FALSE, warning=FALSE}
@@
 
-## iCLIP
+## iCLIP of minigenes
+
+Read in the processed data (following alignment with Bowtie2, and filtering for high-confidence alignments), and generate
+a plot of the overall distribution of crosslinks, and the rolling average difference between the 2x Risk and 2x Healthy
+
+TODO: add script which does this filtering
+
+```{r message=FALSE, warning=FALSE}
+library(patchwork)
+
+exon_snp <- 556 
+intron_snp <- 1106
+intron_start <- 141 
+intron_end <- 1429
+acceptor1 <- 394
+acceptor2 <- 444
+donor <- 572
+mg_l <- 1624 
+cpoor_start <- 490
+cpoor_end <- 1325
+risk_colour <- "#00BFC4"
+healthy_colour <- "#F8766D"
+
+filtered_df <- read_csv(paste0(here(), "/data/iCLIP/bowtie2_aligned_2_filtered_df.csv"))
+
+rolling_window <- 20
+
+df2 <- filtered_df %>%
+  ungroup() %>%
+  group_by(condition, pos) %>%
+  mutate(mean_counts = sum(normalised_counts)/2) %>%
+  select(pos, condition, mean_counts) %>%
+  unique() %>%
+  pivot_wider(names_from = "condition", values_from ="mean_counts")
+
+df2[is.na(df2)] <- 0
+
+df2 <- df2 %>% 
+  mutate(diff = R-H) %>%
+  mutate(mean_4 = (H+R)/2) %>%
+  arrange(pos)
+
+top <- ggplot(df2, aes(x=pos, y=diff)) +
+  geom_area(aes(x=pos, y = mean_4), colour="grey50", fill = "grey50", alpha=0.2) +
+  geom_vline(xintercept = exon_snp, linetype="dashed") +
+  geom_vline(xintercept = intron_snp, linetype="dashed") +
+  xlim(1,1600) +
+  #ylim(-1,5) +
+  theme_minimal() +
+  ylab("XLs per 1000") +
+  xlab("") +
+  geom_vline(xintercept = cpoor_start) +
+  geom_vline(xintercept = cpoor_end)
+
+bottom <- ggplot(df2, aes(x=pos, y=zoo::rollmean(diff, k = rolling_window, na.pad=T))) +
+  geom_area(colour = "grey50", fill="grey50", alpha = 1) +
+  geom_vline(xintercept =  exon_snp, linetype="dashed") +
+  geom_vline(xintercept =  intron_snp, linetype="dashed") +
+  xlim(1,1600) +
+  ylim(-0.5,0.75) +
+  theme_minimal() +
+  ylab("Change in XL density") +
+  xlab("Distance from start of minigene") +
+  geom_vline(xintercept =  intron_start) +
+  geom_vline(xintercept =  intron_end) +
+  geom_vline(xintercept =  acceptor1) +
+  geom_vline(xintercept =  acceptor2) +
+  geom_vline(xintercept =  donor)
+
+combined <- top/bottom
+combined
+```
+
+Next, filter for peaks in the iCLIP signal, and see how they change in 2x Risk versus 2x Healthy. Highlight those which are close to the SNPs.
+
+```{r message=FALSE, warning=FALSE}
+
+find_ratio_bound <- function(e_x, sem_x, e_y, sem_y, quantile_value, randoms = 10000){
+  # monte carlo confidence interval estimate for a ratio. Assumes the two variables
+  # x and y are normally distributed. Finds confidence interval for (x-y)/y
+
+  # randomly generate possible global means for x and y
+
+  xs = rnorm(randoms, mean = e_x, sd=sem_x)  # risk
+  ys = rnorm(randoms, mean = e_y, sd=sem_y)  # healthy
+
+  ratios = (xs-ys)/ys
+
+  return(quantile(ratios, quantile_value))
+}
+
+## Plot big changes
+
+confidence_interval <- 0.75  # fracion, i.e. 0.75 = 75%.
+
+peak_threshold <- 5 # how many times above the mean to be a "peak"
+region_length <- 50
+
+peak_t <- peak_threshold*1000/mg_l
+
+peak_df <- filtered_df %>%
+  ungroup() %>%
+  group_by(pos) %>%
+  mutate(max_at_pos = max(normalised_counts)) %>%
+  ungroup() %>%
+  filter(max_at_pos >= peak_t) %>%
+  group_by(condition, pos) %>%
+  mutate(peak_mean = mean(normalised_counts), 
+         peak_sd = sd(normalised_counts)) %>%
+  ungroup() %>%
+  select(condition, pos, peak_mean, peak_sd) %>%
+  unique() %>%
+  pivot_wider(names_from = "condition", values_from = c("peak_mean", "peak_sd")) %>%
+  mutate(frac_increase = (peak_mean_R-peak_mean_H)/peak_mean_H) %>%
+  mutate(peak_rank = rank(frac_increase)) %>%
+  mutate(colour = ifelse(abs(pos-exon_snp)<region_length, "gold", 
+                         ifelse(abs(pos-intron_snp)<region_length, "purple", "grey50")))
+
+peak_df$upper_bound <- mapply(find_ratio_bound, 
+                        peak_df$peak_mean_R,
+                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
+                        peak_df$peak_mean_H,
+                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
+                        0.5*(1+confidence_interval))
+
+peak_df$lower_bound <- mapply(find_ratio_bound, 
+                        peak_df$peak_mean_R,
+                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
+                        peak_df$peak_mean_H,
+                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
+                        0.5*(1-confidence_interval))
+
+ggplot(peak_df, aes(x = peak_rank, y = 100*frac_increase, fill=colour)) +
+  geom_bar(stat="identity") +
+  scale_fill_identity() +
+  theme_minimal() +
+  ggpubr::theme_pubr()+
+  ylab("% change in 2x Risk") +
+  xlab("Rank") +
+  geom_errorbar(aes(ymin = 100*lower_bound, ymax = 100*upper_bound, x = peak_rank), 
+                width = 0.3)
+
+```
 

### 2021-03-14 oscar_figures.Rmd
add some comments

@@
 ## Ribosome profiling quality control
+
+TODO: convert all libraries to p_load.
 
@@
 Read in the processed data (following alignment with Bowtie2, and filtering for high-confidence alignments), and generate
-a plot of the overall distribution of crosslinks, and the rolling average difference between the 2x Risk and 2x Healthy
+a plot of the overall distribution of crosslinks, and the rolling average difference between the 2x Risk and 2x Healthy.
+
+Next, filter for peaks in the iCLIP signal, and see how they change in 2x Risk versus 2x Healthy. Highlight those which are close to the SNPs.
 
 TODO: add script which does this filtering
+TODO: add fasta files for the minigenes
 
@@
 healthy_colour <- "#F8766D"
+confidence_interval <- 0.75  # fracion, i.e. 0.75 = 75%.
+peak_threshold <- 5 # how many times above the mean to be a "peak"
+region_length <- 50
 
@@
 rolling_window <- 20
+peak_t <- peak_threshold*1000/mg_l  # by definition...
 
@@
   xlim(1,1600) +
-  #ylim(-1,5) +
   theme_minimal() +
@@
 combined
-```
-
-Next, filter for peaks in the iCLIP signal, and see how they change in 2x Risk versus 2x Healthy. Highlight those which are close to the SNPs.
-
-```{r message=FALSE, warning=FALSE}
 
@@
 
-## Plot big changes
-
-confidence_interval <- 0.75  # fracion, i.e. 0.75 = 75%.
-
-peak_threshold <- 5 # how many times above the mean to be a "peak"
-region_length <- 50
-
-peak_t <- peak_threshold*1000/mg_l
-
 peak_df <- filtered_df %>%
@@
   group_by(pos) %>%
-  mutate(max_at_pos = max(normalised_counts)) %>%
+  mutate(max_at_pos = max(normalised_counts)) %>%  # at least 1 of the 4 samples above the threshold
   ungroup() %>%
-  filter(max_at_pos >= peak_t) %>%
+  filter(max_at_pos >= peak_t) %>%  # at least 1 of the 4 samples above the threshold
   group_by(condition, pos) %>%
@@
 ```
-
-##

### 2021-03-14 preprocessing_scripts/Targeted_RNA_seq/specific_RT/analyse_unc_trnaseq_final.R
Add processing script for specific RT

@@ -1,0 +1,225 @@
+library(tidyverse)
+library(Rsamtools)
+library(DescTools)
+library(clustringr)
+library(stringdist)
+
+# The purpose of this script is to find the number of unique UMIs  (i.e. unique cDNAs) in the 
+# targeted RNA sequencing of UNC13A, using a specific RT primer.
+#
+# This script takes .bam files aligned to the human (HG38) genome. It filters for alignments near
+# the UNC13A cryptic exon. It works out how many unique UMIs there are, and whether these 
+# cDNAs contain the risk or non-risk alleles of rs12973192
+#
+# The script first identifies whether reads contain the cryptic exon, or are "correctly" spliced. 
+# It then extracts the UMIs associated with each read from the read name (these were already moved
+# from the read header during demultiplexing with Ultraplex)
+#
+# Because these libraries were very low complexity/over-amplified (especially for the reads which contained the 
+# cryptic exon) there is a high risk of erroneous false-positive UMIs due to sequencing/PCR errors.
+# To minimise this, the code first ranks UMIs by their abundance (because the most abundant UMIs
+# are unlikely to be derived from sequencing errors). It then works down the list of UMIs, 
+# ensuring that each one is sufficiently different from the UMIs which have been added to the 
+# list of "true" UMIs. It does this by finding the minimum string distance of each UMI to 
+# the UMIs currently in the list of "true" UMIs. It requires that for a UMI sequence which 
+# has a string distance of n bases, it is at least 1/(10^n) times the abundance of 
+# the most similar UMI. It also requires that each UMI is at least 1/100 times as abundant
+# as the most abundant, and is present in at least 50 reads.
+
+get_sample <- function(string){
+    # Detects sample name from the file name
+    return(str_split(string,":")[[1]][2])
+}
+
+find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100,
+                             min_n = 50){
+    # df should have two columns - "strings" which are the UMIs, and "n" which
+    # is the copy number of each
+    #
+    # higher values of distance multiplier increase ability to detect low abundance
[…]

### 2021-03-14 preprocessing_scripts/Targeted_RNA_seq/random_hexamer/analyse_unc_trnaseq_3.R
add preprocessing script for random hexamer targeted rnaseq

@@ -5,15 +5,14 @@ library(clustringr)
 library(stringdist)
 
 # The purpose of this script is to find the number of unique UMIs  (i.e. unique cDNAs) in the 
-# targeted RNA sequencing of UNC13A, using a specific RT primer.
+# targeted RNA sequencing of UNC13A, using a random hexamer RT primer.
 #
 # This script takes .bam files aligned to the human (HG38) genome. It filters for alignments near
 # the UNC13A cryptic exon. It works out how many unique UMIs there are, and whether these 
 # cDNAs contain the risk or non-risk alleles of rs12973192
 #
-# The script first identifies whether reads contain the cryptic exon, or are "correctly" spliced. 
-# It then extracts the UMIs associated with each read from the read name (these were already moved
-# from the read header during demultiplexing with Ultraplex)
+# The script extracts the UMIs associated with each read from the read name (these were already moved
+# from the read header during demultiplexing with Ultraplex), and checks each read aligns to the cryptic.
 #
 # Because these libraries were very low complexity/over-amplified (especially for the reads which contained the 
 # cryptic exon) there is a high risk of erroneous false-positive UMIs due to sequencing/PCR errors.
@@ -26,53 +25,69 @@ library(stringdist)
 # the most similar UMI. It also requires that each UMI is at least 0.01x times as abundant
 # as the most abundant, and is present in at least 50 reads.
 #
-# This script differs from the random hexamer script in that it detects whether reads are cryptic
-# or correctly spliced by looking at the CIGAR string.
+# This script differs from the specific RT script in that it does not detect correctly spliced transcripts.
+# It also doesn't have a lower limit on the number of reads containing each UMI (though 
+# each must still be at least 0.01x the abundance of the most abundant).
+#
+# Note that although the forward reads in the fastq files for these samples correspond to the forward
+# strand of the RNA, and thus the MINUS strand of the genome, the bam file format automatically reverse
+# complements the reads, thus "C" is still healthy, and "G" is still risk.
+
 
 get_sample <- function(string){
-    # Detects sample name from the file name
     return(str_split(string,":")[[1]][2])
 }
 
[…]

### 2021-03-14 preprocessing_scripts/Targeted_RNA_seq/random_hexamer/analyse_unc_trnaseq_3.R
add comments

@@ -68,10 +68,8 @@ find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100){
 
         this_n <- df_remaining$n[which.max(df_remaining$n)]
 
-
         # find its distance from others
         if(nrow(df_added) > 0){
-
 
             df_added <- df_added %>%
                 mutate(distances = stringdist::stringdist(strings, largest))
@@ -79,12 +77,6 @@ find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100){
 
             # find how many there are of these
             already_added_sum <- sum(df_added$n[which(df_added$distances==min_distance)]) # PERHAPS THIS SHOULD BE ALL OF THEM? AND MAYBE SHOULD BE MAX
-
-            # this_df <- df %>%
-            #   mutate(distances == stringdist(strings, largest)) %>%
-            #   filter(!distances == 0) %>% # remove self
-            #   filter(distances == min(distances))
-            # already_added_sum <- sum(this_df$n) # in this case not "already added", instead all of them
 
         } else {
             min_distance <- 8 # set to large so defo included
@@ -117,8 +109,6 @@ find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100){
         return(df_added)
     }
 }
-
-
 
 ##### RUN
 
@@ -231,9 +221,8 @@ just_C_or_G_tmp[is.na(just_C_or_G_tmp)] <- 0
 just_C_or_G <- just_C_or_G_tmp %>%
     group_by(sample_name) %>%
     mutate(total = sum(n)) %>%
-    mutate(n_healthy = ifelse(nt=="C", n, 0)) %>%
+    mutate(n_healthy = ifelse(nt=="C", n, 0)) %>% # C is healthy in forward genomic strand
     mutate(p_value = pbinom(max(n_healthy), total, 0.5))
 
 # Write final data out
-
 write_csv(just_C_or_G, "values_for_plot.csv")

### 2021-03-14 oscar_figures.Rmd
removed pointless line of code

@@
   mutate(p_value = pbinom(combined_healthy, sum(combined_n), 0.5)) %>%  # single-tailed bionomial test
-  mutate(colour = ifelse(risk == "Non-Risk", "lightblue", "red")) %>%
   ungroup() %>%

### 2021-03-16 oscar_figures.Rmd
add heptamer analysis

@@
 ggplot(ribo_df, aes(x=distance_from_start, y=n_this_dist)) +
-    geom_bar(stat="identity") +
+    geom_line() +
+    geom_point() +
     xlim(-50, 50) +
@@
 ```
+
+## Heptameric analysis
+
+Analyse E-values (higher E value = stronger binding enrichment) for the motifs present at the exon SNP.
+
+```{r message=FALSE, warning=FALSE}
+
+# Read in supplementary 4 from Ray et al 2013
+
+data <- read_tsv(paste0(here(), "/data/heptamers/41586_2013_BFnature12311_MOESM334_ESM.txt"))
+
+for(type in c("exon", "intron")){
+
+    if(type == "exon"){
+        snp <-     "TAAAAGCATGGAT"
+        healthy <- "TAAAAGGATGGAT"
+    } else {
+        snp <-     "GGATGGGTGGATA"
+        healthy <- "GGATGGTTGGATA"
+    }
+
+    all_healthy <- c()
+    all_risk <- c()
+
+    # Make dataframe containing E scores for each relevant 7mer
+
+    for(i in 1:(str_length(snp)-6)){
+        this_snp <- str_sub(snp, i, i+6)
+        this_healthy <- str_sub(healthy, i, i+6)
+
+        all_healthy <- c(all_healthy, str_replace_all(this_healthy, "T", "U"))
+        all_risk <- c(all_risk, str_replace_all(this_snp, "T", "U"))
+
+        row1 <- data %>% filter(`7mer` == str_replace_all(this_snp, "T", "U")) %>%
+            mutate(snp = T, pos = i)
+
+        row2 <- data %>% filter(`7mer` == str_replace_all(this_healthy, "T", "U")) %>%
+            mutate(snp = F, pos = i)
+
+        if(i == 1){
+            full <- bind_rows(row1, row2)
+        } else {
+            full <- bind_rows(full, row1, row2)
+        }
+    }
+
+    # Find average change in E score for the two groups of heptamers
+
+    full2 <- full %>%
+        group_by(pos) %>%
+        filter(n() == 2) %>%
+        ungroup() %>%
+        pivot_longer(cols = contains("set")) %>%
+        select(-`7mer`) %>%
+        pivot_wider(names_from = "snp", values_from = "value") %>%
+        mutate(abs_change = `TRUE`-`FALSE`) %>%
+        filter(!is.na(`TRUE`) & !is.na(`FALSE`)) %>%
+        group_by(name) %>%
+        mutate(mean_abs_change = mean(abs_change)) %>%
+        select(name, mean_abs_change) %>%
+        unique() %>%
+        arrange(mean_abs_change) %>%
+        mutate(is_tdp43 = ifelse(str_detect(name, "RNCMPT00076"), "red", "grey80")) %>%
+        mutate(short_name = str_sub(name, 1, 11)) %>%
+        ungroup() %>% group_by(short_name) %>%
+        mutate(average_delta = mean(mean_abs_change)) %>%
+        select(short_name, is_tdp43, average_delta) %>%
+        unique()
+
+    # Plot ranks
+
+    plot <- ggplot(full2, aes(x=rank(average_delta), y = average_delta,
+                      fill = is_tdp43)) +
+        geom_bar(stat="identity") +
+        scale_fill_identity() +
+        theme_minimal() +
+        ggtitle(paste0("Average change in binding E value for each dataset with SNP for ", type, " SNP")) +
+        xlab("") +
+        ylab("Average change in E score") +
+        ggpubr::theme_pubr() +
+        xlab("Rank")
+
+    print(plot)
+
+    ggsave(paste0("C:/Users/ogw/Google Drive/UCL PhD/Year 2/unc13a/figures for paper/heptamter ranking ",
+                  type, ".svg"), plot = plot)
+
+    full_just_tdp <- full %>%
+        group_by(pos) %>%
+        filter(n() == 2) %>%
+        ungroup() %>%
+        select(`7mer`, contains("RNCMPT00076"), snp, pos) %>%
+        mutate(mean_E = (RNCMPT00076_e_setA + RNCMPT00076_e_setB)/2) %>%
+        pivot_longer(cols = contains("RNCMP")) %>%
+        mutate(ypos = ifelse(snp, -0.35, -0.27))
+
+    plot2 <- ggplot(full_just_tdp, aes(x = pos, y = value, colour = snp,
+                              label = `7mer`)) +
+        geom_point(size = 3) +
+        xlab("7mer") +
+        ylab("E value (similar to rank?)") +
+        ggeasy::easy_add_legend_title("Contains Risk SNP") +
+        ggtitle(paste0("Binding E value for TDP43 for all 7x 7mers with or without risk SNP for ", type, " SNP")) +
+        geom_label(aes(y = ypos), size = 2.7) +
+        xlim(0.5, 7.5) +
+        theme_bw() + 
+        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  +
+        ggeasy::easy_remove_x_axis()
+
+    print(plot2)
+
+}
+
+```

### 2021-03-18 paper_figures.Rmd
updated tables and script
added two samples for each condition to gain significance - added statistical analysis on script and figures

--- 2021-03-03 paper_figures.Rmd
@@
 title: "Figures for Paper"
-author: "Anna-Leigh Brown + Matteo Zanovello"
+author: "Anna-Leigh Brown"
 date: "06/10/2020"
@@
 
-for_bar_plot_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
-
-for_bar_plot_sy <- data.frame(for_bar_plot_sy)
-for_bar_plot_sy$condition <- as.factor(for_bar_plot_sy$condition)
-for_bar_plot_sy$condition <- factor(for_bar_plot_sy$condition,levels(for_bar_plot_sy$condition)[c(2,1)])
-
-nmd_plot = ggbarplot(for_bar_plot_sy,
+fbp_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
+fbp_sy$condition <- as.factor(fbp_sy$condition)
+fbp_sy$condition <- factor(fbp_sy$condition,levels(fbp_sy$condition)[c(2,1)])
+
+stattest_sy <- fbp_sy %>%
+    group_by(gene) %>%
+    t_test(value ~ 1, mu = 1) %>%
+    add_significance() %>%
+    add_xy_position(x="gene")
+stattest_sy
+
+nmd_plot_sy <- ggbarplot(fbp_sy,
                      x = 'gene',
+                     y = 'value',
                      add = c("mean_se","jitter"),
-                        y = 'value',
                      fill = 'condition',
@@
     ggpubr::theme_pubr() + 
-  scale_fill_manual(
-    values = c("#40B0A6","#E1BE6A")
-  ) + 
-  scale_color_manual(
-    values = c("#1C2617","#262114")
-  )
-
-plot(nmd_plot)
+    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) + 
+    scale_color_manual(values = c("#1C2617","#262114")) +
+    ylab("Percent of transcript after CHX treatment") +
+    xlab("") +
+    stat_pvalue_manual(stattest_sy, label="p.signif", y.position = 9)
+nmd_plot_sy
+
+```
+
+```{r}
+
+fbp_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
+fbp_dz$condition <- as.factor(fbp_dz$condition)
+fbp_dz$condition <- factor(fbp_dz$condition,levels(fbp_dz$condition)[c(2,1)])
+
+stattest_dz <- fbp_dz %>%
+    group_by(gene) %>%
+    t_test(value ~ 1, mu = 1) %>%
+    add_significance() %>%
+    add_xy_position(x="gene")
+stattest_dz
+
+nmd_plot_dz <- ggbarplot(fbp_dz,
+                        x = 'gene',
+                        y = 'value',
+                        add = c("mean_se","jitter"),
+                        fill = 'condition',
+                        color = 'condition',
+                        position = position_dodge(0.8),
+                        dot.size = 10) +
+    scale_y_continuous(labels = scales::percent) + 
+    ggpubr::theme_pubr() + 
+    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) + 
+    scale_color_manual(values = c("#1C2617","#262114")) +
+    ylab("Percent of transcript after CHX treatment") +
+    xlab("") +
+    stat_pvalue_manual(stattest_dz, label="p.signif", y.position = 26)
+nmd_plot_dz
+
 ```

### 2021-03-18 oscar_figures.Rmd
add itc bar graph

@@
 knitr::opts_chunk$set(echo = TRUE)
-library(tidyverse)
-library(DESeq2)
-library(here)
-library(ggrepel)
+if (!require("pacman")) install.packages("pacman")
+library(pacman)
+p_load("tidyverse", "DESeq2", "here", "ggrepel", "patchwork")
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-library(patchwork)
-
 exon_snp <- 556 
@@
 ```
+
+## Isothermal Titration Calorimetry
+
+Read in values
+
+```{r warning=FALSE, message=FALSE}
+df <- read_csv(paste0(here(), 
+                      "/data/itc/itc final values.csv")) %>%
+    group_by(RNA) %>%
+    mutate(sd =sd(kd)) %>%
+    mutate(mean_kd = mean(kd))
+
+ggplot(df, aes(x = RNA, y = kd)) +
+    geom_bar(aes(x = RNA, y = mean_kd), stat="identity", position="dodge") +
+    geom_point() +
+    geom_errorbar(aes(x=RNA, ymin = mean_kd-sd, ymax = mean_kd+sd),
+                  width = 0.3) +
+    ylim(0, NA) +
+    ggpubr::theme_pubr() +
+    ylab("Kd / nM") +
+    xlab("")
+
+t.test(df$kd[which(df$RNA=="H")], df$kd[which(df$RNA=="R")])
+```

### 2021-03-18 oscar_figures.Rmd
Add supplementary iCLIP figure

@@
 
+Supplementary iCLIP Figures
+
+```{r message=FALSE, warning=FALSE}
+
+bars <- filtered_df %>%
+  ungroup() %>%
+  group_by(condition, pos) %>%
+  mutate(mean_normalised = mean(normalised_counts),
+         sd_normalised = sd(normalised_counts)) %>%
+  mutate(colour = ifelse(condition == "H", healthy_colour, risk_colour))
+
+bar_top <- ggplot(bars %>% filter(abs(pos-exon_snp) < 35), aes(x = pos, y = mean_normalised,
+                                                       fill = colour)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_fill_identity() +
+  geom_errorbar(aes(ymin = mean_normalised-sd_normalised, ymax=mean_normalised+sd_normalised),
+                position="dodge") +
+  geom_vline(xintercept = exon_snp, linetype="dashed") +
+  ylab("XLs per 1000") +
+  ggpubr::theme_pubr() +
+  ggtitle("iCLIP signal around CE SNP") +
+  xlab("")
+
+bar_bottom <- ggplot(bars %>% filter(abs(pos-intron_snp) < 35), aes(x = pos, y = mean_normalised,
+                                                       fill = colour)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_fill_identity() +
+  geom_errorbar(aes(ymin = mean_normalised-sd_normalised, ymax=mean_normalised+sd_normalised),
+                position="dodge") +
+  geom_vline(xintercept = intron_snp, linetype="dashed") +
+  ggpubr::theme_pubr() +
+  ylab("XLs per 1000") +
+  ggtitle("iCLIP signal around intronic SNP") +
+  xlab("")
+
+bar_combined <- bar_top / bar_bottom
+bar_combined
+
+```
+
 ## Heptameric analysis

### 2021-03-19 paper_figures.Rmd
Update paper_figures.Rmd
fix git merge flags

--- 2021-03-09 paper_figures.Rmd
@@
 
-## barplot UNC13B NMD PSI
+## barplot UNC13B frameshift PSI
 
@@
           axis.text.y = element_text(size = 28))
+
 ```
@@
           axis.text.y = element_text(size = 28)) +
-    scale_y_continuous(labels = scales::percent) +
+    scale_y_continuous(labels = scales::percent,expand = expansion(mult = c(0, .1))) +
     geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
@@
 
-## Scatterplot UNC13B-IR and UNC13A Crptic
-
-```{r}
+## Scatterplot UNC13B-IR and UNC13A Cryptic
+
+```{r}
+
 rel_rna_cryptic_amount %>% 
@@
 nmd_plot = ggbarplot(for_bar_plot_sy,
+                        x = 'condition',
+                        add = c("mean_se","jitter"),
+                        y = 'value',
+                        fill = 'condition',
+                        color = 'condition',
+                        position = position_dodge(0.8),
+                        dot.size = 10,facet.by = 'gene') + 
+  scale_y_continuous(labels = scales::percent) + 
+  ggpubr::theme_pubr() +
+  scale_fill_manual(
+    values = c("#40B0A6","#E1BE6A")
+  ) + 
+  scale_color_manual(
+    values = c("#1C2617","#262114")
+  ) +
+  ylab("Percent of transcript after CHX treatment") + 
+  stat_compare_means()
+
+plot(nmd_plot)
+```
+
+```{r}
+
+for_bar_plot_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
+
+for_bar_plot_dz <- data.frame(for_bar_plot_dz)
+for_bar_plot_dz$condition <- as.factor(for_bar_plot_dz$condition)
+for_bar_plot_dz$condition <- factor(for_bar_plot_dz$condition,levels(for_bar_plot_dz$condition)[c(2,1)])
+
+nmd_plot2 = ggbarplot(for_bar_plot_dz,
                         x = 'gene',
@@
     values = c("#1C2617","#262114")
-  ) +
-  ylab("Percent of transcript after CHX treatment")
-
-plot(nmd_plot)
+  )
+
+plot(nmd_plot2)
 ```
@@
     mutate(number_g_alleles = as.numeric(rs12973192) - 1) %>% 
-    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi))
+    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
+    mutate(junction_reads_stmn2 = STMN2_annotated_leaf + STMN2_cryptic_leaf) %>% 
+  mutate(junction_reads_unc13a = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf + UNC13A_annotated_leaf) %>% 
+  unique()
 
@@
 ```{r}
-comps = unique(clean_data_table$disease_group2) %>% gsub("Control", " Control",.) %>% 
+comps = clean_data_table %>% 
+  group_by(disease_group2,tissue_clean) %>% 
+  mutate(n_sample = n_distinct(sample)) %>% 
+  ungroup() %>% 
+  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
+  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
+  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% pull(detection_name) %>% 
+  unique() %>% 
   combn(.,2,simplify = F)
@@
   facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
-  theme(legend.position = 'none')
+  theme(legend.position = 'none') + 
+  stat_compare_means(comparisons = comps)
+```
+
+```{r make_those_plots_function}
+plot_by_tissue = function(tissue_name,dt){
+
+only_tissue = dt %>% 
+  group_by(disease_group2,tissue_clean) %>% 
+  mutate(n_sample = n_distinct(sample)) %>% 
+  ungroup() %>% 
+  filter(tissue_clean == tissue_name) %>% 
+  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
+  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
+  mutate(tissue_clean = gsub("_"," ",tissue_clean))
+
+unique_x_values = only_tissue %>% 
+  pull(detection_name) %>% 
+  unique()
+
+comparisons = combn(unique_x_values,2,simplify = F)
+
+the_plot = only_tissue %>% 
+    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
+    ggpubr::theme_pubr() +
+    geom_boxplot(aes(fill = disease_group2)) + 
+    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
+    ylab("UNC13A TPM") + 
+    xlab("") + 
+    guides(color = FALSE) +
+    theme(text = element_text(size = 20)) +
+  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
+  theme(legend.position = 'none') + 
+  stat_compare_means() + 
+  stat_compare_means(label = "p.signif",comparisons = comparisons,)
+
+  return(the_plot)
+}
+
+dis_tis = clean_data_table[disease_tissue == T,unique(tissue_clean)]
+dis_tis = c("Cerebellum",dis_tis)
+my_plots = purrr::map(dis_tis,plot_by_tissue, clean_data_table)
+ggarrange(plotlist=my_plots)
+
 ```
@@
 clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(UNC13A_annotated_leaf > 15) %>% 
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
@@
   geom_point() + 
-  stat_cor(label.y = 38,size = 8,method = 's',cor.coef.name = 'rho') + 
+  stat_cor(label.y = 20,size = 8,method = 's',cor.coef.name = 'rho') + 
   geom_smooth(method = lm, se = FALSE,color = "black") +
@@
   filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(UNC13A_annotated_leaf > 15) %>% 
   filter(disease_tissue == T) %>% 
@@
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(junction_reads_stmn2 >= 30) %>% 
+  filter(junction_reads_unc13a >= 30) %>%
+  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>% 
+  # filter(STMN2_cryptic_leaf >= 2) %>% 
   filter(grepl("Cortex",tissue_clean)) %>% 
@@
   theme(text = element_text(size = 18)) +
-  scale_y_continuous(labels = scales::percent)  +
-  scale_x_continuous(labels = scales::percent) 
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
+  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))  
 ```
@@
 clean_data_table %>% 
-  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-  # filter(unc13a_cryptic_leaf_psi > 0) %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
+    filter(tissue_clean %in% 
+               c("Frontal Cortex", 
+                 "Lumbar Spinal Cord", 
+                 "Cervical Spinal Cord",
+                 "Motor Cortex", 
+                 "Temporal Cortex",
+                 "Cerebellum") )%>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 1) %>% 
+  filter(STMN2_cryptic_leaf >= 1) %>% 
   dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
   melt() %>% 
-  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
     mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
@@
   ggplot(aes(x = variable, y = value)) +
-  geom_boxplot(aes(color = variable)) +
+  geom_boxplot(aes(color = variable),outlier.shape = NA) +
   geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
@@
   scale_color_manual(values = c("#008574","#cd5582")) +
-  facet_wrap(~tissue_clean,nrow = 3,scales = 'free') +
-    scale_y_continuous(labels = scales::percent)  +
+  facet_wrap(~tissue_clean,nrow = 3, scales = 'free') +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
   ggpubr::theme_pubr() +
@@
   xlab("") +
-  ylab("PSI")  +
+  ylab("CE PSI")  +
   theme(legend.position = 'none')
@@
 # Effect of UNC13A genotype
-## UNC13A psi by genotype - rs12973192
+
+## UNC13A psi by genotype - rs12973192 - RE
+
 ```{r}
 rs12973192_comparisons = clean_data_table %>% 
+  filter(junction_reads_unc13a >=30) %>% 
   filter(disease_tissue == T) %>% 
@@
   ungroup() %>% 
-  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
-  pull(detection_name) %>% 
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )"))
+
+comp_for_gg = rs12973192_comparisons %>% pull(detection_name) %>% 
   unique() %>% 
@@
 
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
-  group_by(rs12973192) %>% 
-  mutate(n_sample = length(unique(sample))) %>% 
-  ungroup() %>% 
-  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
+ce_psi = rs12973192_comparisons %>% 
   ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12973192)) + 
@@
   ylab("UNC13A PSI") +
-  xlab("UNC13A rs12973192 Genotype") +
+  xlab(element_blank()) +
   scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
@@
   stat_compare_means() +
-  stat_compare_means(comparisons = rs12973192_comparisons,label = "p.signif") + 
-  scale_y_continuous(trans='log10')
-```
-
-## UNC13A psi by genotype - rs12608932
-```{r}
-
+  theme(legend.position = "none") +
+  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") + 
+  scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
+  ggtitle("CE SNP (rs12973192)")
+
+```
+
+## UNC13A psi by genotype - rs12608932 - RI
+```{r}
 rs12608932_comparisons = clean_data_table %>% 
+  filter(junction_reads_unc13a >=30) %>% 
   filter(disease_tissue == T) %>% 
@@
   ungroup() %>% 
-  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
-  pull(detection_name) %>% 
+  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))
+
+comp_for_gg = rs12608932_comparisons %>% pull(detection_name) %>% 
   unique() %>% 
@@
 
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
-  group_by(rs12608932) %>% 
-  mutate(n_sample = length(unique(sample))) %>% 
-  ungroup() %>% 
-  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
+in_psi = rs12608932_comparisons %>% 
   ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12608932)) + 
@@
   geom_jitter( height = 0) + 
-  ylab("UNC13A PSI") +
-  xlab("UNC13A rs12608932 Genotype") +
+  ylab(element_blank()) +
+  xlab(element_blank()) +
   scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
   ggpubr::theme_pubr() +
+  theme(legend.position = "none") +
   theme(text = element_text(size = 24)) +
   stat_compare_means() +
-  stat_compare_means(comparisons = rs12608932_comparisons,label = "p.signif") + 
-  scale_y_continuous(trans='log10')
-
+  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") + 
+    scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
+  ggtitle("intronic SNP (rs12973192)")
+```
+
+```{r}
+ggpubr::ggarrange(ce_psi,in_psi)
 ```
@@
 
-## Correlation UNC13A PSI and STMN2 PSI with genotype effect
-
-```{r}
-
-clean_data_table %>% 
-  filter(disease_tissue == T) %>% 
+## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12973192
+
+```{r}
+
+ce_snp = clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+  filter(junction_reads_stmn2 >= 30) %>% 
+  filter(junction_reads_unc13a >= 30) %>%
   filter(grepl("Cortex",tissue_clean)) %>% 
@@
   theme(legend.title = element_blank()) +
-  theme(text = element_text(size = 24),legend.text = element_text(size = 32))
-
+  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
+    ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
+      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
+  ggtitle("CE SNP (rs12973192) genotype")
+
+```
+
+## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12608932
+
+```{r}
+
+in_snp = clean_data_table %>% 
+  filter(disease_tissue == T) %>% 
+    filter(junction_reads_stmn2 >= 30) %>% 
+  filter(junction_reads_unc13a >= 30) %>%
+  filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+    mutate(rs12608932 = fct_relevel(rs12608932,
+                              "A/A", "A/C", "C/C")) %>%
+  filter(!grepl("Cord",tissue_clean)) %>% 
+  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
+    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
+    ggpubr::ggscatter(., 
+                      x = "log10_stmn2",
+                      y = "log10_unc13a",
+                      color = 'rs12608932',                      
+                      add = "reg.line", size = 3
+                      ) +
+    ylab("UNC13A CE PSI ") + 
+    xlab("STMN2 Cryptic PSI ") +
+  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
+                      y = unc13a_cryptic_leaf_psi,
+                      color = rs12608932),
+           method = 'spearman',
+           cor.coef.name = 'rho',
+           show.legend = FALSE,
+           size = 14) + 
+    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  theme(legend.title = element_blank()) +
+  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
+      ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
+      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
+   ggtitle("intronic SNP (rs12608932) genotype")
+
+```
+
+```{r}
+ggpubr::ggarrange(ce_snp,in_snp)
 ```
@@
   filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+  filter(junction_reads_stmn2 >= 30) %>% 
+  filter(junction_reads_unc13a >= 30) %>% 
   filter(grepl("Cortex",tissue_clean)) %>% 
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  group_by(rs12973192) %>% 
+  mutate(n_sample = length(unique(sample))) %>% 
+  ungroup() %>% 
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
+  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))
+
+comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)
+
+detected_table_for_graph %>% 
+  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
+  geom_boxplot(show.legend = F) + 
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
+  stat_compare_means() + 
+  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
+  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
+  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 0,size = 1.5) 
+```
+
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord
+
+```{r}
+detected_table_for_graph = clean_data_table %>% 
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  filter(junction_reads_stmn2 >= 30) %>% 
+  filter(junction_reads_unc13a >= 30) %>% 
   filter(disease_tissue == T) %>% 
@@
 
-## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord
-
-```{r}
-detected_table_for_graph = clean_data_table %>% 
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932
+
+```{r}
+detected_table_for_graph_rs12608932 = clean_data_table %>% 
+  filter(grepl("Cortex",tissue_clean)) %>% 
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
@@
   filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
-  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
-  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
-  group_by(rs12973192) %>% 
-  mutate(n_sample = length(unique(sample))) %>% 
-  ungroup() %>% 
-  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
-  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))
-
-comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)
-
-detected_table_for_graph %>% 
-  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
-  geom_boxplot(show.legend = F) + 
-  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
-  stat_compare_means( label = "p.format") + 
-  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
-  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
-  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
-  xlab("UNC13A rs12973192 Genotype") +
-  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
-  ggpubr::theme_pubr() +
-  theme(text = element_text(size = 24)) +
-  geom_hline(yintercept = 0,size = 1.5) 
-```
-
-## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932
-
-```{r}
-detected_table_for_graph_rs12608932 = clean_data_table %>% 
-  filter(grepl("Cortex",tissue_clean)) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
-  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
-  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
+  filter(junction_reads_stmn2 >= 30) %>% 
+  filter(junction_reads_unc13a >= 30) %>% 
   filter(grepl("Cortex",tissue_clean)) %>% 
@@
    mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2))
+```
+
+```{r}
 liu_vafs %>% 
@@
   stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
-
+```
+
+```{r}
+liu_stmn2 %>% 
+  dplyr::select(stmn2_psi,sample,unc13a_psi) %>% 
+  data.table::melt() %>% 
+   mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
+    filter(!grepl("unsorted",sample_name)) %>% 
+    mutate(condition = sub(".*_", "", sample)) %>% 
+    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>% 
+    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>% 
+  ggbarplot(,
+            x = "condition",
+            add = c("mean_se",'jitter'),
+            y = "value",
+            fill = 'variable',
+            color = 'variable',
+            position = position_dodge(0.8),dot.size = 6) + 
+  ggpubr::theme_pubr() +
+  scale_color_manual(values = c("#000000","#000000")) +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
+  theme(text = element_text(size = 24))  +
+  scale_y_continuous(labels = scales::percent) + 
+  xlab("") +
+  ylab("PSI")  +
+  theme(legend.position =  'top') +
+  theme(legend.title=element_blank()) +
+  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
+```
+
+```{r}
 liu_stmn2 %>% 

### 2021-03-23 paper_figures.Rmd
Update paper_figures.Rmd
Ignore the person who is not hom for both snps

@@
                fill ="red")) + 
-    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA)) + 
+    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA) + 
     geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
@@
 detected_table_for_graph = clean_data_table %>% 
+  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>% 
   filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
@@
   mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
-  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))
+  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3)) %>% as.data.table()
+
+#get out that person who is discongruous
 

### 2021-03-24 paper_figures.Rmd

@@
 
-for_bar_plot_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
-
-for_bar_plot_sy <- data.frame(for_bar_plot_sy)
-for_bar_plot_sy$condition <- as.factor(for_bar_plot_sy$condition)
-for_bar_plot_sy$condition <- factor(for_bar_plot_sy$condition,levels(for_bar_plot_sy$condition)[c(2,1)])
-
-nmd_plot = ggbarplot(for_bar_plot_sy,
-                        x = 'condition',
+fbp_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
+fbp_sy$condition <- as.factor(fbp_sy$condition)
+fbp_sy$condition <- factor(fbp_sy$condition,levels(fbp_sy$condition)[c(2,1)])
+
+stattest_sy <- fbp_sy %>%
+    group_by(gene) %>%
+    t_test(value ~ 1, mu = 1) %>%
+    add_significance() %>%
+    add_xy_position(x="gene")
+stattest_sy
+
+nmd_plot_sy <- ggbarplot(fbp_sy,
+                     x = 'gene',
+                     y = 'value',
                      add = c("mean_se","jitter"),
-                        y = 'value',
-                        fill = 'condition',
-                        color = 'condition',
-                        position = position_dodge(0.8),
-                        dot.size = 10,facet.by = 'gene') + 
-  scale_y_continuous(labels = scales::percent) + 
-  ggpubr::theme_pubr() +
-  scale_fill_manual(
-    values = c("#40B0A6","#E1BE6A")
-  ) + 
-  scale_color_manual(
-    values = c("#1C2617","#262114")
-  ) +
-  ylab("Percent of transcript after CHX treatment") + 
-  stat_compare_means()
-
-plot(nmd_plot)
-```
-
-```{r}
-
-for_bar_plot_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
-
-for_bar_plot_dz <- data.frame(for_bar_plot_dz)
-for_bar_plot_dz$condition <- as.factor(for_bar_plot_dz$condition)
-for_bar_plot_dz$condition <- factor(for_bar_plot_dz$condition,levels(for_bar_plot_dz$condition)[c(2,1)])
-
-nmd_plot2 = ggbarplot(for_bar_plot_dz,
-                        x = 'gene',
-                        add = c("mean_se","jitter"),
-                        y = 'value',
                      fill = 'condition',
@@
     ggpubr::theme_pubr() +
-  scale_fill_manual(
-    values = c("#40B0A6","#E1BE6A")
-  ) + 
-  scale_color_manual(
-    values = c("#1C2617","#262114")
-  )
-
-plot(nmd_plot2)
+    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) +
+    scale_color_manual(values = c("#1C2617","#262114")) +
+    ylab("Percent of transcript after CHX treatment") +
+    xlab("") +
+    stat_pvalue_manual(stattest_sy, label="p.signif", y.position = 9)
+nmd_plot_sy
+
+```
+
+```{r}
+
+fbp_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
+fbp_dz$condition <- as.factor(fbp_dz$condition)
+fbp_dz$condition <- factor(fbp_dz$condition,levels(fbp_dz$condition)[c(2,1)])
+
+stattest_dz <- fbp_dz %>%
+    group_by(gene) %>%
+    t_test(value ~ 1, mu = 1) %>%
+    add_significance() %>%
+    add_xy_position(x="gene")
+stattest_dz
+
+nmd_plot_dz <- ggbarplot(fbp_dz,
+                        x = 'gene',
+                        y = 'value',
+                        add = c("mean_se","jitter"),
+                        fill = 'condition',
+                        color = 'condition',
+                        position = position_dodge(0.8),
+                        dot.size = 10) +
+    scale_y_continuous(labels = scales::percent) +
+    ggpubr::theme_pubr() +
+    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) +
+    scale_color_manual(values = c("#1C2617","#262114")) +
+    ylab("Percent of transcript after CHX treatment") +
+    xlab("") +
+    stat_pvalue_manual(stattest_dz, label="p.signif", y.position = 26)
+nmd_plot_dz
+
 ```

### 2021-03-25 paper_figures.Rmd
Add Oscar figs - add rstatix for Matteo

@@
 title: "Figures for Paper"
-author: "Anna-Leigh Brown + Matteo Zanovello"
+author: "Anna-Leigh Brown + Matteo Zanovello + Oscar Wilkins"
 date: "06/10/2020"
@@
        "ggthemes",
-       "ggrepel")
+       "ggrepel",
+       "DESeq2", 
+       "patchwork", 
+       "rstatix")
 ```
@@
 ```
+
+# Ribosome profiling
+## Ribosome profiling quality control
+
+TODO: convert all libraries to p_load.
+
+Perform quality control to check that reads are periodic.
+
+```{r message=FALSE, warning=FALSE}
+read_length_filter <- 29  # length of reads of interest
+gencode_v29_info <- read_tsv(paste0(here(), "/data/longest_proteincoding_transcript_hs_details.txt"))
+ribo_df <- read_tsv(paste0(here(), "/data/riboseq/sh_tdpb.Aligned.toTranscriptome.out.bam.bed.gz"), 
+                    col_names = c("transcript_id", "start0", "end", "strand")) %>%
+    inner_join(gencode_v29_info, by = "transcript_id") %>%
+    dplyr::filter(end-start0 == read_length_filter, strand == "+") %>%
+    mutate(distance_from_start = start0-cds_start) %>%
+    group_by(distance_from_start) %>%
+    mutate(n_this_dist = n()) %>%
+    dplyr::select(distance_from_start, n_this_dist) %>%
+    unique()
+
+```
+
+```{r, warning=FALSE}
+ggplot(ribo_df, aes(x=distance_from_start, y=n_this_dist)) +
+    geom_line() +
+    geom_point() +
+    xlim(-50, 50) +
+    ylab("Counts") + 
+    xlab("Distance of 5' end from annotated start codon") +
+    ggtitle(paste0("Periodicity of ", read_length_filter, "nt reads from TDP-43 KD replicate B"))
+```
+
+## Ribosome profiling volcano plot
+
+Use DESeq to detect differential ribosome footprints and create volcano plots
+
+```{r message = FALSE, warning=FALSE}
+counts_df <- read_csv(paste0(here(), "/data/riboseq/cds_feature_counts.csv"))
+counts <- as.matrix(counts_df %>% column_to_rownames("gene_id"))
+ipsc_splicing <- read_csv(paste0(here(), "/data/ipsc_splicing_results.csv"))
+ipsc_de = read_csv(paste0(here(), "/data/ipsc_differential_expression.csv"))
+
+col_data <- data.frame(sample = colnames(counts)) %>%
+  mutate(condition = ifelse(str_detect(sample, "tdp"), "tdp_ko", "control"))
+
+dds <- DESeqDataSetFromMatrix(countData = counts,
+                              colData = col_data,
+                              design= ~condition)
+dds <- DESeq(dds)
+results_name <- resultsNames(dds)[2]
+res <- results(dds, name=results_name)
+
+res_df <- data.frame(res) %>%
+  rownames_to_column("gene_id") %>%
+  left_join(gencode_v29_info, by = "gene_id")
+
+cryptic_df <- ipsc_splicing %>%
+  dplyr::select(gene_name, cryptic_junction) %>%
+  unique() %>%
+  filter(cryptic_junction)
+
+res_df2 <- res_df %>%
+  mutate(cryptic_junction = gene_name %in% cryptic_df$gene_name) %>%
+  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
+                                                     "UNC13B","PFKP","SETD5",
+                                                     "ATG4B","STMN2","TARDBP") ~ gene_name)) %>%
+  mutate(colour = ifelse(cryptic_junction, "#fe6101",  "#648FFF")) %>%
+  arrange(colour)
+```
+
+```{r, warning=FALSE} 
+p1 <- ggplot(res_df2, aes(x = log2FoldChange, y = -log10(pvalue), label = label_junction)) +
+  geom_point(data = res_df2 %>% filter(!cryptic_junction),
+             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#648FFF") +
+  geom_point(data = res_df2 %>% filter(cryptic_junction),
+             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#fe6101") +
+  ggpubr::theme_pubr() +
+  ylab("-Log10(p-value)") +
+  xlab("Log2 fold change") +
+  geom_text_repel(box.padding = 1.2, max.overlaps = Inf)
+
+p2 <- p1 + ylim(0,10)
+p1
+p2
+```
+
+# Targeted RNA-seq
+
+Read in data from processed bam files
+
+TODO: add scripts which deduplicate UMIs for each.
+
+```{r message=FALSE, warning=FALSE}
+specific_rt <- read_csv(paste0(here(), "/data/targeted_RNA_seq/specific_RT_primer/for_plot.csv")) %>%
+    select(risk, sample_name, n)
+
+random_hexamer <- read_csv(paste0(here(), "/data/targeted_RNA_seq/random_hexamer/values_for_plot.csv")) %>%
+    select(-sample_name) %>%
+    select(sample_name = actual_sample_name, risk, n)
+
+tg_rnaseq_df <- bind_rows(specific_rt, random_hexamer) %>%
+  group_by(sample_name) %>%
+  mutate(n_healthy = ifelse(risk == "Non-Risk", n, 0)) %>%
+  mutate(combined_healthy = sum(n_healthy)) %>%
+  select(-n_healthy) %>%
+  mutate(combined = sum(n)) %>%
+  group_by(sample_name, risk) %>%
+  mutate(combined_n = sum(n)) %>%
+  ungroup() %>%
+  select(-n) %>%
+  unique() %>%
+  group_by(sample_name) %>%
+  mutate(p_value = pbinom(combined_healthy, sum(combined_n), 0.5)) %>%  # single-tailed bionomial test
+  ungroup() %>%
+  mutate(sample_name_ordered = fct_reorder(sample_name, dplyr::desc(combined_n)))
+
+ggplot(tg_rnaseq_df %>%
+         filter(combined_n > 0), aes(x = sample_name_ordered, y = combined_n, fill = risk)) +
+  geom_bar(stat="identity", position="dodge") +
+  geom_text(aes(x=sample_name, y=-2.5, label=signif(p_value,2)), size = 3.5) +
+  xlab("") +
+  ylab("Unique cDNAs") +
+  ggeasy::easy_add_legend_title("Allele") +
+  theme_minimal() +
+  ggpubr::theme_pubr() +
+  ggeasy::easy_rotate_x_labels()
+```
+
+# iCLIP of minigenes
+
+Read in the processed data (following alignment with Bowtie2, and filtering for high-confidence alignments), and generate
+a plot of the overall distribution of crosslinks, and the rolling average difference between the 2x Risk and 2x Healthy.
+
+Next, filter for peaks in the iCLIP signal, and see how they change in 2x Risk versus 2x Healthy. Highlight those which are close to the SNPs.
+
+TODO: add script which does this filtering
+TODO: add fasta files for the minigenes
+
+```{r message=FALSE, warning=FALSE}
+exon_snp <- 556 
+intron_snp <- 1106
+intron_start <- 141 
+intron_end <- 1429
+acceptor1 <- 394
+acceptor2 <- 444
+donor <- 572
+mg_l <- 1624 
+cpoor_start <- 490
+cpoor_end <- 1325
+risk_colour <- "#00BFC4"
+healthy_colour <- "#F8766D"
+confidence_interval <- 0.75  # fracion, i.e. 0.75 = 75%.
+peak_threshold <- 5 # how many times above the mean to be a "peak"
+region_length <- 50
+
+filtered_df <- read_csv(paste0(here(), "/data/iCLIP/bowtie2_aligned_2_filtered_df.csv"))
+
+rolling_window <- 20
+peak_t <- peak_threshold*1000/mg_l  # by definition...
+
+df2 <- filtered_df %>%
+  ungroup() %>%
+  group_by(condition, pos) %>%
+  mutate(mean_counts = sum(normalised_counts)/2) %>%
+  select(pos, condition, mean_counts) %>%
+  unique() %>%
+  pivot_wider(names_from = "condition", values_from ="mean_counts")
+
+df2[is.na(df2)] <- 0
+
+df2 <- df2 %>% 
+  mutate(diff = R-H) %>%
+  mutate(mean_4 = (H+R)/2) %>%
+  arrange(pos)
+
+top <- ggplot(df2, aes(x=pos, y=diff)) +
+  geom_area(aes(x=pos, y = mean_4), colour="grey50", fill = "grey50", alpha=0.2) +
+  geom_vline(xintercept = exon_snp, linetype="dashed") +
+  geom_vline(xintercept = intron_snp, linetype="dashed") +
+  xlim(1,1600) +
+  theme_minimal() +
+  ylab("XLs per 1000") +
+  xlab("") +
+  geom_vline(xintercept = cpoor_start) +
+  geom_vline(xintercept = cpoor_end)
+
+bottom <- ggplot(df2, aes(x=pos, y=zoo::rollmean(diff, k = rolling_window, na.pad=T))) +
+  geom_area(colour = "grey50", fill="grey50", alpha = 1) +
+  geom_vline(xintercept =  exon_snp, linetype="dashed") +
+  geom_vline(xintercept =  intron_snp, linetype="dashed") +
+  xlim(1,1600) +
+  ylim(-0.5,0.75) +
+  theme_minimal() +
+  ylab("Change in XL density") +
+  xlab("Distance from start of minigene") +
+  geom_vline(xintercept =  intron_start) +
+  geom_vline(xintercept =  intron_end) +
+  geom_vline(xintercept =  acceptor1) +
+  geom_vline(xintercept =  acceptor2) +
+  geom_vline(xintercept =  donor)
+
+combined <- top/bottom
+combined
+
+find_ratio_bound <- function(e_x, sem_x, e_y, sem_y, quantile_value, randoms = 10000){
+  # monte carlo confidence interval estimate for a ratio. Assumes the two variables
+  # x and y are normally distributed. Finds confidence interval for (x-y)/y
+
+  # randomly generate possible global means for x and y
+
+  xs = rnorm(randoms, mean = e_x, sd=sem_x)  # risk
+  ys = rnorm(randoms, mean = e_y, sd=sem_y)  # healthy
+
+  ratios = (xs-ys)/ys
+
+  return(quantile(ratios, quantile_value))
+}
+
+peak_df <- filtered_df %>%
+  ungroup() %>%
+  group_by(pos) %>%
+  mutate(max_at_pos = max(normalised_counts)) %>%  # at least 1 of the 4 samples above the threshold
+  ungroup() %>%
+  filter(max_at_pos >= peak_t) %>%  # at least 1 of the 4 samples above the threshold
+  group_by(condition, pos) %>%
+  mutate(peak_mean = mean(normalised_counts), 
+         peak_sd = sd(normalised_counts)) %>%
+  ungroup() %>%
+  select(condition, pos, peak_mean, peak_sd) %>%
+  unique() %>%
+  pivot_wider(names_from = "condition", values_from = c("peak_mean", "peak_sd")) %>%
+  mutate(frac_increase = (peak_mean_R-peak_mean_H)/peak_mean_H) %>%
+  mutate(peak_rank = rank(frac_increase)) %>%
+  mutate(colour = ifelse(abs(pos-exon_snp)<region_length, "gold", 
+                         ifelse(abs(pos-intron_snp)<region_length, "purple", "grey50")))
+
+peak_df$upper_bound <- mapply(find_ratio_bound, 
+                        peak_df$peak_mean_R,
+                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
+                        peak_df$peak_mean_H,
+                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
+                        0.5*(1+confidence_interval))
+
+peak_df$lower_bound <- mapply(find_ratio_bound, 
+                        peak_df$peak_mean_R,
+                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
+                        peak_df$peak_mean_H,
+                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
+                        0.5*(1-confidence_interval))
+
+ggplot(peak_df, aes(x = peak_rank, y = 100*frac_increase, fill=colour)) +
+  geom_bar(stat="identity") +
+  scale_fill_identity() +
+  theme_minimal() +
+  ggpubr::theme_pubr()+
+  ylab("% change in 2x Risk") +
+  xlab("Rank") +
+  geom_errorbar(aes(ymin = 100*lower_bound, ymax = 100*upper_bound, x = peak_rank), 
+                width = 0.3)
+
+```
+
+Supplementary iCLIP Figures
+
+```{r message=FALSE, warning=FALSE}
+
+bars <- filtered_df %>%
+  ungroup() %>%
+  group_by(condition, pos) %>%
+  mutate(mean_normalised = mean(normalised_counts),
+         sd_normalised = sd(normalised_counts)) %>%
+  mutate(colour = ifelse(condition == "H", healthy_colour, risk_colour))
+
+bar_top <- ggplot(bars %>% filter(abs(pos-exon_snp) < 35), aes(x = pos, y = mean_normalised,
+                                                       fill = colour)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_fill_identity() +
+  geom_errorbar(aes(ymin = mean_normalised-sd_normalised, ymax=mean_normalised+sd_normalised),
+                position="dodge") +
+  geom_vline(xintercept = exon_snp, linetype="dashed") +
+  ylab("XLs per 1000") +
+  ggpubr::theme_pubr() +
+  ggtitle("iCLIP signal around CE SNP") +
+  xlab("")
+
+bar_bottom <- ggplot(bars %>% filter(abs(pos-intron_snp) < 35), aes(x = pos, y = mean_normalised,
+                                                       fill = colour)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_fill_identity() +
+  geom_errorbar(aes(ymin = mean_normalised-sd_normalised, ymax=mean_normalised+sd_normalised),
+                position="dodge") +
+  geom_vline(xintercept = intron_snp, linetype="dashed") +
+  ggpubr::theme_pubr() +
+  ylab("XLs per 1000") +
+  ggtitle("iCLIP signal around intronic SNP") +
+  xlab("")
+
+bar_combined <- bar_top / bar_bottom
+bar_combined
+
+```
+
+# Heptameric analysis
+
+Analyse E-values (higher E value = stronger binding enrichment) for the motifs present at the exon SNP.
+
+```{r message=FALSE, warning=FALSE}
+
+# Read in supplementary 4 from Ray et al 2013
+
+data <- read_tsv(paste0(here(), "/data/heptamers/41586_2013_BFnature12311_MOESM334_ESM.txt"))
+
+for(type in c("exon", "intron")){
+
+    if(type == "exon"){
+        snp <-     "TAAAAGCATGGAT"
+        healthy <- "TAAAAGGATGGAT"
+    } else {
+        snp <-     "GGATGGGTGGATA"
+        healthy <- "GGATGGTTGGATA"
+    }
+
+    all_healthy <- c()
+    all_risk <- c()
+
+    # Make dataframe containing E scores for each relevant 7mer
+
+    for(i in 1:(str_length(snp)-6)){
+        this_snp <- str_sub(snp, i, i+6)
+        this_healthy <- str_sub(healthy, i, i+6)
+
+        all_healthy <- c(all_healthy, str_replace_all(this_healthy, "T", "U"))
+        all_risk <- c(all_risk, str_replace_all(this_snp, "T", "U"))
+
+        row1 <- data %>% filter(`7mer` == str_replace_all(this_snp, "T", "U")) %>%
+            mutate(snp = T, pos = i)
+
+        row2 <- data %>% filter(`7mer` == str_replace_all(this_healthy, "T", "U")) %>%
+            mutate(snp = F, pos = i)
+
+        if(i == 1){
+            full <- bind_rows(row1, row2)
+        } else {
+            full <- bind_rows(full, row1, row2)
+        }
+    }
+
+    # Find average change in E score for the two groups of heptamers
+
+    full2 <- full %>%
+        group_by(pos) %>%
+        filter(n() == 2) %>%
+        ungroup() %>%
+        pivot_longer(cols = contains("set")) %>%
+        select(-`7mer`) %>%
+        pivot_wider(names_from = "snp", values_from = "value") %>%
+        mutate(abs_change = `TRUE`-`FALSE`) %>%
+        filter(!is.na(`TRUE`) & !is.na(`FALSE`)) %>%
+        group_by(name) %>%
+        mutate(mean_abs_change = mean(abs_change)) %>%
+        select(name, mean_abs_change) %>%
+        unique() %>%
+        arrange(mean_abs_change) %>%
+        mutate(is_tdp43 = ifelse(str_detect(name, "RNCMPT00076"), "red", "grey80")) %>%
+        mutate(short_name = str_sub(name, 1, 11)) %>%
+        ungroup() %>% group_by(short_name) %>%
+        mutate(average_delta = mean(mean_abs_change)) %>%
+        select(short_name, is_tdp43, average_delta) %>%
+        unique()
+
+    # Plot ranks
+
+    plot <- ggplot(full2, aes(x=rank(average_delta), y = average_delta,
+                      fill = is_tdp43)) +
+        geom_bar(stat="identity") +
+        scale_fill_identity() +
+        theme_minimal() +
+        ggtitle(paste0("Average change in binding E value for each dataset with SNP for ", type, " SNP")) +
+        xlab("") +
+        ylab("Average change in E score") +
+        ggpubr::theme_pubr() +
+        xlab("Rank")
+
+    print(plot)
+
+    full_just_tdp <- full %>%
+        group_by(pos) %>%
+        filter(n() == 2) %>%
+        ungroup() %>%
+        select(`7mer`, contains("RNCMPT00076"), snp, pos) %>%
+        mutate(mean_E = (RNCMPT00076_e_setA + RNCMPT00076_e_setB)/2) %>%
+        pivot_longer(cols = contains("RNCMP")) %>%
+        mutate(ypos = ifelse(snp, -0.35, -0.27))
+
+    plot2 <- ggplot(full_just_tdp, aes(x = pos, y = value, colour = snp,
+                              label = `7mer`)) +
+        geom_point(size = 3) +
+        xlab("7mer") +
+        ylab("E value (similar to rank?)") +
+        ggeasy::easy_add_legend_title("Contains Risk SNP") +
+        ggtitle(paste0("Binding E value for TDP43 for all 7x 7mers with or without risk SNP for ", type, " SNP")) +
+        geom_label(aes(y = ypos), size = 2.7) +
+        xlim(0.5, 7.5) +
+        theme_bw() + 
+        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  +
+        ggeasy::easy_remove_x_axis()
+
+    print(plot2)
+
+}
+
+```
+
+# Isothermal Titration Calorimetry
+
+Read in values
+
+```{r warning=FALSE, message=FALSE}
+df <- read_csv(paste0(here(), 
+                      "/data/itc/itc final values.csv")) %>%
+    group_by(RNA) %>%
+    mutate(sd =sd(kd)) %>%
+    mutate(mean_kd = mean(kd))
+
+ggplot(df, aes(x = RNA, y = kd)) +
+    geom_bar(aes(x = RNA, y = mean_kd), stat="identity", position="dodge") +
+    geom_point() +
+    geom_errorbar(aes(x=RNA, ymin = mean_kd-sd, ymax = mean_kd+sd),
+                  width = 0.3) +
+    ylim(0, NA) +
+    ggpubr::theme_pubr() +
+    ylab("Kd / nM") +
+    xlab("")
+
+t.test(df$kd[which(df$RNA=="H")], df$kd[which(df$RNA=="R")])
+```

### 2021-03-26 paper_figures.Rmd

@@
         toc_float: true
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'docs/index.html')) })
 ---
@@
        "patchwork",
-       "rstatix")
+       "rstatix",
+       "ggeasy")
 ```

### 2021-06-28 paper_figures.Rmd
alsod
Code for scrapping ALS mutations from ALSoD website + technical variates

@@
 
+<!-- # iPSC WTC11 parent line genotype -->
+
+<!-- ```{r confirm_no_als_ftd_wtc11 mutations} -->
+<!-- library(vcf) -->
+<!-- wtc11_vcf = "/Users/annaleigh/Downloads/3546dc62_AH77TTBBXX_DS-229105_GCCAAT_recalibrated.vcf.gz" -->
+<!-- vcf <- read.vcfR( wtc11_vcf, verbose = FALSE ) -->
+<!-- ``` -->
+
 # UNC13A CE PSI across TDP-43 KD experiments
@@
 
+## scatterplot UNC13A CE and UNC13B NMD
+
+```{r}
+rel_rna_cryptic_amount %>%
+    filter(condition != "control") %>%
+    ggplot() +
+    geom_point(aes(x = cryptic_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
+    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
+    theme(legend.position = 'bottom')+
+    facet_wrap(~source)
+
+```
 # NMD experiment on SH-SY5Y cells
@@
 
-```{r}
+```{r Inclusion reads by if TDP-potential}
 ####Inclusion reads by if TDP-potential####
@@
 
-boxplot_table %>%
-  filter(disease_tissue == T) %>%
+clean_data_table %>%
+  filter(disease_tissue == T) %>%
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
   mutate(detected = inclusion_reads > 0) %>%
@@
   xlab("N individuals")
+```
+
+## Effect of library size on CE discovery
+
+```{r unc detection by library size  }
+# library size
+
+clean_data_table %>%
+   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
+  filter(disease_tissue == T) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
+  ggplot(aes(x = unc13a_detected,y = log10(library_size) )) + 
+  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+  geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
+  labs(title = "Library size and UNC13A CE detection ",subtitle = "Wilcox test") + 
+  facet_wrap(~ tissue + disease_group2) + 
+  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
+  labs(y = "log10 library size", x = "UNC13A CE detected") +
+  theme(axis.text.x = element_text(size = 15)) + 
+  ggpubr::theme_pubr()
+```
+
+## Effect of UNC13A TPM onCE discovery
+
+```{r unc detection by UNC13A tpm }
+# library size
+
+clean_data_table %>%
+   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
+  filter(disease_tissue == T) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
+  ggplot(aes(x = unc13a_detected,y = UNC13A_TPM )) + 
+  geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+  geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
+  labs(title = "UNC13A TPM and UNC13A CE detection ",subtitle = "Wilcox test") + 
+  facet_wrap(~ tissue + disease_group2, scales = 'free_y') + 
+  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
+  labs(y = "UNC13A expression (TPM)", x = "UNC13A CE detection") +
+  theme(axis.text.x = element_text(size = 15)) +
+  ggpubr::theme_pubr()
+
+```
+
+## Effect of RIN and UNC13A discovery
+
+```{r rin discovery}
+## effect of RIN?
+clean_data_table %>%
+  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
+  filter(disease_tissue == T) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  filter(!is.na(rin)) %>%
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
+  ggplot(aes(x = unc13a_detected,y = rin )) + 
+  geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+   geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
+  labs(title = "Sample RIN and UNC13A CE detection ",subtitle = "Wilcox test") + 
+  facet_wrap(~disease_group2 + tissue) + 
+  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center" ) +
+  labs(y = "RIN", x = "UNC13A CE detection") +
+  theme(axis.text.x = element_text(size = 15)) +
+  ggpubr::theme_pubr()
+
+```
+
+```{r unc13a tpm and ce detection}
+## effect of RIN?
+clean_data_table %>%
+  filter(disease_tissue == T) %>% 
+  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
+  add_count(tissue_clean,disease_group2) %>% 
+  mutate(plot_name = glue::glue("{disease_group2} \n ({n})")) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
+  ggplot(aes(x = plot_name, fill = unc13a_detected,y = UNC13B_TPM )) + 
+  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
+  geom_boxplot(outlier.colour = NA) +
+  labs(title = "UNC13A CE detection and UNC13B TPM",subtitle = "Wilcox test") + 
+  facet_wrap(~ tissue_clean,scales = 'free') + 
+  ggpubr::stat_compare_means(label = "p.signif",) +
+  labs(y = "UNC13B TPM", x = "UNC13A CE detection") +
+  theme(axis.text.x = element_text(size = 15)) +
+  ggpubr::theme_pubr()
+```
+
+## Effect of celltype composition on UNC13A CE detection
+
+```{r}
+conversion_table <- data.frame(short = c("Ast","End", "Mic", "Ex", "Oli", "Per" ),
+                               long = c("Astrocytes", "Endothelial", "Microglia", "Neurons", "Oligos", "Pericytes")
+  )
+
+cell_types = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/MuSiC_residual_results.tsv")
+
+cell_types %>% 
+  left_join(clean_data_table %>%
+   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
+  filter(disease_tissue == T) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ))
+
 ```
@@
 
+# Relationship between STMN2 and UNC13B fsE PSI
+## Scatter plot showing the correlation in non-log space in cortex UNC13B fsE & STMN2 CE
+
+```{r}
+clean_data_table %>%
+  filter(disease_tissue == T) %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  filter(junction_reads_stmn2 >= 30) %>%
+  filter(junction_reads_unc13a >= 30) %>%
+  mutate(junction_reads_unc13b = UNC13B_nmd_junction_5prime + UNC13B_annotated + UNC13B_nmd_junction_3prime) %>% 
+  filter(junction_reads_unc13b >= 30) %>%
+  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
+  # filter(STMN2_cryptic_leaf >= 2) %>%
+  filter(grepl("Cortex",tissue_clean)) %>%
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+  filter(unc13a_cryptic_leaf_psi > 0) %>%
+  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = unc13b_nmd_psi)) +
+  geom_point() +
+  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
+  geom_smooth(method = lm, se = FALSE,color = "black") +
+  xlab("UNC13A CE PSI") +
+  ylab("UNC13B fsE PSI") +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 18)) +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
+  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
+```
+
 ## Side by side comparison UNC13A & STMN2 CE PSI

### 2021-06-30 paper_figures.Rmd
add wtc11 snp cleaner

@@
 
-<!-- # iPSC WTC11 parent line genotype -->
-
-<!-- ```{r confirm_no_als_ftd_wtc11 mutations} -->
-<!-- library(vcf) -->
-<!-- wtc11_vcf = "/Users/annaleigh/Downloads/3546dc62_AH77TTBBXX_DS-229105_GCCAAT_recalibrated.vcf.gz" -->
-<!-- vcf <- read.vcfR( wtc11_vcf, verbose = FALSE ) -->
-<!-- ``` -->
-
 # UNC13A CE PSI across TDP-43 KD experiments
@@
 clean_data_table %>%
-   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
   filter(disease_tissue == T) %>% 
   mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
-  ggplot(aes(x = unc13a_detected,y = log10(library_size) )) + 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+  mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
+  ggplot(aes(x = disease_group2,y = log10(library_size), fill = unc13a_detected)) + 
   geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
-  geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
+  geom_boxplot(colour = "black", outlier.colour = NA) +
   labs(title = "Library size and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  facet_wrap(~ tissue + disease_group2) + 
+  facet_wrap(~ tissue, scales = 'free_x') + 
   ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
-  labs(y = "log10 library size", x = "UNC13A CE detected") +
+  labs(y = "log10 library size", fill= "UNC13A CE detected", x = element_blank()) +
   theme(axis.text.x = element_text(size = 15)) + 
@@
 clean_data_table %>%
-   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
     filter(disease_tissue == T) %>% 
   mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
-  ggplot(aes(x = unc13a_detected,y = UNC13A_TPM )) + 
-  geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
-  geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+  mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
+  ggplot(aes(x = disease_group2,y = UNC13A_TPM, fill = unc13a_detected)) + 
+  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+  geom_boxplot(colour = "black", outlier.colour = NA) +
   labs(title = "UNC13A TPM and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  facet_wrap(~ tissue + disease_group2, scales = 'free_y') + 
+  facet_wrap(~ tissue, scales = 'free') + 
   ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
-  labs(y = "UNC13A expression (TPM)", x = "UNC13A CE detection") +
+  labs(y = "UNC13A TPM", fill= "UNC13A CE detected", x = element_blank()) +
   theme(axis.text.x = element_text(size = 15)) + 
@@
 
-```{r unc13a tpm and ce detection}
+## Effect of UNC13A discovery and UNC13B TPM
+
+```{r unc13a ce and UNC13B_TPM}
 ## effect of RIN?
@@
 cell_types %>% 
-  left_join(clean_data_table %>%
+  left_join(clean_data_table, by = 'sample') %>%
   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
@@
   mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ))
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+  filter(!is.na(inclusion_reads)) %>% 
+  select(sample,cell,deconv, tissue_clean,unc13a_detected) %>% 
+  unique() %>% 
+  ggplot() + 
+  geom_boxplot(aes(x = unc13a_detected, y = deconv, fill = cell)) + 
+  facet_wrap(~ tissue_clean,scales = 'free')
 

### 2021-06-30 cleaning_wtc11_vcf.R

@@ -1,0 +1,64 @@
+library(data.table)
+library(VariantAnnotation)
+library(tidyverse)
+#get the genomic coords of the revelvant ALS/FTD genes 
+als_ftd_risk_genes = c("AARS2", "ALS2", "ALS3", "ALS7", "ANG", "APBB2", "APOE", "APP", 
+                       "ATP13A2", "ATXN2", "C19orf12", "C6orf10", "C9orf72", "CAPZA1", 
+                       "CCNF", "CDH22", "CFAP410", "CHCHD10", "CHMP2B", "CHRNA3,4,B4", 
+                       "CHRNA4", "CNTN6", "CRYM", "CSF1R", "CTSF", "CX3CR1", "DAO", 
+                       "DCTN1", "ELP3", "EPB41L1", "ERBB4", "EWSR1", "FIG4", "FLNC", 
+                       "FUS", "GFRA2", "GLE1", "GRN", "GSN", "HNRNPA1", "HNRNPA2B1", 
+                       "ITM2B", "KIF5A", "LMNB1", "LOC101929163", "LRRK2", "MAPT", "MATR3", 
+                       "MPO", "NEFH", "NEK1", "NIPA1", "NOTCH3", "NPC2", "OPTN", "PFN1", 
+                       "PLD3", "PON1-3", "PRKAR1B", "PRNP", "PRPH", "PSEN1", "PSEN2", 
+                       "RNPLA6", "SETX", "SIGMAR1", "SOD1", "SORT1", "SPG11", "SQSTM1", 
+                       "SRR", "SS18L1", "TAF15", "TARDBP", "TBK1", "TIA1", "TMEM106B", 
+                       "TNIP1", "TREM2", "TRPM2", "TSC1", "TUBA4A", "UBQLN2", "UNC13A", 
+                       "VAPB", "VCP")
+
+gene_locations = annotables::grch38 %>% filter(symbol %in% als_ftd_risk_genes) %>% dplyr::select(symbol,chr,start,end) %>% filter(
+    chr %in% as.character(1:22)
+) %>% mutate(chr = paste0("chr",chr))
+
+wtc11_vcf = "/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/wtc11_vcf.gz"
+
+gene_gr = makeGRangesFromDataFrame(gene_locations,keep.extra.columns = T)
+
+tab <- TabixFile(wtc11_vcf)
[…]

### 2021-07-06 paper_figures.Rmd
STR genotype + more metrics graphs

@@
 
+## Effect of seqeuncer (read length) size on CE discovery
+
+```{r unc detection by sequencer  }
+# library size 
+#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
+unc13a_sequencing_platform <- clean_data_table %>%
+  filter(tissue_clean %in% c("Frontal_Cortex", 
+                             "Lumbar_Spinal_Cord", 
+                             "Cervical_Spinal_Cord", 
+                             "Motor_Cortex", 
+                             "Temporal_Cortex")) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+  mutate(tissue = as.character(tissue)) %>%
+  filter(sequencing_platform != "") %>% 
+  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP"))
+
+quick_fisher = function(tbl){
+
+  if(nrow(tbl) <2){
+    return(NA)
+  }else{
+    comparison_table = table(unc13a = tbl$unc13a_detected, 
+                             sequencing_platform = tbl$sequencing_platform)
+    p_value = fisher.test(comparison_table)$p.value
+
+    return(p_value)
+  }
+}
+
+unc13a_data_fisher = unc13a_sequencing_platform %>% 
+  filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
+  split(.,list(unique(unc13a_sequencing_platform$tissue_clean), unique(unc13a_sequencing_platform$disease_group2))) %>% 
+  purrr::map_df( quick_fisher, .id = "tissue_clean" ) %>% 
+  melt() %>% 
+  select(variable,value) %>% 
+  separate(variable,sep = "\\.", into = c("tissue_clean","disease_group2"))
+
+unc13a_sequencing_platform_pct = unc13a_sequencing_platform %>%
+  group_by(disease_group2,tissue_clean, sequencing_platform, unc13a_detected) %>% tally() %>%
+  spread(key = unc13a_detected, value = n,fill = 0) %>%
+  mutate(detection = `+` / ( `-` + `+`) ) %>% 
+  mutate(total =  ( `-` + `+`) ) %>% 
+  mutate(plot_name = glue::glue("{sequencing_platform} \n  ({total})")) %>% 
+  left_join(unc13a_data_fisher)
+
+unc13a_sequencing_platform_pct %>%
+  ggplot(aes(x = plot_name, y = detection )) + 
+  geom_col(fill = "firebrick") +
+  labs(title = "UNC13A CE detection and sequencing platform", y = "UNC13A CE detected") + 
+  facet_wrap(~disease_group2 + tissue_clean,scales = 'free_x') +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
+  geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(value,digits = 2) ) )) +
+  theme(legend.text = element_text(size = 15))
+ # scale_y_continuous(expand = c(0,0) ) 
+```
 ## Effect of library size on CE discovery
@@
 clean_data_table %>%
-  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
   filter(disease_tissue == T) %>% 
@@
   mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
-  ggplot(aes(x = unc13a_detected,y = rin )) + 
+    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
+  ggplot(aes(x = disease_group2,y = rin, fill = unc13a_detected)) + 
   geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
-   geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
-  labs(title = "Sample RIN and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  facet_wrap(~disease_group2 + tissue) + 
-  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center" ) +
-  labs(y = "RIN", x = "UNC13A CE detection") +
+   geom_boxplot(colour = "black", outlier.colour = NA) +
+  labs(title = "Sample RIN and UNC13A CE detection ") + 
+  facet_wrap(~ tissue, scales = 'free_x') + 
+  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center") +
+  labs(y = "RIN", x = element_blank(), fill = "UNC13A CE detection") +
   theme(axis.text.x = element_text(size = 15)) +
@@
   theme(axis.text.x = element_text(size = 15)) +
-  ggpubr::theme_pubr()
+  ggpubr::theme_pubr() +
+  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center")
 ```
@@
 
-```{r}
-conversion_table <- data.frame(short = c("Ast","End", "Mic", "Ex", "Oli", "Per" ),
-                               long = c("Astrocytes", "Endothelial", "Microglia", "Neurons", "Oligos", "Pericytes")
-  )
-
-cell_types = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/MuSiC_residual_results.tsv")
+```{r celltype and detection}
+
+cell_types = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/All_1917_samples_Darmanis_dtangle_deconv.tsv")
 
 cell_types %>% 
-  left_join(clean_data_table, by = 'sample') %>%
+  left_join(clean_data_table[,.(sample,
+                               UNC13A_3prime_leaf,
+                               UNC13A_5prime_1_leaf,
+                               disease_group2,
+                               disease_tissue)], by = 'sample') %>%
   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
   filter(disease_tissue == T) %>% 
+  filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
   mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
   mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+  add_count(disease_group2,tissue_clean) %>% 
+  mutate(n = n / 5) %>% 
+  mutate(plot_name = glue::glue("{disease_group2} \n {n}")) %>% 
   filter(!is.na(inclusion_reads)) %>% 
-  select(sample,cell,deconv, tissue_clean,unc13a_detected) %>% 
   unique() %>% 
-  ggplot() + 
-  geom_boxplot(aes(x = unc13a_detected, y = deconv, fill = cell)) + 
-  facet_wrap(~ tissue_clean,scales = 'free')
+  ggplot(aes(x = cell, y = deconv, fill = unc13a_detected)) + 
+  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
+  geom_boxplot(outlier.color = NA) + 
+  facet_wrap(~plot_name+tissue,scales = 'free') +
+  ggpubr::stat_compare_means(label = "p.signif") +
+  ggpubr::stat_compare_means(label = "p.format",vjust = 1.2) +
+  labs(title = "Cell type decomposition and UNC13A CE detection ",subtitle = "Wilcox test") + 
+  ggpubr::theme_pubr()
 
@@
     scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
-  ggtitle("intronic SNP (rs12973192)")
+  ggtitle("intronic SNP (rs12608932)")
 ```
@@
 
+```{r}
+pretty_effect_plot = function(model, p_value_cutoff = 0.05, title = "Model 1"){
+    p_value = Tmisc::lmp(model)
+    r_squared = summary(model)$adj.r.squared
+    pretty_plot = broom::tidy(model) %>% 
+        filter(p.value < p_value_cutoff) %>%
+        mutate(term = fct_reorder(term, -abs(estimate))) %>%
+        ggplot(aes(x = term, y = estimate,fill = estimate > 0 )) +
+        geom_col() +
+        coord_flip() +
+        ggtitle(title,subtitle = glue::glue("Model p-value: {p_value} \n Adj. R-squared: {r_squared}"))
+}
+
+detect2 = clean_data_table %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+  filter(junction_reads_stmn2 >= 30) %>%
+  filter(junction_reads_unc13a >= 30) %>%
+  filter(disease_tissue == T) %>% 
+  mutate(call = fct_relevel(rs12973192,
+                              "C/C", "C/G", "G/G")) %>% 
+    mutate(number_g_alleles = as.numeric(call)) %>% 
+  mutate(age_at_death = as.numeric(age_at_death))
+
+m1 = lm(unc13a_cryptic_leaf_psi ~ stmn_2_cryptic_psi_leaf * 
+          number_g_alleles + 
+          tissue_clean + 
+          age_at_death + 
+          sex, 
+              detect2)
+
+k = pretty_effect_plot(m1,p_value_cutoff = 0.05)
+print(k)
+
+```
+
 ## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932
@@
   geom_hline(yintercept = 0,size = 1.5)
+```
+
+## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - STR - cortex and cord
+
+```{r str and log2 graph}
+str_genotype = janitor::clean_names(fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/rs56041637.genotypes.377samples.txt")) %>% 
+  dplyr::rename(presumed_unique_id = number_sample_id)
+
+detected_table_for_graph = clean_data_table %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+  filter(junction_reads_stmn2 >= 30) %>%
+  filter(junction_reads_unc13a >= 30) %>%
+  filter(disease_tissue == T) %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  dplyr::select(sample,participant_id,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
+  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
+  left_join(str_genotype) %>% 
+    filter(!is.na(genotype_ci)) %>% unique()
+
+comparisons = combn(unique(detected_table_for_graph$genotype),2,simplify = F)
+
+detected_table_for_graph %>%
+  ggplot(aes(y = log2_unc, x = genotype,fill = genotype)) +
+  geom_boxplot(show.legend = F) +
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
+  stat_compare_means( label = "p.format") +
+  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
+  ylab("Ratio UNC13A CE PSI / STMN2 PSI") +
+  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
+  xlab("UNC13A STR Genotype") +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 0,size = 1.5)
+```
+
+## STR by SNP genotype
+
+```{r}
+clean_data_table %>% 
+  mutate(presumed_unique_id = gsub("-b38","",participant_id)) %>% 
+  select(participant_id,presumed_unique_id,rs12973192,rs12608932) %>% 
+  unique() %>% 
+  left_join(str_genotype) %>% 
+  unique() %>% 
+  mutate(genotype = fct_relevel(genotype,"6/6", "6/7", "6/8", "6/9",
+                                "6/10","6/12","7/8","8/8","8/10","9/10")) %>% 
+  ggplot(aes(x = genotype,fill = rs12973192)) + 
+  geom_bar() + 
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ylab("N individuals") +
+  xlab("STR genotype")
+
 ```
@@
 read_length_filter <- 29  # length of reads of interest
-gencode_v29_info <- read_tsv(paste0(here(), "/data/longest_proteincoding_transcript_hs_details.txt"))
-ribo_df <- read_tsv(paste0(here(), "/data/riboseq/sh_tdpb.Aligned.toTranscriptome.out.bam.bed.gz"),
-                    col_names = c("transcript_id", "start0", "end", "strand")) %>%
+gencode_v29_info <- fread(paste0(here(), "/data/longest_proteincoding_transcript_hs_details.txt"))
+ribo_df <- fread(paste0(here(), "/data/riboseq/sh_tdpb.Aligned.toTranscriptome.out.bam.bed.gz")) 
+names(ribo_df)  = c("transcript_id", "start0", "end", "strand")
+ribo_df = ribo_df %>% 
     inner_join(gencode_v29_info, by = "transcript_id") %>%
@@
 ```
+
+```{r als_ftd_genes_venn}
+als_ftd_genomics = c("APP", "CHMP2B", "CSF1R", "DNAJC5","UNC13A",
+                     "DNMT1", "EPM2A", "GRN", "ITM2B",
+                     "MAPT", "NHLRC1", "NOTCH3", "PRNP", 
+                     "PSEN1", "PSEN2", "TARDBP", "TBK1", 
+                     "TYROBP", "UBQLN2", "VCP", "ATN1", 
+                     "ATXN10", "ATXN1", "ATXN2", "C9orf72", 
+                     "HTT", "JPH3", "NOP56", "TBP", "ALS2", 
+                     "ANG", "DCTN1", "FIG4", "FUS", "HNRNPA1", 
+                     "OPTN", "PFN1", "SETX", "SIGMAR1", "SLC52A2", 
+                     "SLC52A3", "SOD1", "VAPB", "AR")
+
+ipsc_de = as.data.table(ipsc_de)
+ipsc_splicing = as.data.table(ipsc_splicing)
+
+ipsc_de[gene_name %in% als_ftd_risk_genes][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique()
+```

### 2021-07-08 preprocessing_scripts/Nanopore/nanopore_analysis.R
add nanopore analysis

@@ -1,0 +1,223 @@
+library(tidyverse)
+library(Rsamtools)
+
+find_coverage_density <- function(vec, q_start, q_end, 
+                                  buffer = 3, l = 100000){
+    # The purpose of this function is to detect 
+    
+    q_start <- q_start + buffer
+    q_end <- q_end - buffer
+    if(q_end < q_start){
+        q_start <- round((q_start+q_end)/2)
+        q_end <- q_start
+    }
+    
+    counts_this_region = sum(vec[q_start:q_end])
+    density <- counts_this_region/(q_end-q_start+1)
+    
+    return(density)
+}
+
+make_coverage_vector <- function(cigar, start, vector_l=1000000,
+                                 find_pos = F){
+    # The purpose of this function is to return a vector of 0s and 1s where
+    # if the read is aligned at the position it's a 1, otherwise it's a 0
+    
+    # First split the cigar string into its components
+    # S = soft clip, M = match (though could contain mismatches), 
+    # I = insert in record, # D = deletion in record, N = skipped region 
+    # from the reference, H = hard clip
+    cig_split <- unlist(strsplit(cigar, "(?<=S|M|I|D|N|H)", perl=TRUE))
+    
+    # generate a vector of 0s
+    v <- rep(0, vector_l)
+    
+    # initialise the first position
+    pos <- start
+    
+    for(cig in cig_split){
+        # Determine which type of alignment this section is
+        type <- str_sub(cig, str_length(cig), str_length(cig))
+        
+        # Find the length of this section
+        cig_length <- as.numeric(str_sub(cig, 1, str_length(cig)-1))
+        
+        
+        # If it's aligned (note that this may include mismatches) then it's 1
+        if(type == "M"){
+            v[pos:(pos+cig_length-1)] <- 1
+        } 
+        
+        # Move the position along the reference by the cig length, unless 
+        # this section is an insert or clipping
+        if(!type %in% c("I", "S", "H")){
+            pos <- pos + cig_length
+        }
+    }
+    
+    if(find_pos){
+        return(pos)
+    } else{
+        return(v)
+    }
+    
+}
+
+### State the positions of interest for the reference
+
+# The intron retention event that is TDP-43 regulated
+ir_start <- 61106 
+ir_end <- 62405
+
+# the cryptic exon (just the short one)  CHECK THIS
+cryptic_start <- 47804
[…]

### 2021-07-08 preprocessing_scripts/Nanopore/nanopore_analysis.R
Update nanopore_analysis.R

@@ -2,17 +2,28 @@ library(tidyverse)
 library(Rsamtools)
 
 find_coverage_density <- function(vec, q_start, q_end, 
-                                  buffer = 3, l = 100000){
-    # The purpose of this function is to detect 
+                                  buffer = 3){
+    # The purpose of this function is to calculate the fraction of the region
+    # to which the record is aligned. Eg if this is a 100 nt region and 80 nt
+    # of the reference are aligned then it would return 0.8 .
     
+    # Buffer is a small amount that is trimmed from the 5' and 3' end to reduce
+    # noise at the ends of regions.
+    
+    # Update start and ends with the buffer
     q_start <- q_start + buffer
     q_end <- q_end - buffer
+    
+    # if, after applying the buffer, the size is now very small, set the size 
+    # to 1
     if(q_end < q_start){
         q_start <- round((q_start+q_end)/2)
         q_end <- q_start
     }
     
+    # Find the number of aligned bases in this region
     counts_this_region = sum(vec[q_start:q_end])
+    # Divide by the length of the region
     density <- counts_this_region/(q_end-q_start+1)
     
     return(density)
@@ -69,7 +80,7 @@ make_coverage_vector <- function(cigar, start, vector_l=1000000,
 ir_start <- 61106 
 ir_end <- 62405
 
-# the cryptic exon (just the short one)  CHECK THIS
+# the cryptic exon (just the short one)  TODO check this
 cryptic_start <- 47804
 cryptic_end <- cryptic_start+127
 
@@ -155,6 +166,9 @@ for(sample in samples){
                                        cryptic_end) # last 20 nucleotides of IR
     }
     
+    # Classify each read based on whether it contains the cryptic or IR
+    # Due to the relatively high error rate of nanopore this is performed 
+    # in a permissive way (in both directions)
     full_reads <- bam_df %>%
         filter(upstream_d > 0.7 | end_cryptic_d > 0.9) %>% # some -ve strand reads may terminate within cryptic
         filter(downstream_d > 0.5 | start_ir_d > 0.3) %>% # some +ve strand reads might terminate in intron - should still include these
@@ -163,13 +177,13 @@ for(sample in samples){
                                               ifelse(end_cryptic_d < 0.1 & start_ir_d > 0.3, "IR",
                                                      ifelse(end_cryptic_d > 0.8 & start_ir_d > 0.3, "both", NA)))))
     
[…]

### 2021-07-14 paper_figures.Rmd
Update paper_figures.Rmd

@@
 
-## Effect of seqeuncer (read length) size on CE discovery
+## Systematic differences in  (read length) between cases and controls
+
+```{r sequences and case/controls}
+# library size 
+#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
+case_control_sequencing <- clean_data_table %>%
+  filter(tissue_clean %in% c("Frontal_Cortex", 
+                             "Lumbar_Spinal_Cord", 
+                             "Cervical_Spinal_Cord", 
+                             "Motor_Cortex", 
+                             "Temporal_Cortex",
+                             "Cerebellum")) %>% 
+  filter(sequencing_platform != "")
+
+quick_chiq = function(tbl){
+
+    if(length(unique(tbl$sequencing_platform)) < 2){
+      return(NA)
+    }
+    comparison_table = table(disease = tbl$disease_group2, 
+                             sequencing_platform = tbl$sequencing_platform)
+    print(comparison_table)
+
+    p_value = chisq.test(comparison_table) %>%  broom::tidy() %>% pull(p.value)
+    browser()
+    print(p_value)
+
+    return(p_value)
+}
+
+case_control_sequencing %>% 
+  group_by(tissue_clean) %>% 
+  nest()  %>% 
+  mutate(chiq_pvalue = map_dbl(data, quick_chiq))
+
+unc13a_data_fisher$data = NULL
+
+unc13a_sequencing_platform_pct = unc13a_sequencing_platform %>%
+  group_by(disease_group2,tissue_clean, sequencing_platform, unc13a_detected) %>% tally() %>%
+  spread(key = unc13a_detected, value = n,fill = 0) %>%
+  mutate(detection = `+` / ( `-` + `+`) ) %>% 
+  mutate(total =  ( `-` + `+`) ) %>% 
+  mutate(plot_name = glue::glue("{sequencing_platform} \n  ({total})")) %>% 
+  left_join(unc13a_data_fisher)
+
+unc13a_sequencing_platform_pct %>%
+  ggplot(aes(x = plot_name, y = detection )) + 
+  geom_col(fill = "firebrick") +
+  labs(title = "UNC13A CE detection and sequencing platform", 
+       y = "UNC13A CE detected",x = element_blank(), subtitle = "Fisher exact test") + 
+  facet_wrap(~disease_group2 + tissue_clean,scales = 'free_x') +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
+  geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(fisher_pvalue,digits = 2) ) )) +
+  theme(legend.text = element_text(size = 15)) +
+  ggpubr::theme_pubr()
+ # scale_y_continuous(expand = c(0,0) ) 
+```
+
+## Effect of sequencing machine (read length) on CE discovery
 
@@
 
-  if(nrow(tbl) <2){
+    if(length(unique(tbl$sequencing_platform)) < 2){
       return(NA)
-  }else{
+    }
     comparison_table = table(unc13a = tbl$unc13a_detected, 
@@
 }
-}
 
@@
   filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
-  split(.,list(unique(unc13a_sequencing_platform$tissue_clean), unique(unc13a_sequencing_platform$disease_group2))) %>% 
-  purrr::map_df( quick_fisher, .id = "tissue_clean" ) %>% 
-  melt() %>% 
-  select(variable,value) %>% 
-  separate(variable,sep = "\\.", into = c("tissue_clean","disease_group2"))
+  group_by(disease_group2,tissue_clean) %>% 
+  nest() %>% 
+  mutate(fisher_pvalue = map_dbl(data, quick_fisher))
+
+unc13a_data_fisher$data = NULL
 
@@
   geom_col(fill = "firebrick") +
-  labs(title = "UNC13A CE detection and sequencing platform", y = "UNC13A CE detected") + 
+  labs(title = "UNC13A CE detection and sequencing platform", 
+       y = "UNC13A CE detected",x = element_blank(), subtitle = "Fisher exact test") + 
   facet_wrap(~disease_group2 + tissue_clean,scales = 'free_x') +
   scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
-  geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(value,digits = 2) ) )) +
-  theme(legend.text = element_text(size = 15))
+  geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(fisher_pvalue,digits = 2) ) )) +
+  theme(legend.text = element_text(size = 15)) +
+  ggpubr::theme_pubr()
  # scale_y_continuous(expand = c(0,0) ) 
@@
 # library size 
-
-clean_data_table %>%
-  filter(disease_tissue == T) %>% 
+force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
+names(force_colors) = c("Control","ALS \n non-TDP","ALS-TDP","FTLD \n non-TDP","FTLD-TDP")
+
+plot_list = list()
+
+lib_comps = c("Cervical_Spinal_Cord", "Frontal_Cortex", 
+              "Lumbar_Spinal_Cord", "Temporal_Cortex",
+              "Motor_Cortex", "Cerebellum")
+
+plot_list = list()
+for(t in 1:length(lib_comps)){
+  plt = clean_data_table %>%
+     filter(tissue_clean == lib_comps[t]) %>% 
     mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
@@
     mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
-  ggplot(aes(x = disease_group2,y = log10(library_size), fill = unc13a_detected)) + 
-  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+    ggplot(aes(x = disease_group2,y = log10(library_size),fill = disease_group2)) + 
     geom_boxplot(colour = "black", outlier.colour = NA) +
-  labs(title = "Library size and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  facet_wrap(~ tissue, scales = 'free_x') + 
-  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
+    geom_point(pch = 21, size = 2, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+    facet_wrap(~ tissue_clean,scales = 'free_x') + 
+    # ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
+    ggpubr::stat_compare_means(label = "p.signif",
+                               label.x.npc = "center",
+                               method = "wilcox",
+                               comparisons = list(
+                                                  c("Control","ALS-TDP"),
+                                                  c("ALS \n non-TDP","ALS-TDP"))) +
+      ggpubr::stat_compare_means(label = "p.signif", 
+                               label.x.npc = "center", 
+                               method = "wilcox",
+                               comparisons = list(
+                                                  c("Control","FTLD-TDP"),
+                                                  c("FTLD-TDP","FTLD \n non-TDP"))) + 
     labs(y = "log10 library size", fill= "UNC13A CE detected", x = element_blank()) +
     theme(axis.text.x = element_text(size = 15)) + 
-  ggpubr::theme_pubr()
-```
-
-## Effect of UNC13A TPM onCE discovery
-
-```{r unc detection by UNC13A tpm }
-# library size
-
-clean_data_table %>%
-  filter(disease_tissue == T) %>% 
-  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+    ggpubr::theme_pubr() + 
+    scale_fill_manual(values = force_colors) +
+            theme(axis.ticks.x = element_blank(),
+              axis.text.x = element_blank(),
+              legend.position = "none")
+
+  plot_list[[t]] <- plt
+}
+
+cowplot::plot_grid(plotlist = plot_list)
+```
+
+## Effect of RIN and UNC13A discovery
+
+```{r rin discovery}
+force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
+names(force_colors) = c("Control","ALS \n non-TDP","ALS-TDP","FTLD \n non-TDP","FTLD-TDP")
+
+plot_list = list()
+
+lib_comps = c("Cervical_Spinal_Cord", "Frontal_Cortex", 
+              "Lumbar_Spinal_Cord", "Temporal_Cortex",
+              "Motor_Cortex", "Cerebellum")
+
+plot_list = list()
+for(t in 1:length(lib_comps)){
+    if(grepl("Cord",lib_comps[t]) | lib_comps[t] == "Motor_Cortex"){
+        groups = c("Control","ALS-TDP","ALS \n non-TDP")
+    }else if(lib_comps[t] %in% c("Frontal_Cortex","Temporal_Cortex")){
+        groups = c("Control","FTLD-TDP","FTLD \n non-TDP")
+    }else{
+        groups = c("ALS-TDP", "Control", "ALS \n non-TDP", "FTLD-TDP", "FTLD \n non-TDP")
+    }
+
+    plt = clean_data_table %>%
+        filter(tissue_clean == lib_comps[t]) %>% 
         mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
-  ggplot(aes(x = disease_group2,y = UNC13A_TPM, fill = unc13a_detected)) + 
-  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+        filter(disease_group2 %in% groups) %>%
+
+        ggplot(aes(x = disease_group2,y = rin,fill =disease_group2 )) + 
         geom_boxplot(colour = "black", outlier.colour = NA) +
-  labs(title = "UNC13A TPM and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  facet_wrap(~ tissue, scales = 'free') + 
-  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
-  labs(y = "UNC13A TPM", fill= "UNC13A CE detected", x = element_blank()) +
+        geom_point(pch = 21, size = 2, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+        facet_wrap(~ tissue_clean,scales = 'free_x') + 
+        # ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
+        ggpubr::stat_compare_means(label = "p.signif",
+                                   label.x.npc = "center",
+                                   method = "wilcox",
+                                   comparisons = list(
+                                       c("Control","ALS-TDP"),
+                                       c("ALS \n non-TDP","ALS-TDP"))) +
+        ggpubr::stat_compare_means(label = "p.signif", 
+                                   label.x.npc = "center", 
+                                   method = "wilcox",
+                                   comparisons = list(
+                                       c("Control","FTLD-TDP"),
+                                       c("FTLD-TDP","FTLD \n non-TDP"))) + 
+        labs(y = "RIN", x = element_blank()) +
         theme(axis.text.x = element_text(size = 15)) + 
-  ggpubr::theme_pubr()
-
-```
-
-## Effect of RIN and UNC13A discovery
-
-```{r rin discovery}
-## effect of RIN?
-clean_data_table %>%
-  filter(disease_tissue == T) %>% 
-  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  filter(!is.na(rin)) %>%
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
-    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
-  ggplot(aes(x = disease_group2,y = rin, fill = unc13a_detected)) + 
-  geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
-   geom_boxplot(colour = "black", outlier.colour = NA) +
-  labs(title = "Sample RIN and UNC13A CE detection ") + 
-  facet_wrap(~ tissue, scales = 'free_x') + 
-  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center") +
-  labs(y = "RIN", x = element_blank(), fill = "UNC13A CE detection") +
-  theme(axis.text.x = element_text(size = 15)) +
-  ggpubr::theme_pubr()
-
+        ggpubr::theme_pubr() + 
+        theme(axis.ticks.x = element_blank(),
+              axis.text.x = element_blank(),
+              legend.position = "none")
+
+    plot_list[[t]] <- plt
+}
+
+cowplot::plot_grid(plotlist = plot_list,ncol = 2)    
 ```
@@
   facet_wrap(~ tissue_clean,scales = 'free') + 
-  ggpubr::stat_compare_means(label = "p.signif",) +
+  ggpubr::stat_compare_means(label = "p.signif") +
   labs(y = "UNC13B TPM", x = "UNC13A CE detection") +
@@
   ggpubr::theme_pubr() +
-  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center")
+  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center",label.y.npc = 0.9)
+```
+
+## Celltype composition between cases and controls
+
+```{r}
+force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
+names(force_colors) = c("Control","ALS \n non-TDP","ALS-TDP","FTLD \n non-TDP","FTLD-TDP")
+
+plot_list = list()
+
+lib_comps = c("Cervical_Spinal_Cord", "Frontal_Cortex", 
+              "Lumbar_Spinal_Cord", "Temporal_Cortex",
+              "Motor_Cortex", "Cerebellum")
+
+for(t in 1:length(lib_comps)){
+
+    if(grepl("Cord",lib_comps[t]) | lib_comps[t] == "Motor_Cortex"){
+        groups = c("Control","ALS-TDP","ALS \n non-TDP")
+    }else if(lib_comps[t] %in% c("Frontal_Cortex","Temporal_Cortex")){
+        groups = c("Control","FTLD-TDP","FTLD \n non-TDP")
+    }else{
+        groups = c("ALS-TDP", "Control", "ALS \n non-TDP", "FTLD-TDP", "FTLD \n non-TDP")
+    }
+
+    plt = cell_types %>% 
+        left_join(clean_data_table[,.(sample,
+                                      UNC13A_3prime_leaf,
+                                      UNC13A_5prime_1_leaf,
+                                      disease_group2,
+                                      disease_tissue)], by = 'sample') %>%
+        filter(!is.na(disease_group2)) %>% 
+        unique() %>% 
+        filter(tissue_clean == lib_comps[t]) %>% 
+        filter(disease_group2 %in% groups) %>% 
+        mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
+        ggplot(aes(x = disease_group2,y = deconv,fill =disease_group2 )) + 
+        geom_boxplot(colour = "black", outlier.colour = NA) +
+        facet_wrap(~cell,scales = 'free_x',nrow = 1) +
+        ggpubr::stat_compare_means(label = "p.signif", 
+                                   label.x.npc = "center", 
+                                   method = "wilcox",
+                                   comparisons = list(
+                                       c("Control","FTLD-TDP"),
+                                       c("FTLD-TDP","FTLD \n non-TDP"))) +
+        ggpubr::stat_compare_means(label = "p.signif",
+                                   label.x.npc = "center",
+                                   method = "wilcox",
+                                   comparisons = list(
+                                       c("Control","ALS-TDP"),
+                                       c("ALS \n non-TDP","ALS-TDP"))) +
+        ggtitle(lib_comps[t]) + 
+        labs(fill = element_blank(),x = element_blank()) + 
+        geom_point(pch = 21, size = 2, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
+        ggpubr::theme_pubr() +
+        theme(axis.ticks.x = element_blank(),
+              axis.text.x = element_blank(),
+              legend.position = "none") + 
+        scale_fill_manual(values = force_colors)
+plot_list[[t]] <- plt
+
+}
+
+cowplot::plot_grid(plotlist = plot_list,ncol = 2)
+
 ```
@@
 ```{r onlyshort}
-boxplot_table %>%
+only_short_plotter = boxplot_table %>%
     filter(tissue_clean %in%
@@
                                       "Temporal Cortex",
-                                      "Motor Cortex")) %>%
+                                      "Motor Cortex"))
+
+comps = only_short_plotter %>% pull(detection_name) %>%
+  unique() %>%
+  combn(.,2,simplify = F)
+
+only_short_plotter %>% 
     ggplot(aes(x = detection_name,
@@
 
-## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12973192
-
-```{r}
-
-ce_snp = clean_data_table %>%
+## Correlation UNC13A PSI and STMN2 PSI with genotype effect - concordants samples
+
+```{r}
+
+clean_data_table %>%
+  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
+  mutate(snp_status = case_when(rs12973192 == "C/C" ~ "healthy/healthy",
+                                rs12973192 == "C/G" ~ "healthy/risk",
+                                rs12973192 == "G/G" ~ "risk/risk")) %>% 
   filter(disease_tissue == T) %>%
@@
                       y = "log10_unc13a",
-                      color = 'rs12973192',
-                      add = "reg.line", size = 3
+                      fill = 'snp_status',
+                      color = 'black', 
+                      size = 3,shape = 21,
                       ) +
@@
     xlab("STMN2 Cryptic PSI ") +
+  geom_smooth(method = 'lm',aes(color = snp_status),se = F) + 
   stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                       y = unc13a_cryptic_leaf_psi,
-                      color = rs12973192),
+                      color = snp_status),
            method = 'spearman',
@@
     scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+      scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+
   theme(legend.title = element_blank()) +
@@
     ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
-      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
-  ggtitle("CE SNP (rs12973192) genotype")
-
-```
-
-## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12608932
-
-```{r}
-
-in_snp = clean_data_table %>%
-  filter(disease_tissue == T) %>%
-    filter(junction_reads_stmn2 >= 30) %>%
-  filter(junction_reads_unc13a >= 30) %>%
-  filter(grepl("Cortex",tissue_clean)) %>%
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
-    mutate(rs12608932 = fct_relevel(rs12608932,
-                              "A/A", "A/C", "C/C")) %>%
-  filter(!grepl("Cord",tissue_clean)) %>%
-  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>%
-    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>%
-    ggpubr::ggscatter(.,
-                      x = "log10_stmn2",
-                      y = "log10_unc13a",
-                      color = 'rs12608932',
-                      add = "reg.line", size = 3
-                      ) +
-    ylab("UNC13A CE PSI ") +
-    xlab("STMN2 Cryptic PSI ") +
-  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
-                      y = unc13a_cryptic_leaf_psi,
-                      color = rs12608932),
-           method = 'spearman',
-           cor.coef.name = 'rho',
-           show.legend = FALSE,
-           size = 14) +
-    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
-  theme(legend.title = element_blank()) +
-  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
-      ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
-      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
-   ggtitle("intronic SNP (rs12608932) genotype")
-
-```
-
-```{r}
-ggpubr::ggarrange(ce_snp,in_snp)
+      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI")))
+
 ```

### 2021-07-14 preprocessing_scripts/Nanopore/nanopore_analysis.R
Update nanopore_analysis.R

@@ -1,5 +1,7 @@
 library(tidyverse)
 library(Rsamtools)
+library(patchwork)
+
 
 find_coverage_density <- function(vec, q_start, q_end, 
                                   buffer = 3){
@@ -185,7 +187,6 @@ for(sample in samples){
 
 # read in processed files
 csvs <- Sys.glob("processed_again/*.csv")
-
 for(csv in csvs){
     if(csv == csvs[1]){
         combined <- read_csv(csv)
@@ -194,15 +195,17 @@ for(csv in csvs){
     }
 }
 
+
+# make a dataframe linking barcodes to the sample names
 sample_names <- data.frame(barcode = c("bc05", "bc06", "bc07", "bc08", "bc09", "bc10", 
                                        "bc11", "bc12"),
                            name = c("SHSY_NT1", "SHSY_Dox1", 
                                     "SHSY_NT2", "SHSY_Dox2",
-                                    "P56", "P45", "P63", "P28"))
+                                    "FTD1", "FTD2", "FTD3", "FTD4"))
 
-
+# Find the number of reads of each type for each sample
 df <- combined %>%
-    filter(!is.na(classification)) %>%
+    filter(!is.na(classification)) %>%  # remove reads that didn't classify clearly
     dplyr::select(barcode, classification) %>%
     group_by(barcode, classification) %>%
     mutate(n = n()) %>%
@@ -212,26 +215,63 @@ df <- combined %>%
     ungroup() %>%
     complete(barcode, nesting(classification), fill = list(fraction = 0)) %>%
     left_join(sample_names) %>%
-    unique()
+    unique() %>%
+    mutate(class2 = ifelse(classification == "both", "CE+IR", classification)) %>%
+    mutate(type = ifelse(str_detect(name, "SHSY"), "SHSY5Y", "Patient")) %>%
+    ungroup()
 
-p1 <- ggplot(df, aes(x = name, y = 100*fraction, fill = classification)) +
+write_csv(df, "combind_nanopore_dataframe.csv")
+
+
+p2_patient <- ggplot(df %>% filter(classification != "Neither",
+                                   type=="Patient"), aes(x = name, y = 100*fraction, fill = class2)) +
     geom_bar(stat="identity", position="dodge") +
-    ggeasy::easy_rotate_x_labels() +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
[…]

### 2021-07-14 paper_figures.Rmd
nanopore updates

--- 2021-06-30 paper_figures.Rmd
@@
 
-## scatterplot UNC13A CE and UNC13B NMD
-
-```{r}
-rel_rna_cryptic_amount %>%
-    filter(condition != "control") %>%
-    ggplot() +
-    geom_point(aes(x = cryptic_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
-    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
-    theme(legend.position = 'bottom')+
-    facet_wrap(~source)
-
-```
 # NMD experiment on SH-SY5Y cells
@@
 
-```{r Inclusion reads by if TDP-potential}
+```{r}
 ####Inclusion reads by if TDP-potential####
@@
 
-clean_data_table %>%
+boxplot_table %>%
   filter(disease_tissue == T) %>%
-  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
   mutate(detected = inclusion_reads > 0) %>%
@@
   xlab("N individuals")
-```
-
-## Effect of library size on CE discovery
-
-```{r unc detection by library size  }
-# library size
-
-clean_data_table %>%
-  filter(disease_tissue == T) %>% 
-  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
-  mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
-  ggplot(aes(x = disease_group2,y = log10(library_size), fill = unc13a_detected)) + 
-  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
-  geom_boxplot(colour = "black", outlier.colour = NA) +
-  labs(title = "Library size and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  facet_wrap(~ tissue, scales = 'free_x') + 
-  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
-  labs(y = "log10 library size", fill= "UNC13A CE detected", x = element_blank()) +
-  theme(axis.text.x = element_text(size = 15)) + 
-  ggpubr::theme_pubr()
-```
-
-## Effect of UNC13A TPM onCE discovery
-
-```{r unc detection by UNC13A tpm }
-# library size
-
-clean_data_table %>%
-    filter(disease_tissue == T) %>% 
-  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
-  mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
-  ggplot(aes(x = disease_group2,y = UNC13A_TPM, fill = unc13a_detected)) + 
-  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
-  geom_boxplot(colour = "black", outlier.colour = NA) +
-  labs(title = "UNC13A TPM and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  facet_wrap(~ tissue, scales = 'free') + 
-  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
-  labs(y = "UNC13A TPM", fill= "UNC13A CE detected", x = element_blank()) +
-  theme(axis.text.x = element_text(size = 15)) + 
-  ggpubr::theme_pubr()
-
-```
-
-## Effect of RIN and UNC13A discovery
-
-```{r rin discovery}
-## effect of RIN?
-clean_data_table %>%
-  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
-  filter(disease_tissue == T) %>% 
-  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  filter(!is.na(rin)) %>%
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
-  ggplot(aes(x = unc13a_detected,y = rin )) + 
-  geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
-   geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
-  labs(title = "Sample RIN and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  facet_wrap(~disease_group2 + tissue) + 
-  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center" ) +
-  labs(y = "RIN", x = "UNC13A CE detection") +
-  theme(axis.text.x = element_text(size = 15)) +
-  ggpubr::theme_pubr()
-
-```
-
-## Effect of UNC13A discovery and UNC13B TPM
-
-```{r unc13a ce and UNC13B_TPM}
-## effect of RIN?
-clean_data_table %>%
-  filter(disease_tissue == T) %>% 
-  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
-  add_count(tissue_clean,disease_group2) %>% 
-  mutate(plot_name = glue::glue("{disease_group2} \n ({n})")) %>% 
-  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
-  ggplot(aes(x = plot_name, fill = unc13a_detected,y = UNC13B_TPM )) + 
-  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
-  geom_boxplot(outlier.colour = NA) +
-  labs(title = "UNC13A CE detection and UNC13B TPM",subtitle = "Wilcox test") + 
-  facet_wrap(~ tissue_clean,scales = 'free') + 
-  ggpubr::stat_compare_means(label = "p.signif",) +
-  labs(y = "UNC13B TPM", x = "UNC13A CE detection") +
-  theme(axis.text.x = element_text(size = 15)) +
-  ggpubr::theme_pubr()
-```
-
-## Effect of celltype composition on UNC13A CE detection
-
-```{r}
-conversion_table <- data.frame(short = c("Ast","End", "Mic", "Ex", "Oli", "Per" ),
-                               long = c("Astrocytes", "Endothelial", "Microglia", "Neurons", "Oligos", "Pericytes")
-  )
-
-cell_types = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/MuSiC_residual_results.tsv")
-
-cell_types %>% 
-  left_join(clean_data_table, by = 'sample') %>%
-  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
-  filter(disease_tissue == T) %>% 
-  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
-  filter(!is.na(inclusion_reads)) %>% 
-  select(sample,cell,deconv, tissue_clean,unc13a_detected) %>% 
-  unique() %>% 
-  ggplot() + 
-  geom_boxplot(aes(x = unc13a_detected, y = deconv, fill = cell)) + 
-  facet_wrap(~ tissue_clean,scales = 'free')
-
 ```
@@
 
-# Relationship between STMN2 and UNC13B fsE PSI
-## Scatter plot showing the correlation in non-log space in cortex UNC13B fsE & STMN2 CE
-
-```{r}
-clean_data_table %>%
-  filter(disease_tissue == T) %>%
-  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
-  filter(junction_reads_stmn2 >= 30) %>%
-  filter(junction_reads_unc13a >= 30) %>%
-  mutate(junction_reads_unc13b = UNC13B_nmd_junction_5prime + UNC13B_annotated + UNC13B_nmd_junction_3prime) %>% 
-  filter(junction_reads_unc13b >= 30) %>%
-  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
-  # filter(STMN2_cryptic_leaf >= 2) %>%
-  filter(grepl("Cortex",tissue_clean)) %>%
-  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
-  filter(unc13a_cryptic_leaf_psi > 0) %>%
-  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = unc13b_nmd_psi)) +
-  geom_point() +
-  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
-  geom_smooth(method = lm, se = FALSE,color = "black") +
-  xlab("UNC13A CE PSI") +
-  ylab("UNC13B fsE PSI") +
-  ggpubr::theme_pubr() +
-  theme(text = element_text(size = 18)) +
-  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
-  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
-```
-
 ## Side by side comparison UNC13A & STMN2 CE PSI
@@
 
-Read in values
+Read in values, plot and perform a t-test
 
@@
 ```
+
+# Nanopore
+
+Read in pre-processed data (see separate script) and plot
+```{r warning=FALSE, message=FALSE}
+df <- read_csv(paste0(here(),
+                      "/data/nanopore/combined_nanopore_dataframe.csv"))
+
+p2_patient <- ggplot(df %>% filter(classification != "Neither",
+                                   type=="Patient"), aes(x = name, y = 100*fraction, fill = class2)) +
+    geom_bar(stat="identity", position="dodge") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("Patient RNA") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+p2_shsy <- ggplot(df %>% filter(classification != "Neither",
+                                type=="SHSY5Y"), aes(x = name, y = 100*fraction, fill = class2)) +
+    geom_bar(stat="identity", position="dodge") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("SH-SY5Y") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+p2_shsy | p2_patient
+
+# without neither removed
+
+p2_patient <- ggplot(df %>% filter(classification != "",
+                                   type=="Patient"), aes(x = name, y = 100*fraction, fill = class2,
+                                                         colour = class2)) +
+    geom_bar(stat="identity", position="dodge") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("Patient RNA") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+p2_shsy <- ggplot(df %>% filter(classification != "",
+                                type=="SHSY5Y"), aes(x = name, y = 100*fraction, fill = class2,
+                                                     colour = class2)) +
+    geom_bar(stat="identity", position="dodge") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("SH-SY5Y") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("") 
+p2_shsy | p2_patient
+```

### 2021-07-14 preprocessing_scripts/Nanopore/nanopore_analysis.R

@@ -220,7 +220,7 @@ df <- combined %>%
     mutate(type = ifelse(str_detect(name, "SHSY"), "SHSY5Y", "Patient")) %>%
     ungroup()
 
-write_csv(df, "combind_nanopore_dataframe.csv")
+write_csv(df, "combined_nanopore_dataframe.csv")
 
 
 p2_patient <- ggplot(df %>% filter(classification != "Neither",
@@ -244,7 +244,7 @@ p2_shsy <- ggplot(df %>% filter(classification != "Neither",
     ggeasy::easy_add_legend_title("") 
 
 p2_shsy | p2_patient
-ggsave("neither_removed.svg", height = 15, width = 20, units="cm")
+ggsave("nanopore_neither_removed.svg", height = 15, width = 20, units="cm")
 
 # without neither removed
 
@@ -271,7 +271,7 @@ p2_shsy <- ggplot(df %>% filter(classification != "",
     xlab("") +
     ggeasy::easy_add_legend_title("") 
 p2_shsy | p2_patient
-ggsave("neither_not_removed.svg", height = 15, width = 20, units="cm")
+ggsave("nanopore_neither_not_removed.svg", height = 15, width = 20, units="cm")
 
 
 

### 2021-07-19 preprocessing_scripts/Targeted_RNA_seq/random_hexamer/analyse_unc_trnaseq_3.R
Update analyse_unc_trnaseq_3.R

@@ -22,12 +22,12 @@ library(stringdist)
 # list of "true" UMIs. It does this by finding the minimum string distance of each UMI to 
 # the UMIs currently in the list of "true" UMIs. It requires that for a UMI sequence which 
 # has a string distance of n bases, it is at least 1/(10^n) times the abundance of 
-# the most similar UMI. It also requires that each UMI is at least 0.01x times as abundant
+# the most similar UMI. It also requires that each UMI is at least 0.02x times as abundant
 # as the most abundant, and is present in at least 50 reads.
 #
 # This script differs from the specific RT script in that it does not detect correctly spliced transcripts.
 # It also doesn't have a lower limit on the number of reads containing each UMI (though 
-# each must still be at least 0.01x the abundance of the most abundant).
+# each must still be at least 0.02x the abundance of the most abundant).
 #
 # Note that although the forward reads in the fastq files for these samples correspond to the forward
 # strand of the RNA, and thus the MINUS strand of the genome, the bam file format automatically reverse
@@ -38,7 +38,7 @@ get_sample <- function(string){
     return(str_split(string,":")[[1]][2])
 }
 
-find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100){
+find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 50){
     # df should have two columns - "strings" which are the UMIs, and "n" which
     # is the copy number of each
     #
@@ -103,7 +103,7 @@ find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100){
         }
     }
 
-    if(nrow(df_added) > 0){
+    if(nrow(df_added) > 1){
         return(df_added %>% select(-distances))
     } else {
         return(df_added)
@@ -112,16 +112,20 @@ find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100){
 
 ##### RUN
 
-setwd("C:/Users/ogw/Google Drive/UCL PhD/Year 2/unc13a/analyse_targeted_RNASEQ/tg3/")
+#setwd("C:/Users/ogw/Google Drive/UCL PhD/Year 2/unc13a/analyse_targeted_RNASEQ/tg3/")
 
-sample_names = read_csv("sample_names.csv")  # which barcode was used for each sample
[…]

### 2021-07-19 paper_figures.Rmd
Add some analysis
pfkp, rap1gap, etc

--- 2021-07-14 paper_figures.Rmd
@@
               axis.text.x = element_blank(),
-              legend.position = "none")
+              legend.position = "none") + 
+    ylab(element_blank())
 
@@
 
-cowplot::plot_grid(plotlist = plot_list)
+cowplot::plot_grid(plotlist = plot_list,nrow = 3)
 ```
@@
 
+cell_types = fread(paste0(here::here(),"/data/All_1917_samples_Darmanis_dtangle_deconv.tsv"))
 for(t in 1:length(lib_comps)){
@@
 
+# Detection rate as a linear model
+```{r}
+
+```
+
 # UNC13A RNA expression (TPM)
@@
 # Relationship between STMN2 and UNC13A CE PSI
-## Scatter plot showing the correlation in non-log space in cortex UNC13A & STMN2 CE
+## Scatter plot showing the correlation in non-log space UNC13A & STMN2 CE
+
+```{r}
+clean_data_table %>%
+  filter(disease_tissue == T) %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  filter(junction_reads_stmn2 >= 30) %>%
+  filter(junction_reads_unc13a >= 30) %>%
+  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
+  # filter(STMN2_cryptic_leaf >= 2) %>%
+  filter(grepl("Cortex",tissue_clean)) %>%
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+  filter(unc13a_cryptic_leaf_psi > 0) %>%
+  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
+  geom_point() +
+  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
+  geom_smooth(method = lm, se = FALSE,color = "black") +
+  xlab("STMN2 Cryptic PSI") +
+  ylab("UNC13A CE PSI") +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 18)) +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
+  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
+```
+## Scatter plot showing the correlation in non-log space RAP1GAP1 & STMN2 CE
 
@@
 ```{r als_ftd_genes_venn}
-als_ftd_genomics = c("APP", "CHMP2B", "CSF1R", "DNAJC5","UNC13A",
-                     "DNMT1", "EPM2A", "GRN", "ITM2B",
-                     "MAPT", "NHLRC1", "NOTCH3", "PRNP", 
-                     "PSEN1", "PSEN2", "TARDBP", "TBK1", 
-                     "TYROBP", "UBQLN2", "VCP", "ATN1", 
-                     "ATXN10", "ATXN1", "ATXN2", "C9orf72", 
-                     "HTT", "JPH3", "NOP56", "TBP", "ALS2", 
-                     "ANG", "DCTN1", "FIG4", "FUS", "HNRNPA1", 
-                     "OPTN", "PFN1", "SETX", "SIGMAR1", "SLC52A2", 
-                     "SLC52A3", "SOD1", "VAPB", "AR")
+als_ftd_genomics = c("ALS2", "ANG", "APP", "AR", "ATN1", "ATXN1", "ATXN10", "ATXN2", 
+"ATXN3", "C21orf2", "C9orf72", "CHMP2B", "CSF1R", "DCTN1", "DNAJC5", 
+"DNMT1", "EPM2A", "FIG4", "FUS", "GRN", "HNRNPA1", "HTT", "ITM2B", 
+"JPH3", "KIF5A", "LOC101927815", "MAPT", "NHLRC1", "NOP56", "NOTCH3", 
+"OPTN", "PFN1", "PMP22", "PRNP", "PSEN1", "PSEN2", "SCFD1", "SETX", 
+"SIGMAR1", "SLC52A2", "SLC52A3", "SOD1", "TARDBP", "TBK1", "TBP", 
+"TNIP1", "TYROBP", "UBQLN2", "UNC13A", "VAPB", "VCP")
 
@@
 
-ipsc_de[gene_name %in% als_ftd_risk_genes][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique()
-```
+ipsc_de[gene_name %in% als_ftd_genomics][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique() %>% arrange(log2FoldChange) %>% clipr::write_clip()
+```
+
+## Linear model on UNC13A CE PSI with covariates
+```{r the_silly_linear}
+# library size 
+#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
+full_cov = clean_data_table %>%
+    filter(tissue_clean %in% c("Frontal_Cortex", 
+                               "Lumbar_Spinal_Cord", 
+                               "Cervical_Spinal_Cord", 
+                               "Motor_Cortex", 
+                               "Temporal_Cortex")) %>% 
+    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+    filter(sequencing_platform != "") %>% 
+    left_join(cell_types %>% 
+                  dplyr::select(cell,deconv,sample) %>% 
+                  unique() %>% 
+                  pivot_wider(id_cols = "sample",names_from = "cell",values_from = "deconv"), by = 'sample') %>%
+    filter(!is.na(disease_group2)) %>% 
+    unique() %>% 
+    filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>% 
+    dplyr::select(c("unc13a_cryptic_leaf_psi",
+                    "sample",  "tissue_clean",  
+                    "rin",  "sex", 
+                    "age",  "mutations", 
+                    "sequencing_platform", 
+                    "number_g_alleles", 
+                    "stmn_2_cryptic_psi_leaf", "oligodendrocytes", 
+                    "astrocytes", "neurons", "endothelial", "microglia")) %>% 
+    column_to_rownames('sample')
+
+fit <- lm(unc13a_cryptic_leaf_psi ~ ., data = full_cov)
+
+fit %>% broom::tidy() %>% fwrite(paste0(here::here(),"/covariate_unc13a_ce.csv"))
+full_cov %>% 
+    group_by(tissue_clean) %>% 
+    nest() %>% 
+    mutate(lm_obj = map(data, ~lm(unc13a_cryptic_leaf_psi ~., data = .x))) %>% 
+    mutate(broomed = map(lm_obj, broom::tidy)) %>% 
+    dplyr::select(tissue_clean,broomed) %>% unnest() %>% filter(p.value < 0.05) %>% View()
+
+nested %>% 
+    pluck("lm_obj", 1) %>% broom::tidy()
+unc13a_data_fisher = unc13a_sequencing_platform %>% 
+    filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
+    group_by(disease_group2,tissue_clean) %>% 
+    nest() %>% 
+    mutate(fisher_pvalue = map_dbl(data, quick_fisher))
+
+unc13a_data_fisher$data = NULL
+
+unc13a_sequencing_platform_pct = unc13a_sequencing_platform %>%
+    group_by(disease_group2,tissue_clean, sequencing_platform, unc13a_detected) %>% tally() %>%
+    spread(key = unc13a_detected, value = n,fill = 0) %>%
+    mutate(detection = `+` / ( `-` + `+`) ) %>% 
+    mutate(total =  ( `-` + `+`) ) %>% 
+    mutate(plot_name = glue::glue("{sequencing_platform} \n  ({total})")) %>% 
+    left_join(unc13a_data_fisher)
+
+unc13a_sequencing_platform_pct %>%
+    ggplot(aes(x = plot_name, y = detection )) + 
+    geom_col(fill = "firebrick") +
+    labs(title = "UNC13A CE detection and sequencing platform", 
+         y = "UNC13A CE detected",x = element_blank(), subtitle = "Fisher exact test") + 
+    facet_wrap(~disease_group2 + tissue_clean,scales = 'free_x') +
+    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
+    geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(fisher_pvalue,digits = 2) ) )) +
+    theme(legend.text = element_text(size = 15)) +
+    ggpubr::theme_pubr()
+# scale_y_continuous(expand = c(0,0) ) 
+```
+
+# Correlation between STMN2 CE PSI and RAP1GAP1/PFKP
+
+```{r}
+patient_others = fread(file.path(here::here(),"data","nygc_pfkp_unc13b_rap1gap.bed"))
+
+patient_others[,sample := gsub(".SJ.out","",V4)]
+clean_data_table = clean_data_table %>% 
+    left_join(patient_others %>% 
+                  pivot_wider(id_cols = "sample",names_from = "V7",values_from = "V5",values_fill = 0))
+
+clean_data_table[, rap1gap_cryptic_psi := (RAP1GAP_noveldonor + RAP1GAP_novelacceptor)/ (RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated)]
+clean_data_table[, pfkp_cryptic_psi := (PFKP_novelacceptor + PFKP_noveldonor)/ (PFKP_novelacceptor + PFKP_noveldonor + PFKP_annotated)]
+
+pfkp_plot = clean_data_table %>%
+    filter(grepl("Cortex",tissue_clean)) %>%
+    filter(disease_tissue == T) %>%
+    filter((PFKP_novelacceptor + PFKP_noveldonor + PFKP_annotated) >=30) %>% 
+    filter(junction_reads_stmn2 >= 30) %>%
+    filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+    filter(pfkp_cryptic_psi > 0) %>%
+    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = pfkp_cryptic_psi)) +
+    geom_point() +
+    stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
+    geom_smooth(method = lm, se = FALSE,color = 'black') +
+    xlab("STMN2 Cryptic PSI") +
+    ylab("PFKP Cryptic PSI") +
+    ggpubr::theme_pubr() +
+    theme(text = element_text(size = 18)) +
+    scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
+    scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
+
+rap1gap_plot = clean_data_table %>%
+    filter(disease_tissue == T) %>%
+    filter(grepl("Cortex",tissue_clean)) %>%
+    filter((RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated) >=30) %>% 
+    filter(junction_reads_stmn2 >= 30) %>%
+    filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+    filter(rap1gap_cryptic_psi > 0) %>%
+    # ggplot(aes(x = rap1gap_cryptic_psi, y = unc13a_cryptic_leaf_psi)) +
+    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = rap1gap_cryptic_psi)) +
+    geom_point() +
+    stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
+    geom_smooth(method = lm, se = FALSE,color = 'black') +
+    xlab("STMN2 Cryptic PSI") +
+    ylab("RAP1GAP Cryptic PSI") +
+    ggpubr::theme_pubr() +
+    theme(text = element_text(size = 18)) +
+    scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
+    scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
+
+cowplot::plot_grid(rap1gap_plot,pfkp_plot)
+```
+
+```{r reproduce_4a_in_rap1gap}
+detected_table_for_graph_rap1 = clean_data_table %>%
+  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  filter(rap1gap_cryptic_psi > 0) %>%
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+  filter(junction_reads_stmn2 >= 30) %>%
+  filter((RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated) >=30) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>%
+  filter(disease_tissue == T) %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,rap1gap_cryptic_psi) %>%
+  mutate(log2_rap1gap = log2((rap1gap_cryptic_psi)/(stmn_2_cryptic_psi_leaf))) %>%
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
+  group_by(rs12973192) %>%
+  mutate(n_sample = length(unique(sample))) %>%
+  ungroup() %>%
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
+  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3)) %>% as.data.table()
+
+#get out that person who is discongruous
+
+comparisons = combn(unique(detected_table_for_graph_rap1$detection_name),2,simplify = F)
+
+detected_table_for_graph_rap1 %>%
+  ggplot(aes(y = log2_rap1gap, x = detection_name,fill = detection_name)) +
+  geom_boxplot(show.legend = F) +
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
+  stat_compare_means() +
+  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
+  ylab("Ratio RAP1GAP CE PSI / STMN2 CE PSI") +
+  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 0,size = 1.5)
+
+```

### 2021-07-19 paper_figures.Rmd
Update paper_figures.Rmd

--- 2021-07-14 paper_figures.Rmd
@@
 
+df[is.na(df)] <- 0
+
 p2_patient <- ggplot(df %>% filter(classification != "Neither",
@@
 p2_shsy | p2_patient
-```
+
+# fractional
+
+df2 <- df %>%
+    filter(classification != "Neither") %>%
+    group_by(name) %>%
+    mutate(f = n/sum(n))
+
+ggplot(df2, aes(x = name, y = 100*f, fill = classification)) +
+    geom_bar(stat="identity") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+ggsave("C:/Users/ogw/Google Drive/UCL PhD/Year 2/unc13a/nanopore/nanopore_100.svg",
+       height = 10, width = 12, units = "cm")
+
+ggplot(df2 %>%
+           filter(type == "Patient"), aes(x = name, y = 100*f, fill = classification)) +
+    geom_bar(stat="identity") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("") |
+    ggplot(df2 %>%
+           filter(type != "Patient"), aes(x = name, y = 100*f, fill = classification)) +
+    geom_bar(stat="identity") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+ggsave("C:/Users/ogw/Google Drive/UCL PhD/Year 2/unc13a/nanopore/nanopore_100_separate.svg",
+       height = 10, width = 12, units = "cm")
+
+```

### 2021-07-20 preprocessing_scripts/Targeted_RNA_seq/random_hexamer/analyse_unc_trnaseq_3.R
performs WASP filtering

@@ -129,8 +129,16 @@ for(sample in samples$n){
                       sample, ".fastq.gzAligned.sortedByCoord.out.bam")
 
     message(sample)
-
-    bam_df <- data.frame(scanBam(bam_name))
+    
+    # get the WASP tag from each
+    p4 <- ScanBamParam(tag=c("vW"), what="flag")
+    bam_tag <- data.frame(scanBam(bam_name, param=p4)) %>%
+        mutate(n = 1:n())
+    
+    bam_df <- data.frame(scanBam(bam_name)) %>%
+        mutate(n = 1:n()) %>%
+        inner_join(bam_tag, by ="n") %>%
+        filter(vW == 1) # filter for reads which pass WASP filtering
 
     df2 <- bam_df %>%
         mutate(umi = str_sub(qname, str_length(qname)-7, str_length(qname))) %>%

### 2021-07-20 preprocessing_scripts/Targeted_RNA_seq/specific_RT/analyse_unc_trnaseq_final.R
Update analyse_unc_trnaseq_final.R
with vcf stuff

@@ -23,7 +23,7 @@ library(stringdist)
 # list of "true" UMIs. It does this by finding the minimum string distance of each UMI to 
 # the UMIs currently in the list of "true" UMIs. It requires that for a UMI sequence which 
 # has a string distance of n bases, it is at least 1/(10^n) times the abundance of 
-# the most similar UMI. It also requires that each UMI is at least 0.01x times as abundant
+# the most similar UMI. It also requires that each UMI is at least 0.02x times as abundant
 # as the most abundant, and is present in at least 50 reads.
 #
 # This script differs from the random hexamer script in that it detects whether reads are cryptic
@@ -34,7 +34,7 @@ get_sample <- function(string){
     return(str_split(string,":")[[1]][2])
 }
 
-find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100,
+find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 50,
                              min_n = 50){
     # df should have two columns - "strings" which are the UMIs, and "n" which
     # is the copy number of each
@@ -104,22 +104,35 @@ find_unique_umis <- function(df, distance_multiplier = 20, break_ratio = 100,
 ##### RUN
 
 # a csv containing the names of the samples
+# samples <- read_csv("//data.thecrick.org/lab-ulej/home/users/wilkino/Unc13a/Targeted_RNAseq/barcodes2.csv",
+#                    col_names = "sample_with_barcode")
+
 samples <- read_csv("//data.thecrick.org/lab-ulej/home/users/wilkino/Unc13a/Targeted_RNAseq/barcodes2.csv",
                    col_names = "sample_with_barcode")
 
 # the directory containing the bam files
-dir <- "C:/Users/ogw/Downloads/bams/ultraplex_demux_"
+#dir <- "C:/Users/ogw/Downloads/bams/ultraplex_demux_"
+dir <- "//data.thecrick.org/lab-ulej/home/users/wilkino/Unc13a/Targeted_RNAseq/demultiplexed_unc/ultraplex_demux_"
 
 samples$sample <- unlist(lapply(samples$sample_with_barcode, get_sample))
+
 
 for(sample in samples$sample){
 
     message(sample)
 
     bam_name <- paste0(dir,
-                        sample, ".fastq.gzAligned.out.sam.bam")
[…]

### 2021-07-20 preprocessing_scripts/Targeted_RNA_seq/specific_RT/analyse_unc_trnaseq_final.R
Update analyse_unc_trnaseq_final.R

@@ -131,8 +131,8 @@ for(sample in samples$sample){
     
     bam_df <- data.frame(scanBam(bam_name)) %>%
         mutate(n = 1:n()) %>%
-        inner_join(bam_tag, by ="n") #%>%
-        #filter(vW == 1) # filter for reads which pass WASP filtering
+        inner_join(bam_tag, by ="n") %>%
+        filter(vW == 1) # filter for reads which pass WASP filtering
 
     df2 <- bam_df %>%
         mutate(umi = str_sub(qname, str_length(qname)-7, str_length(qname))) %>%
@@ -234,4 +234,4 @@ just_C_or_G <- just_C_or_G_tmp %>%
 
 # Write final data out
 
-write_csv(just_C_or_G, "//data.thecrick.org/lab-ulej/home/users/wilkino/Unc13a/Targeted_RNAseq/demultiplexed_unc/for_plot_without_vcf.csv")
+write_csv(just_C_or_G, "//data.thecrick.org/lab-ulej/home/users/wilkino/Unc13a/Targeted_RNAseq/demultiplexed_unc/for_plot_vcf.csv")

### 2021-07-21 paper_figures.Rmd
Update paper_figures.Rmd

--- 2021-07-19 paper_figures.Rmd
@@
 
-## Systematic differences in  (read length) between cases and controls
-
-```{r sequences and case/controls}
+## Read length affects detection probability in cases
+
+```{r readlength_detection_in_cases}
 # library size 
@@
     p_value = chisq.test(comparison_table) %>%  broom::tidy() %>% pull(p.value)
-    browser()
+
     print(p_value)
@@
 for(t in 1:length(lib_comps)){
-    if(grepl("Cord",lib_comps[t]) | lib_comps[t] == "Motor_Cortex"){
-        groups = c("Control","ALS-TDP","ALS \n non-TDP")
-    }else if(lib_comps[t] %in% c("Frontal_Cortex","Temporal_Cortex")){
-        groups = c("Control","FTLD-TDP","FTLD \n non-TDP")
-    }else{
+
         groups = c("ALS-TDP", "Control", "ALS \n non-TDP", "FTLD-TDP", "FTLD \n non-TDP")
-    }
 
@@
                                        c("ALS \n non-TDP","ALS-TDP"))) +
-        ggpubr::stat_compare_means(label = "p.signif", 
+        ggpubr::stat_compare_means( label = "p.signif",
                                    label.x.npc = "center", 
@@
               axis.text.x = element_blank(),
-              legend.position = "none")
+              legend.position = "none") +
+      scale_color_manual(values = force_colors) + 
+      scale_fill_manual(values = force_colors)
 
@@
 
-    if(grepl("Cord",lib_comps[t]) | lib_comps[t] == "Motor_Cortex"){
-        groups = c("Control","ALS-TDP","ALS \n non-TDP")
-    }else if(lib_comps[t] %in% c("Frontal_Cortex","Temporal_Cortex")){
-        groups = c("Control","FTLD-TDP","FTLD \n non-TDP")
-    }else{
         groups = c("ALS-TDP", "Control", "ALS \n non-TDP", "FTLD-TDP", "FTLD \n non-TDP")
-    }
 
@@
   geom_boxplot(outlier.color = NA) + 
-  facet_wrap(~plot_name+tissue,scales = 'free') +
+  facet_wrap(~plot_name+tissue_clean,scales = 'free') +
   ggpubr::stat_compare_means(label = "p.signif") +
@@
 ```{r onlyshort}
+
+star_maker = function (p.value) 
+{
+    unclass(symnum(p.value, corr = FALSE, na = FALSE, cutpoints = c(0, 0.0001,
+        0.001, 0.01, 0.05, 0.1, 1), symbols = c("****","***", "**", 
+        "*", ".", " ")))
+}
 only_short_plotter = boxplot_table %>%
@@
 
-comps = only_short_plotter %>% pull(detection_name) %>%
+only_short_plotter %>% 
+  dplyr::select(sample,tissue_clean,inclusion_reads,disease_group2) %>% 
+  mutate(disease_group2 = gsub("\n", "",disease_group2)) %>% 
   unique() %>% 
-  combn(.,2,simplify = F)
+  group_by(tissue_clean) %>% 
+  nest() %>%
+  mutate(wilcox_res = map(data, ~pairwise.wilcox.test(.x$inclusion_reads,.x$disease_group2) %>% broom::tidy())) %>%
+  dplyr::select(-data) %>% 
+  unnest() %>% 
+  mutate(star_plot = star_maker(p.value)) %>% fwrite("~/Desktop/stars_for_weaverly_fig3b.csv")
 
@@
     theme(text = element_text(size = 18))  +
-    theme(legend.position = "bottom")
+    theme(legend.position = "bottom") + 
+    stat_compare_means(comparisons = comps)
 ```
@@
   ylab("CE PSI")  +
-  theme(legend.position = 'none')
+  theme(legend.position = 'none') + 
+  stat_compare_means(label = 'p.signif')
 
@@
   mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
-  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3)) %>% as.data.table()
+ as.data.table()
 
@@
 
-detected_table_for_graph_rap1 %>%
+rap1_fold_plot = detected_table_for_graph_rap1 %>%
   ggplot(aes(y = log2_rap1gap, x = detection_name,fill = detection_name)) +
@@
   geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
-  stat_compare_means() +
   stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
-  ylab("Ratio RAP1GAP CE PSI / STMN2 CE PSI") +
-  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
+  ylab(expression(paste("Lo", g[2], " fold CE PSI / STMN2 CE PSI "))) +
   xlab("UNC13A rs12973192 Genotype") +
@@
   theme(text = element_text(size = 24)) +
-  geom_hline(yintercept = 0,size = 1.5)
-
-```
+  geom_hline(yintercept = 0,size = 1.5) + 
+  ggtitle("RAP1GAP")
+
+```
+
+```{r reproduce_4a_in_pfkp}
+detected_table_for_graph_pfkp = clean_data_table %>%
+  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  filter(pfkp_cryptic_psi > 0) %>%
+  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
+  filter(junction_reads_stmn2 >= 30) %>%
+    filter((PFKP_novelacceptor + PFKP_noveldonor + PFKP_annotated) >=30) %>% 
+  filter(grepl("Cortex",tissue_clean)) %>%
+  filter(disease_tissue == T) %>%
+  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
+  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,pfkp_cryptic_psi) %>%
+  mutate(log2_pfkp = log2((pfkp_cryptic_psi)/(stmn_2_cryptic_psi_leaf))) %>%
+  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
+  group_by(rs12973192) %>%
+  mutate(n_sample = length(unique(sample))) %>%
+  ungroup() %>%
+  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
+  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3)) %>% as.data.table()
+
+#get out that person who is discongruous
+
+comparisons = combn(unique(detected_table_for_graph_pfkp$detection_name),2,simplify = F)
+
+pfkp_fold_plot = detected_table_for_graph_pfkp %>%
+  ggplot(aes(y = log2_pfkp, x = detection_name,fill = detection_name)) +
+  geom_boxplot(show.legend = F) +
+  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
+  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
+  xlab("UNC13A rs12973192 Genotype") +
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggpubr::theme_pubr() +
+  theme(text = element_text(size = 24)) +
+  geom_hline(yintercept = 0,size = 1.5) + 
+  ggtitle("PFKP") + 
+  ylab(element_blank())
+
+cowplot::plot_grid(rap1_fold_plot,pfkp_fold_plot)
+
+```

### 2021-07-21 paper_figures.Rmd

@@
 
-#get out that person who is discongruous
-
 comparisons = combn(unique(detected_table_for_graph_pfkp$detection_name),2,simplify = F)
@@
 cowplot::plot_grid(rap1_fold_plot,pfkp_fold_plot)
-
-```
+```
+# Nanopore
+
+Read in pre-processed data (see separate script) and plot
+```{r warning=FALSE, message=FALSE}
+df <- read_csv(paste0(here(),
+                      "/data/nanopore/combined_nanopore_dataframe.csv"))
+
+df[is.na(df)] <- 0
+
+p2_patient <- ggplot(df %>% filter(classification != "Neither",
+                                   type=="Patient"), aes(x = name, y = 100*fraction, fill = class2)) +
+    geom_bar(stat="identity", position="dodge") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("Patient RNA") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+p2_shsy <- ggplot(df %>% filter(classification != "Neither",
+                                type=="SHSY5Y"), aes(x = name, y = 100*fraction, fill = class2)) +
+    geom_bar(stat="identity", position="dodge") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("SH-SY5Y") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+p2_shsy | p2_patient
+
+# without neither removed
+
+p2_patient <- ggplot(df %>% filter(classification != "",
+                                   type=="Patient"), aes(x = name, y = 100*fraction, fill = class2,
+                                                         colour = class2)) +
+    geom_bar(stat="identity", position="dodge") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("Patient RNA") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+p2_shsy <- ggplot(df %>% filter(classification != "",
+                                type=="SHSY5Y"), aes(x = name, y = 100*fraction, fill = class2,
+                                                     colour = class2)) +
+    geom_bar(stat="identity", position="dodge") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("SH-SY5Y") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("") 
+p2_shsy | p2_patient
+
+# fractional
+
+df2 <- df %>%
+    filter(classification != "Neither") %>%
+    group_by(name) %>%
+    mutate(f = n/sum(n))
+
+ggplot(df2, aes(x = name, y = 100*f, fill = classification)) +
+    geom_bar(stat="identity") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+ggplot(df2 %>%
+           filter(type == "Patient"), aes(x = name, y = 100*f, fill = classification)) +
+    geom_bar(stat="identity") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("") |
+    ggplot(df2 %>%
+           filter(type != "Patient"), aes(x = name, y = 100*f, fill = classification)) +
+    geom_bar(stat="identity") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("")
+
+```

### 2021-08-05 paper_figures.Rmd
Update paper_figures.Rmd
added plot for 2F and S2C

@@
 ```
-# NMD experiment on SH-SY5Y cells
-
-```{r}
-
-fbp_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
-fbp_sy$condition <- as.factor(fbp_sy$condition)
-fbp_sy$condition <- factor(fbp_sy$condition,levels(fbp_sy$condition)[c(2,1)])
-
-stattest_sy <- fbp_sy %>%
-    group_by(gene) %>%
-    t_test(value ~ 1, mu = 1) %>%
+# NMD experiment on SH-SY5Y cells - CHX
+
+```{r}
+
+for_bar_plot_input <- fread(file.path(here::here(),"data","SY5Y_Tidy_4.csv"))
+for_bar_plot <- data.frame(for_bar_plot_input)
+for_bar_plot$Treatment <- factor(for_bar_plot$Treatment, levels = c("DMSO", "CHX"))
+
+shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "HNRNPL")$Value)
+shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "STMN2")$Value)
+shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "UNC13A")$Value)
+shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "UNC13B")$Value)
+
+stattest <- for_bar_plot %>%
+  group_by(Gene) %>%
+  pairwise_t_test(Value ~ Treatment, ref.group = "DMSO", alternative = "greater") %>%
   add_significance() %>% 
-    add_xy_position(x="gene")
-stattest_sy
-
-nmd_plot_sy <- ggbarplot(fbp_sy,
-                         x = 'gene',
-                         y = 'value',
+  add_xy_position(x="Treatment")
+stattest
+
+nmd_plot_chx = ggbarplot(for_bar_plot,
+                     x = 'Gene',
                      add = c("mean_se","jitter"),
-                         fill = 'condition',
-                         color = 'condition',
+                     y = 'Value',
+                     color = 'Treatment',
                      position = position_dodge(0.8),
                      dot.size = 10) +
-    scale_y_continuous(labels = scales::percent) +
-    ggpubr::theme_pubr() +
-    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) +
-    scale_color_manual(values = c("#1C2617","#262114")) +
-    ylab("Percent of transcript after CHX treatment") +
-    xlab("") +
-    stat_pvalue_manual(stattest_sy, label="p.signif", y.position = 9)
-nmd_plot_sy
-
-```
-
-```{r}
-
-fbp_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
-fbp_dz$condition <- as.factor(fbp_dz$condition)
-fbp_dz$condition <- factor(fbp_dz$condition,levels(fbp_dz$condition)[c(2,1)])
-
-stattest_dz <- fbp_dz %>%
-    group_by(gene) %>%
-    t_test(value ~ 1, mu = 1) %>%
+  scale_y_continuous() + 
+  ggpubr::theme_pubr() + stat_pvalue_manual(stattest, label="p.signif" , x = "Gene")
+plot(nmd_plot_chx)
+
+```
+
+# NMD experiment on SH-SY5Y cells - sTDP43+sUPF1
+
+```{r}
+
+for_bar_plot_sy_input <- fread(file.path(here::here(),"data","results_upf1.csv"))
+
+for_bar_plot_sy_df <- data.frame(for_bar_plot_sy_input)
+for_bar_plot_sy_df$Condition <- as.factor(for_bar_plot_sy_df$Condition)
+
+#assess TDP-43 and UPF1 KD
+for_bar_plot_sz <- for_bar_plot_sy_df %>%
+  filter(Gene == "TDP43" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1" | Condition == "Non-treated"))
+
+shapiro.test(filter(for_bar_plot_sz, Condition == "siTDP43+siCTRL" & Gene == "TDP43")$Value)
+shapiro.test(filter(for_bar_plot_sz, Condition == "siTDP43+siUPF1" & Gene == "TDP43")$Value)
+
+stattest_sz <- for_bar_plot_sz %>%
+  pairwise_t_test(Value ~ Condition, ref.group = "Non-treated", alternative = "less") %>%
   add_significance() %>% 
-    add_xy_position(x="gene")
-stattest_dz
-
-nmd_plot_dz <- ggbarplot(fbp_dz,
-                         x = 'gene',
-                         y = 'value',
+  add_xy_position(x="Condition")
+stattest_sz
+
+for_bar_plot_sx <- for_bar_plot_sy_df %>%
+  filter(Gene == "UPF1" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1"))
+
+shapiro.test(filter(for_bar_plot_sx, Condition == "siTDP43+siUPF1" & Gene == "UPF1")$Value)
+
+stattest_sx <- for_bar_plot_sx %>%
+  pairwise_t_test(Value ~ Condition, ref.group = "siTDP43+siCTRL", alternative = "less") %>%
+  add_significance() %>% 
+  add_xy_position(x="Condition")
+stattest_sx
+
+for_bar_plot_szx <- for_bar_plot_sy_df %>%
+  filter(Gene == "TDP43" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1" | Condition == "Non-treated")
+         | Gene == "UPF1" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1"))
+
+#fix position of stats for plot
+stattest_sx$xmin[1] <- 1.8
+stattest_sx$xmax[1] <- 2.2
+stattest_sx$y.position[1] <- 1.3
+stattest_sz$xmin[1] <- 0.75
+stattest_sz$xmin[2] <- 0.75
+stattest_sz$xmax[1] <- 1
+stattest_sz$xmax[2] <- 1.25
+stattest_sz$y.position[1] <- 1.2
+stattest_sz$y.position[2] <- 1.3
+
+nmd_plotz = ggbarplot(for_bar_plot_szx,
+                     x = 'Gene',
                      add = c("mean_se","jitter"),
-                         fill = 'condition',
-                         color = 'condition',
+                     y = 'Value',
+                     color = 'Condition',
                      position = position_dodge(0.8),
                      dot.size = 10) +
-    scale_y_continuous(labels = scales::percent) +
-    ggpubr::theme_pubr() +
-    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) +
-    scale_color_manual(values = c("#1C2617","#262114")) +
-    ylab("Percent of transcript after CHX treatment") +
-    xlab("") +
-    stat_pvalue_manual(stattest_dz, label="p.signif", y.position = 26)
-nmd_plot_dz
+  scale_y_continuous() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") + 
+  ggpubr::theme_pubr() + stat_pvalue_manual(stattest_sz) + stat_pvalue_manual(stattest_sx)
+plot(nmd_plotz)
+
+#assess rescue of NMD-sensitive transcripts after UPF1 inhibition
+for_bar_plot_sy <- for_bar_plot_sy_df %>%
+  filter(Gene == "hnRNPL" | Gene == "STMN2" | Gene == "UNC13A" | Gene == "UNC13B")
+
+shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "hnRNPL")$Value)
+shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "STMN2")$Value)
+shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "UNC13A")$Value)
+shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "UNC13B")$Value)
+
+stattest_sy <- for_bar_plot_sy %>%
+  group_by(Gene) %>%
+  pairwise_t_test(Value ~ Condition, ref.group = "siTDP43+siCTRL", alternative = "greater") %>%
+  add_significance() %>% 
+  add_xy_position(x="Condition")
+stattest_sy
+
+nmd_plot = ggbarplot(for_bar_plot_sy, 
+                     x = 'Gene', 
+                     add = c("mean_se","jitter"), 
+                     y = 'Value',
+                     color = 'Condition', 
+                     position = position_dodge(0.8), 
+                     dot.size = 10) +
+  scale_y_continuous() + 
+  ggpubr::theme_pubr() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") +
+  stat_pvalue_manual(stattest_sy, label="p.signif", x = "Gene")
+plot(nmd_plot)
 

### 2021-08-13 paper_figures.Rmd
Add NYGC IR plots

--- 2021-07-21 paper_figures.Rmd
@@
     geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)
-
 ```
@@
 rel_rna_cryptic_amount %>%
+  left_join(text_table[,.(plot_name,source)]) %>% 
+  unique() %>% 
+  mutate(plot_name = fct_relevel(plot_name,"iPSC MN","SH-SY5Y","I3 Neurons","SK-N-DZ_a")) %>% 
     ggbarplot(,
-              x = "source",
+              x = "plot_name",
               add = c("mean_se","jitter"),
@@
     scale_y_continuous(labels = scales::percent,expand = expansion(mult = c(0, .1))) +
-    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
-    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)
+    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = plot_name, y= 1.35,label = padj_plot),size = 8) +
+    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = plot_name, y= 1.2,label = log2FoldChange_plot),size = 8)
 
@@
 ```{r}
+force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
+names(force_colors) = c("Control","ALS \n non-TDP","ALS-TDP","FTLD \n non-TDP","FTLD-TDP")
 clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
@@
 
-## Read length affects detection probability in cases
-
-```{r readlength_detection_in_cases}
-# library size 
-#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
-case_control_sequencing <- clean_data_table %>%
-  filter(tissue_clean %in% c("Frontal_Cortex", 
-                             "Lumbar_Spinal_Cord", 
-                             "Cervical_Spinal_Cord", 
-                             "Motor_Cortex", 
-                             "Temporal_Cortex",
-                             "Cerebellum")) %>% 
-  filter(sequencing_platform != "")
-
-quick_chiq = function(tbl){
-
-    if(length(unique(tbl$sequencing_platform)) < 2){
-      return(NA)
-    }
-    comparison_table = table(disease = tbl$disease_group2, 
-                             sequencing_platform = tbl$sequencing_platform)
-    print(comparison_table)
-
-    p_value = chisq.test(comparison_table) %>%  broom::tidy() %>% pull(p.value)
-
-    print(p_value)
-
-    return(p_value)
-}
-
-case_control_sequencing %>% 
-  group_by(tissue_clean) %>% 
-  nest()  %>% 
-  mutate(chiq_pvalue = map_dbl(data, quick_chiq))
-
-unc13a_data_fisher$data = NULL
-
-unc13a_sequencing_platform_pct = unc13a_sequencing_platform %>%
-  group_by(disease_group2,tissue_clean, sequencing_platform, unc13a_detected) %>% tally() %>%
-  spread(key = unc13a_detected, value = n,fill = 0) %>%
-  mutate(detection = `+` / ( `-` + `+`) ) %>% 
-  mutate(total =  ( `-` + `+`) ) %>% 
-  mutate(plot_name = glue::glue("{sequencing_platform} \n  ({total})")) %>% 
-  left_join(unc13a_data_fisher)
-
-unc13a_sequencing_platform_pct %>%
-  ggplot(aes(x = plot_name, y = detection )) + 
-  geom_col(fill = "firebrick") +
-  labs(title = "UNC13A CE detection and sequencing platform", 
-       y = "UNC13A CE detected",x = element_blank(), subtitle = "Fisher exact test") + 
-  facet_wrap(~disease_group2 + tissue_clean,scales = 'free_x') +
-  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
-  geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(fisher_pvalue,digits = 2) ) )) +
-  theme(legend.text = element_text(size = 15)) +
-  ggpubr::theme_pubr()
- # scale_y_continuous(expand = c(0,0) ) 
-```
-
 ## Effect of sequencing machine (read length) on CE discovery
@@
   theme(legend.text = element_text(size = 15)) +
-  ggpubr::theme_pubr()
+  ggpubr::theme_pubr() +
+  scale_fill_manual(values = ("#b3251f"))
  # scale_y_continuous(expand = c(0,0) ) 
 ```
-## Effect of library size on CE discovery
+
+## Sequencing platform and disease type
+
+Were there systemic differences in sequence platform between ALS and FTD?
+
+```{r platform_compare}
+tmp = clean_data_table %>%
+  filter(tissue_clean %in% c("Frontal_Cortex", 
+                             "Lumbar_Spinal_Cord", 
+                             "Cervical_Spinal_Cord", 
+                             "Motor_Cortex", 
+                             "Temporal_Cortex")) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+  mutate(tissue = as.character(tissue)) %>%
+  filter(sequencing_platform != "") %>% 
+  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>% 
+  dplyr::select(tissue_clean,disease_group2,sequencing_platform) %>% 
+  group_by(tissue_clean) %>% 
+  nest()
+
+tmp %>% 
+  mutate(chisq_res = map(data, ~ .x %>% table() %>% chisq.test() %>% pluck('residuals')))
+```
+
+## Library size by disease status and tissue
 
@@
 # library size
-force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
-names(force_colors) = c("Control","ALS \n non-TDP","ALS-TDP","FTLD \n non-TDP","FTLD-TDP")
 
@@
 
-## Effect of RIN and UNC13A discovery
-
-```{r rin discovery}
+## Effect of library depth on CE discovery
+
+```{r unc detection by library depth}
+
+unc13a_library_depth <- clean_data_table %>%
+  filter(tissue_clean %in% c("Frontal_Cortex", 
+                             "Lumbar_Spinal_Cord", 
+                             "Cervical_Spinal_Cord", 
+                             "Motor_Cortex", 
+                             "Temporal_Cortex")) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP"))
+
+unc13a_library_depth %>%
+  group_by(disease_group2,tissue_clean) %>% 
+  add_count() %>% 
+  mutate(plot_name = glue::glue("{tissue_clean} \n {n}")) %>% 
+  ggplot(aes(x = unc13a_detected, y = log10(library_size))) + 
+  geom_boxplot(aes(fill = unc13a_detected), colour = "black", outlier.colour = NA) + 
+  facet_wrap(~disease_group2+ plot_name) + 
+  geom_jitter(height = 0,size = 2,alpha = 0.3) + 
+  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox", label.y.npc = 0.95) + 
+  ggpubr::theme_pubr() +
+  xlab("UNC13A CE Detected") + 
+  scale_fill_manual(values = c("#EBC3B9","#b3251f")) + 
+  theme(legend.position = 'none')
+```
+
+## RIN between cases and controls
+
+```{r rin_cases_controls}
 force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
@@
 cowplot::plot_grid(plotlist = plot_list,ncol = 2)    
+```
+
+## RIN and detection in cases
+```{r rin_detection_cases}
+
+unc13a_rin <- clean_data_table %>%
+  filter(tissue_clean %in% c("Frontal_Cortex", 
+                             "Lumbar_Spinal_Cord", 
+                             "Cervical_Spinal_Cord", 
+                             "Motor_Cortex", 
+                             "Temporal_Cortex")) %>% 
+  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>% 
+  filter(rin != "")
+
+unc13a_rin %>%
+  group_by(disease_group2,tissue_clean) %>% 
+  add_count() %>% 
+  mutate(plot_name = glue::glue("{tissue_clean} \n {n}")) %>% 
+  ggplot(aes(x = unc13a_detected, y = rin)) + 
+  geom_boxplot(aes(fill = unc13a_detected), colour = "black", outlier.colour = NA) + 
+  facet_wrap(~disease_group2+ plot_name) + 
+  ylab("RIN") +
+  geom_jitter(height = 0,size = 2,alpha = 0.3) + 
+  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox", label.y.npc = 0.95) + 
+  ggpubr::theme_pubr() +
+  xlab("UNC13A CE Detected") +
+  scale_fill_manual(values = c("#EBC3B9","#b3251f")) + 
+  theme(legend.position = 'none')
 ```
@@
                                disease_tissue)], by = 'sample') %>%
-  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
-  filter(disease_tissue == T) %>% 
-  filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
+  filter(tissue_clean %in% c("Frontal_Cortex", 
+                             "Lumbar_Spinal_Cord", 
+                             "Cervical_Spinal_Cord", 
+                             "Motor_Cortex", 
+                             "Temporal_Cortex")) %>% 
   mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
   mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
-  add_count(disease_group2,tissue_clean) %>% 
+  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>% 
+  group_by(disease_group2,tissue_clean) %>% 
+  add_count() %>% 
   mutate(n = n / 5) %>% 
-  mutate(plot_name = glue::glue("{disease_group2} \n {n}")) %>% 
+  mutate(plot_name = glue::glue("{tissue_clean} \n {n}")) %>% 
   filter(!is.na(inclusion_reads)) %>% 
@@
   ggplot(aes(x = cell, y = deconv, fill = unc13a_detected)) + 
-  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
+  # geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
   geom_boxplot(outlier.color = NA) + 
-  facet_wrap(~plot_name+tissue_clean,scales = 'free') +
-  ggpubr::stat_compare_means(label = "p.signif") +
-  ggpubr::stat_compare_means(label = "p.format",vjust = 1.2) +
-  labs(title = "Cell type decomposition and UNC13A CE detection ",subtitle = "Wilcox test") + 
-  ggpubr::theme_pubr()
+  facet_wrap(~disease_group2+plot_name,scales = 'free') +
+  ggpubr::stat_compare_means(label = "p.format",vjust = 1.3) +
+  ggpubr::theme_pubr() +
+  scale_fill_manual(values = c("#EBC3B9","#b3251f")) + 
+  theme(legend.position = 'none') +
+  xlab(element_blank())
 
@@
 
+## UNC13A and UNC13B Intron Retention in NYGC
+
+```{r unc13a_ir_in_newyork}
+stat_test_a = clean_data_table %>% 
+    filter(tissue_clean %in% c("Frontal_Cortex", 
+                               "Lumbar_Spinal_Cord", 
+                               "Cervical_Spinal_Cord", 
+                               "Motor_Cortex", 
+                               "Temporal_Cortex")) %>% 
+    unique() %>% 
+    group_by(tissue_clean) %>% 
+    wilcox_test(IRratioUNC13A ~ disease_group2) %>%
+    add_significance("p")
+
+stat_test_b = clean_data_table %>% 
+    filter(tissue_clean %in% c("Frontal_Cortex", 
+                               "Lumbar_Spinal_Cord", 
+                               "Cervical_Spinal_Cord", 
+                               "Motor_Cortex", 
+                               "Temporal_Cortex")) %>% 
+    unique() %>% 
+    group_by(tissue_clean) %>% 
+    wilcox_test(IRratioUNC13B ~ disease_group2) %>%
+    add_significance("p")
+
+clean_data_table %>% 
+    filter(tissue_clean %in% c("Frontal_Cortex", 
+                               "Lumbar_Spinal_Cord", 
+                               "Cervical_Spinal_Cord", 
+                               "Motor_Cortex", 
+                               "Temporal_Cortex")) %>% 
+    unique() %>% 
+    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
+    ggplot(aes(x = disease_group2,
+               y = IRratioUNC13A)) +
+    geom_boxplot(aes(fill = disease_group2)) + 
+    facet_wrap(~tissue_clean,scales = 'free_x') +
+    stat_pvalue_manual(stat_test_a %>% filter(group1 == "Control" & group2 == "FTLD-TDP"), 
+                       label = "p.signif",y.position = 0.98) +
+    stat_pvalue_manual(
+                       stat_test_a %>% filter(group1 == "ALS-TDP" & group2 == "Control"), 
+                       label = "p.signif",y.position = 0.88) +
+    ylab("UNC13A IRratio") +
+  xlab(element_blank())+
+  scale_fill_manual(values = force_colors) +
+  ggpubr::theme_pubr() + 
+  theme(legend.position = 'none')
+
+```
+
+```{r unc13b_ir_in_newyork}
+clean_data_table %>% 
+    filter(tissue_clean %in% c("Frontal_Cortex", 
+                               "Lumbar_Spinal_Cord", 
+                               "Cervical_Spinal_Cord", 
+                               "Motor_Cortex", 
+                               "Temporal_Cortex")) %>% 
+    unique() %>% 
+    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
+    ggplot(aes(x = disease_group2,
+               y = IRratioUNC13B)) +
+    geom_boxplot(aes(fill = disease_group2)) + 
+    facet_wrap(~tissue_clean,scales = 'free_x') +
+    stat_pvalue_manual(stat_test_b %>% filter(group1 == "Control" & group2 == "FTLD-TDP"), 
+                       label = "p.signif",y.position = 0.98) +
+    stat_pvalue_manual(
+        stat_test_b %>% filter(group1 == "ALS-TDP" & group2 == "Control"), 
+        label = "p.signif",y.position = 0.88) +
+    ylab("UNC13B IRratio") +
+    scale_fill_manual(values = force_colors) +
+    ggpubr::theme_pubr() + 
+    theme(legend.position = 'none')
+
+```
+
 ## Correlation STMN2 Cryptic PSI and UNC13A TPM
@@
   ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
-  geom_point() +
+  geom_point(pch = 21, size = 3) +
   stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
@@
 ```
+## Detection by genotype
+```{r}
+
+clean_data_table %>%
+  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
+  mutate(snp_status = case_when(rs12973192 == "C/C" ~ "healthy/healthy",
+                                rs12973192 == "C/G" ~ "healthy/risk",
+                                rs12973192 == "G/G" ~ "risk/risk")) %>% 
+  filter(disease_tissue == T) %>%
+  filter(junction_reads_stmn2 >= 30) %>%
+  filter(junction_reads_unc13a >= 30) %>%
+
+```
 
@@
   xlab("N individuals") +
-  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))
+  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
+  ggtitle(label = "ALS/FTD-TDP patients",subtitle =  glue::glue("Chi-squared Test: p-value {round(over_p,2)}"))
 
@@
 ```
+
 # Nanopore
@@
 
-p2_shsy <- ggplot(df %>% filter(classification != "",
-                                type=="SHSY5Y"), aes(x = name, y = 100*fraction, fill = class2,
+ggplot(df %>% filter(classification != "",
+                                type=="SHSY5Y"), aes(x = name, y = fraction, fill = class2,
                                                      colour = class2)) +
@@
     xlab("") +
-    ggeasy::easy_add_legend_title("") 
+    ggeasy::easy_add_legend_title("") +
+    scale_y_continuous(labels = scales::percent_format(accuracy = 1L))
+
 p2_shsy | p2_patient
@@
     xlab("") +
-    ggeasy::easy_add_legend_title("") |
+    ggeasy::easy_add_legend_title("")
+
     ggplot(df2 %>%

### 2021-09-01 paper_figures.Rmd
Adding in the raw data from different guides - salmon data

@@
 rel_rna_cryptic_amount %>%
+  left_join(text_table[,.(plot_name,source)]) %>% 
+  unique() %>% 
+  mutate(plot_name = fct_relevel(plot_name,"iPSC MN","SH-SY5Y","I3 Neurons","SK-N-DZ_a")) %>% 
     ggbarplot(,
-              x = "source",
+              x = "plot_name",
               add = c("mean_se","jitter"),
-              y = "UNC13A",
+              y = "UNC13B",
               fill = 'condition',
@@
           axis.text.y = element_text(size = 28)) +
-    scale_y_continuous(labels = scales::percent) +
-    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
-    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)
+    scale_y_continuous(labels = scales::percent,expand = expansion(mult = c(0, .1))) +
+    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = plot_name, y= 1.35,label = padj_plot),size = 8) +
+    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = plot_name, y= 1.2,label = log2FoldChange_plot),size = 8)
 
@@
                y = IRratioUNC13A)) +
-    geom_boxplot(aes(fill = disease_group2)) + 
+    geom_boxplot(aes(fill = disease_group2),outlier.shape = NA) + 
+   geom_jitter(aes(fill = disease_group2),
+               pch = 21,height = 0,
+               width = 0.2,size = 3,alpha = 0.8) +
     facet_wrap(~tissue_clean,scales = 'free_x') +
@@
                y = IRratioUNC13B)) +
-    geom_boxplot(aes(fill = disease_group2)) + 
-    facet_wrap(~tissue_clean,scales = 'free_x') +
+    geom_boxplot(aes(fill = disease_group2),outlier.shape = NA) + 
+   geom_jitter(aes(fill = disease_group2),
+               pch = 21,height = 0,
+               width = 0.2,size = 3,alpha = 0.8) +    facet_wrap(~tissue_clean,scales = 'free_x') +
     stat_pvalue_manual(stat_test_b %>% filter(group1 == "Control" & group2 == "FTLD-TDP"), 
@@
     ylab("UNC13B IRratio") +
+  xlab(element_blank()) +
     scale_fill_manual(values = force_colors) +
@@
 
+# UNC13B isoform expression in-vitro and across the NYGC
+
+```{r unc13b_isoform_cell_lines}
+iso_colors = c("#1972A9", "#28A919", "#F49F0B")
+names(iso_colors) = c("UNC13B-207","UNC13B-208","UNC13B-210")
+#207 -- 1972A9 - normal
+#208 -- 28A919 - NMD
+#210 -- F49F0B - brain
+t_type = fread(file.path(here::here(),"data","transcripts-Summary-Homo_sapiens_Gene_Summary_ENSG00000198722.csv"))
+
+cell_lines = fread(file.path(here::here(),"data","unc13b_salmon_cell_lines.csv"))
+cell_lines %>% 
+    mutate(sample = gsub(".Aligned.sorted.out","",sample)) %>% 
+    left_join(rel_rna_cryptic_amount[,.(condition,plot_name,sample)]) %>% 
+    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
+    filter(Name.y %in% c("UNC13B-207",
+                         "UNC13B-208",
+                         "UNC13B-210")) %>%
+    ggplot(aes(x =condition, y = TPM,fill = Name.y)) + 
+    geom_boxplot() +
+   geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
+    # facet_wrap(~plot_name) +
+  ggpubr::theme_pubr() + 
+  labs(fill="",x="") + 
+  scale_fill_manual(values = iso_colors)
+
+```
+
+```{r unc13b_isoform_nygc}
+
+unc13b_expression = fread(file.path(here::here(),"data","unc13b_transcript_counts.csv"))
+
+clean_data_table %>% 
+    left_join(unc13b_expression) %>% 
+    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
+    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
+    select(sample,disease_group2,tissue_clean,unc13a_detected,TPM,Name) %>% 
+    filter(!is.na(Name)) %>% 
+    unique() %>% 
+    filter(tissue_clean %in% c("Frontal_Cortex", 
+                               "Lumbar_Spinal_Cord", 
+                               "Cervical_Spinal_Cord", 
+                               "Motor_Cortex", 
+                               "Temporal_Cortex")) %>% 
+    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
+        filter(Name.y %in% c("UNC13B-207",
+                         "UNC13B-208",
+                         "UNC13B-210")) %>%
+    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
+    ggplot(aes(x = disease_group2, y = TPM, fill = Name.y)) + 
+    geom_boxplot() +
+    geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
+    facet_wrap(~tissue_clean,scales = 'free') + 
+    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) + 
+    scale_y_continuous(label=comma) +
+    ggpubr::theme_pubr() + 
+    labs(fill="",x="") +
+    scale_fill_manual(values = iso_colors)
+
+```
+
 # Nanopore
@@
 ```{r warning=FALSE, message=FALSE}
-df <- read_csv(paste0(here(),
+df <- read_csv(paste0(here::here(),
                       "/data/nanopore/combined_nanopore_dataframe.csv"))

### 2021-09-01 cleaning_wtc11_vcf.R

@@ -1,64 +1,0 @@
-library(data.table)
-library(VariantAnnotation)
-library(tidyverse)
-#get the genomic coords of the revelvant ALS/FTD genes 
-als_ftd_risk_genes = c("AARS2", "ALS2", "ALS3", "ALS7", "ANG", "APBB2", "APOE", "APP", 
-                       "ATP13A2", "ATXN2", "C19orf12", "C6orf10", "C9orf72", "CAPZA1", 
-                       "CCNF", "CDH22", "CFAP410", "CHCHD10", "CHMP2B", "CHRNA3,4,B4", 
-                       "CHRNA4", "CNTN6", "CRYM", "CSF1R", "CTSF", "CX3CR1", "DAO", 
-                       "DCTN1", "ELP3", "EPB41L1", "ERBB4", "EWSR1", "FIG4", "FLNC", 
-                       "FUS", "GFRA2", "GLE1", "GRN", "GSN", "HNRNPA1", "HNRNPA2B1", 
-                       "ITM2B", "KIF5A", "LMNB1", "LOC101929163", "LRRK2", "MAPT", "MATR3", 
-                       "MPO", "NEFH", "NEK1", "NIPA1", "NOTCH3", "NPC2", "OPTN", "PFN1", 
-                       "PLD3", "PON1-3", "PRKAR1B", "PRNP", "PRPH", "PSEN1", "PSEN2", 
-                       "RNPLA6", "SETX", "SIGMAR1", "SOD1", "SORT1", "SPG11", "SQSTM1", 
-                       "SRR", "SS18L1", "TAF15", "TARDBP", "TBK1", "TIA1", "TMEM106B", 
-                       "TNIP1", "TREM2", "TRPM2", "TSC1", "TUBA4A", "UBQLN2", "UNC13A", 
-                       "VAPB", "VCP")
-
-gene_locations = annotables::grch38 %>% filter(symbol %in% als_ftd_risk_genes) %>% dplyr::select(symbol,chr,start,end) %>% filter(
-    chr %in% as.character(1:22)
-) %>% mutate(chr = paste0("chr",chr))
-
-wtc11_vcf = "/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/wtc11_vcf.gz"
-
-gene_gr = makeGRangesFromDataFrame(gene_locations,keep.extra.columns = T)
-
-tab <- TabixFile(wtc11_vcf)
[…]

### 2021-09-10 paper_figures.Rmd
more figs

@@
 title: "Figures for Paper"
-author: "Anna-Leigh Brown + Matteo Zanovello + Oscar Wilkins"
+author: "Anna-Leigh Brown + Matteo Zanovello + Oscar Wilkins + Matthew Keuss"
 date: "06/10/2020"
@@
                      color = 'Condition',
+                     fill = "Condition",
                      position = position_dodge(0.8), 
-                     dot.size = 10) +
+                     dot.size = 20,
+                     size = 1.5) +
   scale_y_continuous() + 
   ggpubr::theme_pubr() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") +
-  stat_pvalue_manual(stattest_sy, label="p.signif", x = "Gene")
+  stat_pvalue_manual(stattest_sy, label="p.signif", x = "Gene") +
+    theme(axis.ticks.length=unit(0.1,"inch"),
+        axis.line = element_line(colour = 'black', size = 1.5),
+        axis.ticks = element_line(colour = "black", size = 2))+
+  scale_fill_manual(values = c("#fca361ff","#aac043ff")) +
+  scale_color_manual(values = c("#000000ff","#000000ff"))
+
 plot(nmd_plot)
@@
 
+## UNC13B fsE across NYGC cohort
+
+```{r do_stats}
+
+for_stats = clean_data_table %>% 
+    filter(tissue_clean %in%
+               c("Frontal_Cortex",
+                 "Lumbar_Spinal_Cord",
+                 "Cervical_Spinal_Cord",
+                 "Motor_Cortex",
+                 "Temporal_Cortex",
+                 "Cerebellum") )%>%
+    group_by(disease_group2,tissue_clean) %>%
+    mutate(n_sample = n_distinct(sample)) %>%
+    ungroup() %>%
+    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
+    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
+    mutate(tissue_clean = fct_relevel(tissue_clean, "Motor_Cortex",
+                                      "Frontal_Cortex",
+                                      "Temporal_Cortex",
+                                      "Cervical_Spinal_Cord",
+                                      "Lumbar_Spinal_Cord")) %>% 
+  mutate(all_the_fse_reads =  (((UNC13B_nmd_junction_5prime + UNC13B_nmd_junction_3prime)/ library_size) * 10^6)) %>% 
+  dplyr::select(tissue_clean,all_the_fse_reads,disease_group2) %>% 
+  group_by(tissue_clean) %>% 
+  nest()
+
+stars_table = for_stats %>% 
+  mutate(wilcox_test = map(data, ~{pairwise.wilcox.test(.x$all_the_fse_reads,.x$disease_group2) %>% broom::tidy()})) %>% 
+  dplyr::select(-data) %>% 
+  unnest() %>% 
+  ungroup() %>% 
+  mutate(stars = gtools::stars.pval(p.value = p.value)) %>% 
+  filter(p.value < 0.05) %>% 
+  mutate(y.position = seq(from = 1.5,to = 2.6,length.out = 11))
+
+```
+
+```{r plot_unc13b_fsE_nygc}
+clean_data_table %>% 
+    filter(tissue_clean %in%
+               c("Frontal_Cortex",
+                 "Lumbar_Spinal_Cord",
+                 "Cervical_Spinal_Cord",
+                 "Motor_Cortex",
+                 "Temporal_Cortex",
+                 "Cerebellum") )%>%
+    group_by(disease_group2,tissue_clean) %>%
+    mutate(n_sample = n_distinct(sample)) %>%
+    ungroup() %>%
+    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
+    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
+    mutate(tissue_clean = fct_relevel(tissue_clean, "Motor_Cortex",
+                                      "Frontal_Cortex",
+                                      "Temporal_Cortex",
+                                      "Cervical_Spinal_Cord",
+                                      "Lumbar_Spinal_Cord")) %>% 
+  mutate(all_the_fse_reads =  (((UNC13B_nmd_junction_5prime + UNC13B_nmd_junction_3prime)/ library_size) * 10^6)) %>% 
+  ggplot(aes(x = disease_group2,
+               y =all_the_fse_reads,
+               fill ="red")) +
+    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA) +
+    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
+    ggplot2::facet_wrap(~tissue_clean,scales = "free",nrow = 2) +
+    ylab("UNC13B fsE \n reads per million") +
+    theme(text = element_text(size = 12)) +
+    xlab("") +
+    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
+    theme_bw() +
+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+    theme(text = element_text(size = 18))  +
+    theme(legend.position = "bottom") + 
+  stat_pvalue_manual(stars_table, label = "{stars}")
+
+```
+
 ## UNC13A and UNC13B Intron Retention in NYGC
@@
                          "UNC13B-210")) %>%
-    ggplot(aes(x =condition, y = TPM,fill = Name.y)) + 
-    geom_boxplot() +
-   geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
+    ggplot(aes(x =condition, y = TPM,fill = Name.y),color = 'black') + 
+    geom_boxplot(size = 1.2,color = "black") +
+   geom_point(stroke = 1.2,size = 2.3,pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
     # facet_wrap(~plot_name) +
@@
 unc13b_expression = fread(file.path(here::here(),"data","unc13b_transcript_counts.csv"))
-
-clean_data_table %>% 
-    left_join(unc13b_expression) %>% 
-    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
-    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
-    select(sample,disease_group2,tissue_clean,unc13a_detected,TPM,Name) %>% 
-    filter(!is.na(Name)) %>% 
+unc13b_expression %>% 
+    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
+    filter(sample != "") %>% 
+    dplyr::select(sample,
+                  EffectiveLength,
+                  NumReads,
+                  Name.y) %>% 
+    unique() %>% 
+    left_join(clean_data_table[,.(sample,library_size)]) %>% 
+    filter(!is.na(library_size)) %>% 
+    mutate(TPM = (NumReads )/ (library_size / 10^6)) %>% 
+    unique() %>% 
+  left_join(clean_data_table %>% 
+    select(sample,disease_group2,tissue_clean)) %>% 
     unique() %>% 
@@
                                "Temporal_Cortex")) %>% 
-    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
+    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
       filter(Name.y %in% c("UNC13B-207",
@@
                          "UNC13B-210")) %>%
-    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
     ggplot(aes(x = disease_group2, y = TPM, fill = Name.y)) + 
@@
     theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) + 
-    scale_y_continuous(label=comma) +
     ggpubr::theme_pubr() + 
@@
 
-```
+# fractional
+
+df %>%
+    filter(classification != "Neither") %>%
+    group_by(name) %>%
+    mutate(f = n/sum(n)) %>% 
+  filter(grepl("FTD",name)) %>% 
+  ggplot(aes(x = name, y = 100*f, fill = classification)) +
+    geom_bar(stat="identity") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("") +
+  scale_fill_manual(values = c("#006613ff",
+                               "#ffbf00ff",
+                               "#f5671cff")) +
+    theme(axis.ticks.length=unit(0.1,"inch"),
+        axis.line = element_line(colour = 'black', size = 1.5),
+        axis.ticks = element_line(colour = "black", size = 2))
+
+ggplot(df %>% filter(type=="Patient"), 
+       aes(x = name, y = 100*fraction, fill = class2)) +
+    geom_bar(stat="identity", position="dodge",color = "black") +
+    ggpubr::theme_pubr() +
+    ggeasy::easy_rotate_x_labels(side = "right") +
+    ylab("Percentage of reads") +
+    ggtitle("Patient RNA") +
+    xlab("") +
+    ggeasy::easy_add_legend_title("") +
+    scale_fill_manual(values = c("#006613ff",
+                               "#ffbf00ff",
+                               "#f5671cff",
+                               "#290340ff")) +
+    theme(axis.ticks.length=unit(0.1,"inch"),
+        axis.line = element_line(colour = 'black', size = 1.5),
+        axis.ticks = element_line(colour = "black", size = 2))
+
+```
+
+# Different guides strength
+
+```{r}
+guide_strength = fread(file.path(here::here(),"data","different_guides_cryptic_expression.csv"))
+guide_strength %>% 
+  melt() %>% 
+  filter(`cell line` == "i5W") %>% 
+   mutate(condition = fct_relevel(condition,
+                              "control","tdp43_mild")) %>%
+    ggbarplot(,
+              x = "condition",
+              add = c("mean_se","jitter"),
+              y = "value",
+              fill = 'variable',
+              color = 'variable',
+              position = position_dodge(0.8)) +
+    ggpubr::theme_pubr() +
+      scale_color_manual(
+        values = c("#1C2617","#262114","#262114")
+    ) +
+        scale_fill_manual(
+        values = c("#1576b8","#eb7f3f","#975452")
+    ) +
+    scale_y_continuous(labels = scales::percent) +
+  ggtitle("NCRM5")
+
+guide_strength %>% 
+  melt() %>% 
+  filter(`cell line` == "i11w_repeat") %>% 
+   mutate(condition = fct_relevel(condition,
+                              "control","tdp43_mild")) %>%
+    ggbarplot(,
+              x = "condition",
+              add = c("mean_se","jitter"),
+              y = "value",
+              fill = 'variable',
+              color = 'variable',
+              position = position_dodge(0.8)) +
+    ggpubr::theme_pubr() +
+      scale_color_manual(
+        values = c("#1C2617","#262114","#262114")
+    ) +
+        scale_fill_manual(
+        values = c("#1576b8","#eb7f3f","#975452")
+    ) +
+    scale_y_continuous(labels = scales::percent) +
+  ggtitle("WTC11")
+```
+
+# Dox curves and protein/RNA levels
+
+```{r}
+library(tidyverse)
+library(viridis)
+library(dplyr)
+
+Protein <- read.csv(file.path(here::here(),"data","Protein_RNA_Corr_3.csv"), colClasses = c("factor","factor","numeric","factor","factor"))
+
+Protein$Treatment <- factor(Protein$Treatment, levels=c("NT","Dox-1","Dox-2","Dox-3","Dox-4","Dox-5"))
+Protein$Line_Group <- factor(Protein$Line_Group, levels=c("Protein","RNA","Correct_RNA"))
+
+agg=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="mean")
+standard_error <- function(x) sd(x) / sqrt(length(x))
+agg$se=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="standard_error")$Value
+
+limits <- aes(ymax=Value+se, ymin=Value-se)
+dodge <- position_dodge(width=0.9)
+
+agg = agg %>% 
+  filter(Gene != "13A-RNA") %>% 
+  mutate(Line_Group = gsub("Correct_RNA","RNA",Line_Group))
+
+ggplot(agg, aes(y=Value, x=Treatment)) +
+  geom_line(size = 2,mapping = aes(x=Treatment, y=Value,
+                          linetype=Line_Group, group=Gene,color = Color_Group), size=1)+
+  geom_errorbar(limits, 
+                position=dodge, 
+                width=0.24,
+                size = 1.1)+
+  xlab("") +
+  ylab("") +
+  ggpubr::theme_pubr()+
+  scale_y_continuous(labels = scales::percent_format(scale = 1))+
+  scale_colour_viridis_d(begin=0, end=1) +
+  labs(linetype = "") +
+  theme(axis.ticks.length=unit(0.1,"inch"),
+        axis.line = element_line(colour = 'black', size = 1.5),
+        axis.ticks = element_line(colour = "black", size = 2))
+
+```

### 2021-09-13 paper_figures.Rmd
Update paper_figures.Rmd

@@
   ggplot() +
-  geom_col(aes(x = detection_name, y = detection_rate,fill = detection_name),position = position_dodge2(width = 0.8, preserve = "single")) +
+  geom_col(aes(x = detection_name,
+               y = detection_rate,
+               fill = detection_name),color = 'black',position = position_dodge2(width = 0.8, preserve = "single")) +
   ggpubr::theme_pubr() +
@@
   theme(text = element_text(size = 18))  +
-  scale_fill_manual(values = c("#4C97B5","#4C97B5","#4C97B5","#4C97B5","#4C97B5","red","#C87156","#C87156")) +
+  scale_fill_manual(values = c("#00bd7dff","#00bd7dff","#00bd7dff","#00bd7dff","#00bd7dff","red","#e668f3ff","#e668f3ff")) +
   theme(legend.position = "none")
@@
   mutate(peak_rank = rank(frac_increase)) %>%
-  mutate(colour = ifelse(abs(pos-exon_snp)<region_length, "gold",
-                         ifelse(abs(pos-intron_snp)<region_length, "purple", "grey50")))
+  mutate(colour = ifelse(abs(pos-exon_snp)<region_length, "#f7e7abff",
+                         ifelse(abs(pos-intron_snp)<region_length, "#c8e1c4ff", "grey50")))
 
@@
 ggplot(peak_df, aes(x = peak_rank, y = 100*frac_increase, fill=colour)) +
-  geom_bar(stat="identity") +
+  geom_bar(stat="identity",,color = 'black') +
   scale_fill_identity() +
@@
   geom_errorbar(aes(ymin = 100*lower_bound, ymax = 100*upper_bound, x = peak_rank),
-                width = 0.3)
+                width = 0.3,size = 1.1)
 
@@
 guide_strength = fread(file.path(here::here(),"data","different_guides_cryptic_expression.csv"))
+
 guide_strength %>% 
@@
     scale_y_continuous(labels = scales::percent) +
-  ggtitle("NCRM5")
+  ggtitle("NCRM5") +
+         theme(axis.line = element_line(colour = 'black', size = 1.5),
+        axis.ticks = element_line(colour = "black", size = 2))
 
@@
     scale_y_continuous(labels = scales::percent) +
-  ggtitle("WTC11")
+  ggtitle("WTC11") +
+          axis.line = element_line(colour = 'black', size = 1.5),
+        axis.ticks = element_line(colour = "black", size = 2))
 ```

### 2021-09-17 paper_figures.Rmd

@@
 df <- read_csv(paste0(here(),
-                      "/data/itc/itc final values.csv")) %>%
-    group_by(RNA) %>%
-    mutate(sd =sd(kd)) %>%
-    mutate(mean_kd = mean(kd))
-
-ggplot(df, aes(x = RNA, y = kd)) +
-    geom_bar(aes(x = RNA, y = mean_kd), stat="identity", position="dodge") +
-    geom_point() +
-    geom_errorbar(aes(x=RNA, ymin = mean_kd-sd, ymax = mean_kd+sd),
-                  width = 0.3) +
+                      "/data/itc/itc final values.csv"))
+
+df2 <- df %>%
+    group_by(RNA, experiment) %>%
+    mutate(sd2 =ifelse(is.na(sd), sd(kd), sd)) %>%
+    mutate(mean_kd = mean(kd)) %>%
+  ungroup() %>%
+  select(experiment, RNA, mean_kd, sd2) %>%
+  unique()
+
+ggplot(df2, aes(x = experiment, fill = RNA, y = mean_kd)) +
+    geom_bar(stat="identity", position="dodge") +
+    #geom_point(aes(group=RNA), position="dodge") +
+    geom_errorbar(aes(x=experiment, ymin = mean_kd-sd2, ymax = mean_kd+sd2,
+                      fill = RNA), position="dodge") +
     ylim(0, NA) +
@@
 
-t.test(df$kd[which(df$RNA=="H")], df$kd[which(df$RNA=="R")])
+t.test(df$kd[which(df$RNA=="H" & df$experiment=="exon")], df$kd[which(df$RNA=="R" & df$experiment=="exon")])
+t.test(df$kd[which(df$RNA=="H" & df$experiment=="intron")], df$kd[which(df$RNA=="R" & df$experiment=="intron")])  # this is an overestimate of significance - however still borderline significant (~0.05) when fitting error is taken into account
 ```

### 2021-09-17 paper_figures.Rmd
update the index for rendering

@@
     theme(text = element_text(size = 18))  +
-    theme(legend.position = "bottom") + 
-    stat_compare_means(comparisons = comps)
+    theme(legend.position = "bottom") 
 ```
@@
 ```
-## Detection by genotype
-```{r}
-
-clean_data_table %>%
-  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
-  mutate(snp_status = case_when(rs12973192 == "C/C" ~ "healthy/healthy",
-                                rs12973192 == "C/G" ~ "healthy/risk",
-                                rs12973192 == "G/G" ~ "risk/risk")) %>% 
-  filter(disease_tissue == T) %>%
-  filter(junction_reads_stmn2 >= 30) %>%
-  filter(junction_reads_unc13a >= 30) %>%
-
-```
 
@@
 ```{r str and log2 graph}
-str_genotype = janitor::clean_names(fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/rs56041637.genotypes.377samples.txt")) %>% 
-  dplyr::rename(presumed_unique_id = number_sample_id)
+str_genotype = janitor::clean_names(fread(file.path(here::here(),"data/rs56041637.genotypes.377samples.txt"))) %>% 
+  dplyr::rename(participant_id = number_sample_id)
 
@@
 
-ipsc_de[gene_name %in% als_ftd_genomics][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique() %>% arrange(log2FoldChange) %>% clipr::write_clip()
-```
-
-## Linear model on UNC13A CE PSI with covariates
+ipsc_de[gene_name %in% als_ftd_genomics][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique() %>% arrange(log2FoldChange) %>% knitr::kable()
+```
+
+# Linear model on UNC13A CE PSI with covariates
 ```{r the_silly_linear}
@@
     mutate(broomed = map(lm_obj, broom::tidy)) %>% 
-    dplyr::select(tissue_clean,broomed) %>% unnest() %>% filter(p.value < 0.05) %>% View()
-
-nested %>% 
-    pluck("lm_obj", 1) %>% broom::tidy()
+    dplyr::select(tissue_clean,broomed) %>% 
+    unnest() %>% 
+    filter(p.value < 0.05) %>% knitr::kable()
+
+```
+
+```{r detection_seq_platform}
 unc13a_data_fisher = unc13a_sequencing_platform %>% 
@@
   ggtitle("WTC11") +
-          axis.line = element_line(colour = 'black', size = 1.5),
+           theme(axis.line = element_line(colour = 'black', size = 1.5),
         axis.ticks = element_line(colour = "black", size = 2))
@@
 agg=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="mean")
-standard_error <- function(x) sd(x) / sqrt(length(x))
-agg$se=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="standard_error")$Value
+std <- function(x) sd(x) / sqrt(length(x))
+
+tmp = Protein %>% 
+  group_by(Treatment,Gene,Line_Group,Color_Group) %>% 
+  summarise(se = std(Value))
 
@@
   filter(Gene != "13A-RNA") %>% 
-  mutate(Line_Group = gsub("Correct_RNA","RNA",Line_Group))
+  mutate(Line_Group = gsub("Correct_RNA","RNA",Line_Group)) %>% 
+  left_join(tmp)
 

### 2021-09-22 paper_figures.Rmd
Update paper_figures.Rmd
fix color

@@
     ylab("Kd / nM") +
-    xlab("")
+    xlab("") +
+  scale_fill_manual(values = c("#0c9bbdff","#ffb700ff"))
 

### 2021-09-22 paper_figures.Rmd

--- 2021-09-17 paper_figures.Rmd
@@
     geom_bar(stat="identity", position="dodge") +
-    #geom_point(aes(group=RNA), position="dodge") +
     geom_errorbar(aes(x=experiment, ymin = mean_kd-sd2, ymax = mean_kd+sd2,
@@
     ylim(0, NA)  +
+    geom_jitter(data = df, 
+                aes(group=RNA,x = experiment, y = kd,fill = RNA), 
+                position=position_jitterdodge(jitter.width = 0.1),pch = 21,size = 3,) +
     ggpubr::theme_pubr() +
     ylab("Kd / nM") +
-    xlab("")
-
-t.test(df$kd[which(df$RNA=="H" & df$experiment=="exon")], df$kd[which(df$RNA=="R" & df$experiment=="exon")])
-t.test(df$kd[which(df$RNA=="H" & df$experiment=="intron")], df$kd[which(df$RNA=="R" & df$experiment=="intron")])  # this is an overestimate of significance - however still borderline significant (~0.05) when fitting error is taken into account
-```
+    xlab("") +
+  scale_fill_manual(values = c("#0c9bbdff","#ffb700ff"))
+
+exon_pval = t.test(df$kd[which(df$RNA=="H" & df$experiment=="exon")], df$kd[which(df$RNA=="R" & df$experiment=="exon")]) %>% broom::tidy() %>% pull(p.value)
+intron_pval = t.test(df$kd[which(df$RNA=="H" & df$experiment=="intron")], df$kd[which(df$RNA=="R" & df$experiment=="intron")])  %>% broom::tidy() %>% pull(p.value) # this is an overestimate of significance - however still borderline significant (~0.05) when fitting error is taken into account
+```
+
+P-value two sample t-test:
+
+* RE - `r exon_pval`
+* RI - `r intron_pval`
+
+# Expression of ALS/FTD risk genes
 

### 2021-09-22 paper_figures.Rmd
added martina shsy5y iclip

@@
 ```
+
+# Endogenous SHSY5Y iCLIP
+
+```{r}
+# Read in filtered bed file (iMAPs output, filtered by script in pre-processing scripts folder)
+sh <- read_csv(paste0(here(), "/data/iCLIP/shsy5y_tdp43_iclip_unc13a_b_only.csv"))
+
+unc13a <- sh %>%
+  filter(chr == "chr19" & start > 17599327 & ((end.x < 17690344 & end.y == 0) | 
+                                                (end.y < 17690344 & end.x == 0))) %>%
+  select(start,total) %>%
+  full_join(data.frame(start=17599327:17690344)) %>%
+  arrange(start)
+
+unc13a[is.na(unc13a)] <- 0
+
+window <- 20
+p_a <- ggplot(unc13a, aes(x = start, y = zoo::rollmean(total, k = window, na.pad=T))) +
+  geom_line(alpha = 1) +
+  ylim(0, NA) +
+    # blue dashed line shows position of CE
+  geom_vline(xintercept = 17642430, linetype="dashed", alpha= 0.3, size=2, colour="blue") +
+    # red line shows position of retained intron
+  geom_vline(xintercept = 17628569, linetype="dashed", alpha= 0.3, size=2, colour="red") +
+  ggtitle("UNC13a") +
+  ylab(paste("Average XL density in", window, "nt window")) +
+  theme_minimal()
+
+unc13b <- sh %>%
+  filter(chr == "chr9" & start > 35160008 & ((end.x < 35407338 & end.y == 0) | 
+                                                (end.y < 35407338 & end.x == 0))) %>%
+  select(start,total) %>%
+  full_join(data.frame(start=35160008:35407338)) %>%
+  arrange(start)
+
+unc13b[is.na(unc13b)] <- 0
+
+p_b <- ggplot(unc13b, aes(x = start, y = zoo::rollmean(total, k = window, na.pad=T))) +
+  geom_line(alpha = 1) +
+  ylim(0, NA) +
+    # blue dashed line shows rough position of UG-repeat near fsE
+  geom_vline(xintercept = 35364641, linetype="dashed", alpha= 0.3, size=2, colour="blue") +
+  theme_minimal() +
+  ggtitle("UNC13B") +
+  ylab(paste("Average XL density in", window, "nt window"))
+
+p_a / p_b
+
+```

### 2021-09-22 preprocessing_scripts/iCLIP/filter_shsy5y_iclip_bed.R

@@ -1,0 +1,25 @@
+library(tidyverse)
+library(patchwork)
+
+get_bedgraph <- function(filename){
+  bg <- read_tsv(filename, col_names = c("chr", "start", "end", "score"))
+  return(bg)
+}
+
+bg1 <- get_bedgraph("shsy_1_full.bedgraph")
+bg2 <- get_bedgraph("shsy_2_full.bedgraph")
+
+combined <- full_join(bg1 %>% rename(score1 = score), 
+                      bg2 %>% rename(score2 = score),
+                      by=c("chr", "start"))
+
+combined[is.na(combined)] <- 0
+
+combined2 <- combined %>%
+  mutate(total =score1+score2 ) %>%
+  filter((chr == "chr19" & start > 17599327 & ((end.x < 17690344 & end.y == 0) | 
+                                                 (end.y < 17690344 & end.x == 0))) | 
+           chr == "chr9" & start > 35160008 & ((end.x < 35407338 & end.y == 0) | 
+                                                 (end.y < 35407338 & end.x == 0)))
+
+write_csv(combined2, "shsy5y_tdp43_iclip_unc13a_b_only.csv")

### 2021-09-28 paper_figures.Rmd
HERM

--- 2021-09-22 paper_figures.Rmd
@@
 
+WAFFLES 
 ```{css, echo = FALSE}
@@
   facet_wrap(~disease_group2+plot_name,scales = 'free') +
-  ggpubr::stat_compare_means(label = "p.format",vjust = 1.3) +
+  ggpubr::stat_compare_means(label = "p.signif",vjust = 1.3,
+hide.ns=TRUE) +
   ggpubr::theme_pubr() +
@@
 ```
+
+# Endogenous SHSY5Y iCLIP
+
+```{r}
+# Read in filtered bed file (iMAPs output, filtered by script in pre-processing scripts folder)
+sh <- read_csv(paste0(here(), "/data/iCLIP/shsy5y_tdp43_iclip_unc13a_b_only.csv"))
+
+unc13a <- sh %>%
+  filter(chr == "chr19" & start > 17599327 & ((end.x < 17690344 & end.y == 0) | 
+                                                (end.y < 17690344 & end.x == 0))) %>%
+  select(start,total) %>%
+  full_join(data.frame(start=17599327:17690344)) %>%
+  arrange(start)
+
+unc13a[is.na(unc13a)] <- 0
+
+window <- 20
+p_a <- ggplot(unc13a, aes(x = start, y = zoo::rollmean(total, k = window, na.pad=T))) +
+  geom_line(alpha = 1) +
+  ylim(0, NA) +
+    # red dashed line shows position of CE
+  geom_vline(xintercept = 17642430, linetype="dashed", alpha= 0.8, size=2, colour="red") +
+    # blue line shows position of retained intron
+  geom_vline(xintercept = 17628569, linetype="dashed", alpha= 0.8, size=2, colour="blue") +
+  ggtitle("UNC13A") +
+  ylab(paste("Average XL density in", window, "nt window")) +
+  theme_minimal() +
+  theme(panel.grid.minor = element_line(size = 0.5), 
+        panel.grid.major.x = element_line(size = 0.7,color = 'grey')) +
+    theme(axis.text = element_text(size = 28,color = "black"))
+
+unc13b <- sh %>%
+  filter(chr == "chr9" & start > 35160008 & ((end.x < 35407338 & end.y == 0) | 
+                                                (end.y < 35407338 & end.x == 0))) %>%
+  select(start,total) %>%
+  full_join(data.frame(start=35160008:35407338)) %>%
+  arrange(start)
+
+unc13b[is.na(unc13b)] <- 0
+
+p_b = ggplot(unc13b, aes(x = start, y = zoo::rollmean(total, k = window, na.pad=T))) +
+  geom_line(alpha = 1) +
+  ylim(0, NA) +
+    # red dashed line shows rough position of UG-repeat near fsE
+  geom_vline(xintercept = 35364641, linetype="dashed", alpha= 0.8, size=2, colour="red") + 
+      # blue dashed line shows start of intron
+
+  geom_vline(xintercept = 35384315, linetype="dashed", alpha= 0.8, size=2, colour="blue") +
+
+  theme_minimal() + 
+  theme(panel.grid.minor = element_line(size = 0.5), 
+        panel.grid.major.x = element_line(size = 0.7,color = 'grey')) +
+  ggtitle("UNC13B") +
+  ylab(paste("Average XL density in", window, "nt window")) +
+  theme(axis.text = element_text(size = 28,color = "black"))
+
+p_a + p_b
+
+```

### 2021-10-01 paper_figures.Rmd
Update paper_figures.Rmd

@@
 
-WAFFLES 
 ```{css, echo = FALSE}
@@
 p_load("data.table",
+       "ggnewscale",
        "tidyverse",
@@
 ```{r}
+std <- function(x) sd(x)/sqrt(length(x))
 
@@
 
-nmd_plot_chx = ggbarplot(for_bar_plot,
-                     x = 'Gene',
-                     add = c("mean_se","jitter"),
-                     y = 'Value',
-                     color = 'Treatment',
-                     position = position_dodge(0.8),
-                     dot.size = 10) +
-  scale_y_continuous() + 
-  ggpubr::theme_pubr() + stat_pvalue_manual(stattest, label="p.signif" , x = "Gene")
-plot(nmd_plot_chx)
+sum_for_bar_plot <- for_bar_plot %>% 
+  group_by(Gene,Treatment) %>% 
+  summarize(mean = mean(Value),sem = std(Value))
+
+ggplot(sum_for_bar_plot) +
+  geom_col(aes(x = Gene, fill = Treatment, y = mean),position = "dodge2",color = 'black') +
+  scale_fill_manual(values = c("#F9AC66","#B4C640")) +
+  new_scale_fill() +
+  geom_errorbar(aes(x = Gene, ymin = mean - sem, ymax = mean + sem),
+                position = "dodge2",size = 1.8) +
+  geom_point(data = for_bar_plot, aes(x = Gene, fill = Treatment, y = Value),
+             pch = 21, 
+             size = 4, stroke = 1.5,
+             position = position_jitterdodge(jitter.width = 0.2)) +
+  scale_fill_manual(values = c("#000000","#000000")) +
+  ggpubr::theme_pubr() + 
+  theme(axis.ticks.length=unit(0.1,"inch"),
+        axis.line = element_line(colour = 'black', size = 1.5),
+        axis.ticks = element_line(colour = "black", size = 2))
 
@@
 guide_strength = fread(file.path(here::here(),"data","different_guides_cryptic_expression.csv"))
-
-guide_strength %>% 
+std <- function(x) sd(x)/sqrt(length(x))
+
+ncrm5 <- guide_strength %>% 
   melt() %>% 
@@
    mutate(condition = fct_relevel(condition,
-                              "control","tdp43_mild")) %>%
-    ggbarplot(,
-              x = "condition",
-              add = c("mean_se","jitter"),
-              y = "value",
-              fill = 'variable',
-              color = 'variable',
-              position = position_dodge(0.8)) +
-    ggpubr::theme_pubr() +
-      scale_color_manual(
-        values = c("#1C2617","#262114","#262114")
-    ) +
-        scale_fill_manual(
-        values = c("#1576b8","#eb7f3f","#975452")
-    ) +
+                              "control","tdp43_mild"))
+sum_df_ncrm5 <- ncrm5 %>% 
+  group_by(condition, variable) %>% 
+  summarize(mean = mean(value),sem = std(value))
+
+ggplot(sum_df_ncrm5) + 
+  geom_col(aes(x = condition, fill = variable, y = mean),position = "dodge2",color = 'black',stroke = 1.5) +
+    scale_fill_manual(values = c("#1576b8","#eb7f3f","#975452")) +
+    geom_errorbar(aes(x = condition, ymin = mean - sem, ymax = mean + sem),
+                position = "dodge2",size = 1.8) +
+    new_scale_fill() +
+  geom_point(data = ncrm5, aes(x = condition, fill = variable, y = value),
+             pch = 21, 
+             size = 4, stroke = 1.5,
+             position = position_jitterdodge(jitter.width = 0.2))  +
+  scale_fill_manual(values = c("#000000","#000000","#000000")) +
+
+  ggpubr::theme_pubr() + 
   scale_y_continuous(labels = scales::percent) +
@@
 
-guide_strength %>% 
+wtc11 <- guide_strength %>% 
   melt() %>% 
@@
    mutate(condition = fct_relevel(condition,
-                              "control","tdp43_mild")) %>%
-    ggbarplot(,
-              x = "condition",
-              add = c("mean_se","jitter"),
-              y = "value",
-              fill = 'variable',
-              color = 'variable',
-              position = position_dodge(0.8)) +
-    ggpubr::theme_pubr() +
-      scale_color_manual(
-        values = c("#1C2617","#262114","#262114")
-    ) +
-        scale_fill_manual(
-        values = c("#1576b8","#eb7f3f","#975452")
-    ) +
+                              "control","tdp43_mild"))
+sum_df_wtc11 <- wtc11 %>% 
+  group_by(condition, variable) %>% 
+  summarize(mean = mean(value),sem = std(value))
+
+ggplot(sum_df_wtc11) + 
+  geom_col(aes(x = condition, fill = variable, y = mean),position = "dodge2",color = 'black',stroke = 1.5) +
+    scale_fill_manual(values = c("#1576b8","#eb7f3f","#975452")) +
+    geom_errorbar(aes(x = condition, ymin = mean - sem, ymax = mean + sem),
+                position = "dodge2",size = 1.8) +
+    new_scale_fill() +
+  geom_point(data = wtc11, aes(x = condition, fill = variable, y = value),
+             pch = 21, 
+             size = 4, stroke = 1.5,
+             position = position_jitterdodge(jitter.width = 0.2))  +
+  scale_fill_manual(values = c("#000000","#000000","#000000")) +
+
+  ggpubr::theme_pubr() + 
   scale_y_continuous(labels = scales::percent) +
@@
   scale_y_continuous(labels = scales::percent_format(scale = 1))+
-  scale_colour_viridis_d(begin=0, end=1) +
   labs(linetype = "") +
+  scale_color_manual(values = c("#E1BE6A","#CD5582","#A987E8"))+
   theme(axis.ticks.length=unit(0.1,"inch"),

### 2021-10-01 paper_figures.Rmd

@@
 plot(nmd_plot)
+
+```
+
+# lowdox + chx plot - from tapestation data
+
+```{r}
+
+lowdox_chx <- fread(file.path(here::here(),"data","band_plot.csv"))
+lowdox_chx <- data.frame(lowdox_chx)
+lowdox_chx$Condition <- factor(lowdox_chx$Condition, levels = c("Untreated", "Doxycycline", "DMSO", "CHX"))
+lowdox_chx$Band <- factor(lowdox_chx$Band, levels = c("Proper", "Short", "Long"))
+
+stattest <- lowdox_chx %>%
+  group_by(Band) %>%
+  pairwise_t_test(Value ~ Condition, ref.group = "CHX") %>%
+  add_significance() %>%
+  add_y_position(step.increase = 0.05) %>%
+  add_x_position()
+stattest <- filter(stattest, Band == "Proper")
+stattest
+
+nmd_plot_chx <- ggplot(lowdox_chx, aes(x = Condition, y = Value, color = Band, fill = Band)) + 
+  stat_summary(fun = mean, geom = "bar", aes(fill = Band), show.legend = F, position = position_dodge(width = 0.95)) + 
+  geom_point(size = 0.9, position = position_dodge(width = 0.95), color = "#000000") + 
+  stat_summary(fun.data = mean_se,  geom = "errorbar", width = 0.2, show.legend = F, position = position_dodge(width = 0.95), color = "#000000") +
+  stat_pvalue_manual(stattest) +
+  theme_classic() + theme(text = element_text(size = 24)) +
+  scale_color_manual(values = c("#1576B8","#EB7F3F","#975452", "#000000")) + 
+  scale_fill_manual(values = c("#1576B8","#EB7F3F","#975452", "#000000")) +
+  xlab("Condition") + ylab("PSI after treatment")
+plot(nmd_plot_chx)
 
